### 2023-06-20 paper/paper.tex

\documentclass{aa}

%
\usepackage{graphicx}
\usepackage[table]{xcolor}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{bm}
\usepackage{multirow, makecell}
\usepackage{booktabs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\begin{document}

\title{Unveiling the Optimal Approach for Neural Network-Based Cosmological Parameter Inference through Likelihood Implicit Methods}

\author{Denise Lanzieri \inst{1}
\and
Justine Zeghal \inst{2}
\and
Alexandre Boucaud \inst{2}
\and
Fran\c{c}ois Lanusse \inst{3}
\and
Jean-Luc Starck \inst{3}
}
\institute{Université Paris Cité, Université Paris-Saclay, CEA, CNRS, AIM, F-91191, Gif-sur-Yvette, France
\and
Université Paris Cité, CNRS, Astroparticule et Cosmologie, F-75013 Paris, France
\and
Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France
}
\titlerunning{}
\date{Received xxx; accepted xxx}

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {blabla}
  % aims heading (mandatory)
    {blabla}
  % methods heading (mandatory)
    {blabla}
  % results heading (mandatory)
   {blabla}
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{giant planet formation --
                $\kappa$-mechanism --
                stability of gas spheres
               }

   \maketitle
%
%-------------------------------------------------------------------

\section{Introduction}
\section{Theory Background}
\subsection{Weak lensing formalism}
\subsection{Lognormal Modeling}
For several cosmological purposes, the non-Gaussian field can be modeled as a Lognormal field \citep{coles1991lognormal,bohm2017bayesian}.
This model has the advantage of offering the rapid generation of the matter or the convergence field while allowing the extraction of information beyond the Gaussian one. 
Although several studies demonstrated that this model fails in describing the 3D field, it properly describes the 2D convergence field.

Assume a multivariate Gaussian distribution G described by the vector elements of mean $\mu_i$ and covariance matrix $\xi_g^{ij}$, we can define the density contrast $\delta(\bm{x})$ as a shifted lognormal random field :
% The distribution of the convergence $\kappa_i$ in the $i-$th redshift bin can be modeled as:
\begin{equation}
    \delta(\bm{x})=\lambda[e^{g(\bm{x})}-1]. 
\end{equation}\label{Eq:log_norm_kappa}
\begin{equation}\label{Eq:log_norm_kappa}
    \kappa_{i}=e^{y_i}-\lambda_i. 
\end{equation}
where $\lambda_i$ is a free parameter defining the shift of the lognormal distribution.  The density contrast $\delta(\bm{x})$ is fully defined by the shift parameter $\lambda$, the mean of the associated Gaussian field $\mu_i$, and its variance $\sigma_g^2\equiv \xi_g$. That fact that we require $<\delta>=0$ implies that the Gaussian field satisfies the following relation:
\begin{equation}
    \mu_g=-\frac{\sigma_g^2}{2}.
\end{equation}

The parameter $\lambda_i$, is also known as \textit{minimum convergence
parameter}, since defines the lowest values for all possible values $\kappa$ being the lognormal defined of for positive values. We can see from \autoref{Eq:log_norm_kappa} that th ereandom variables $\kappa_i$ is fully defined by the shift parameter $\lambda_i$, the mean of the associated gaussian variables $\mu_i$, and its variance $\sigma_i^2\equiv \xi_g^{ii}$.

 The modelling for the shift parameter can be approached in different ways.  determine
For example, $\lambda$ can be determined following the approach of \citet{xavier2016improving} , i.e.  matching moments of the distribution, or following the approach of \citet{hilbert2011cosmic}, i.e. fitting it as a free parameter. In general, it is a parameter dependent from the redshift, the cosmology, and the scale of the field at which the smoothing is applied.
In this work, we choose the derived $\lambda$ using Cosmomentum.
The lognormal $\xi^{ij}_{ln}$ and the associated Gaussian covariances $ \xi^{ij}_g$ are related via :
\begin{align}
    \xi^{ij}_{ln}(\theta) & \equiv \lambda_i \lambda_j (e^{ \xi^{ij}_g(\theta)}-1) \nonumber \\ 
    \xi^{ij}_g(\theta)&=\log{\left[ \frac{\xi^{ij}_{ln}(\theta)}{\lambda_i \lambda_j}+1\right ]}. \label{Eq:log_norm_corr}
\end{align}
%Note for Denise: The lambda_i and the \lambda_j in Eq:log_norm_corr are the same parameters defined in Eq:log_norm_kappa only for Gaussian distribution with mean zero. Otherwise, these parameters should be defined as alpha_i and alpha_j where alpha_i= <y>+lambda_i. See Xavier et al. 2015 for a detailed description.

One of the advantages of modeling the convergence field through a lognormal model is the fact that it enables the constraining of the convergence field in different redshift bins simultaneously while taking into account the correlation between the bins. Indeed, in the Fourier space, the power spectrum of $y$ is defined as:
\begin{equation}\label{Eq:log_norm_cls}
    C^{ij}_y(\ell)=2\pi \int_0^{\pi} d\theta \sin{\theta}P_{\ell}(\cos{\theta})\xi^{ij}_{y}(\theta)
\end{equation}
with $P_{\ell}$ Legendre polynomial with order $\ell$.

Having defined the survey to reproduce in terms of galaxy number density, redshifts, and shape noise, we can sample the Lognormal random map $\kappa$.  using \autoref{Eq:log_norm_kappa} and \autoref{Eq:log_norm_corr}.  First, we need to compute the theoretical auto $C_{\ell}^{i}$ and cross-correlated $C_{\ell}^{ij}$ angular power spectrum for each tomographic bin. These theoretical predictions are computed using the public library  \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-mosmo}}. Then the 1-D $C_{\ell}$ are projected into 2-D grids of the same size as the desired final convergence maps and the Gaussian covariances $\xi^{ij}_y(\theta)$ are computed through \autoref{Eq:log_norm_corr}. To take into account the cross-correlation between different redshift bins, we first stack together the vector $\xi^{ij}_y(\theta)$ to build a covariance matrix having the auto-correlation on the main diagonal and the cross-correlation off-diagonal. Then we perform an eigenvalues decomposition. This decomposition enables us to decorrelate the maps

\begin{equation}
    \bm{\Sigma}= 
    \begin{pmatrix}
    C_{\ell}^{11} & C_{\ell}^{12} & \cdots & C_{\ell}^{1n} \\
    C_{\ell}^{21} & C_{\ell}^{22} & \cdots & C_{\ell}^{2n} \\
    \vdots  & \vdots  & \ddots & \vdots  \\
    C_{\ell}^{n1} & C_{\ell}^{n2} & \cdots & C_{\ell}^{nn} 
    \end{pmatrix}
\end{equation}
%--------------------------------------------------------------------
\section{Simulations}
\subsection{Data generation}
Our analysis is based on a standard flat $\Lambda$CDM cosmological model, with the following parameters: the baryonic density fraction $\Omega_b$, the total matter density fraction $\Omega_m$, the Hubble parameter $h_0$, the spectral index $n_s$, the amplitude of the primordial power spectrum $\sigma_8$ and the dark energy parameter $w_0$. The priors used in the simulations and in inference process are listed in \autoref{tab:prior}, according to \citet{zhang2022transitioning}.
We compute the theoretical power and cross spectra using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023}. To calculate the 
lognormal shift parameter, we adopt the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which utilizes perturbation theory to compute the cosmology-dependent shift parameters. Therefore, the calculation of the shift parameters is based on a window function, while our pixels are rectangular. Following, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$  with $\Delta L$, the resolution of the pixels.  It is important to note that, as pointed out by \citet{boruah2022map}, this perturbation theory-based approach may not provide accurate results at small scales. However, since the main objective of this paper is to compare different compression strategies, we are not concerned about the potential implications of this approximation. In future applications of the presented inference pipeline, the simulation procedure will be based on N-body simulations.

Simulating the log-normal field for one single convergence map is a straightforward process. 
However, for a tomographic analysis, as described in the previous section, it is necessary to consider the cross-correlation between different tomographic bins. To account for this correlation, we employ the sampling strategy described in Section X. \\
Each map is reproduced on a regular grid with dimensions of $256^2$ pixels and covers an area of $10\times 10$ deg$^2$.

\begin{table}
	\centering
	\caption{ Prior and fiducial values used for the analyses. 
 The symbol $\mathcal{N}_T$ represents a Truncated Normal distribution. The lower bound of the support for the $\Omega_c$ distribution is set to zero, while the lower and upper bounds for the $w_0$ distribution are set to -2.0 and -0.3, respectively.}
	\begin{tabular}{lcc} 
		\hline \hline
		Parameter  & Prior & Fiducial value \\
		$\Omega_c$ & $\mathcal{N}_T$ (0.2664, 0.2) & 0.2664 \\
		$\Omega_b$ & $\mathcal{N}$ (0.0492, 0.006) & 0.0492 \\
		$\sigma_8$ & $\mathcal{N}$ (0.831, 0.14) & 0.831 \\
		$h$ & $\mathcal{N}$ (0.6727, 0.063) & 0.6727\\
		$n_s$ & $\mathcal{N}$ (0.9645, 0.08) & 0.9645 \\
		$w_{0}$ &  $\mathcal{N}_T$ (-1.0, 0.9) &  -1.0 \\
		\hline
	\end{tabular}
	\label{tab:prior}
\end{table}
%--------------------------------------------------------------------

\subsection{Noise and survey setting}
We conduct a tomographic study to reproduce the redshift distribution and the expected noise for the LSST-Y10 data release.
Specifically, following \citet{zhang2022transitioning}, we model the underlying redshift distribution using the parametrized Smail distribution \citep{smail1995deep}:
\begin{equation}
    n(z)=\propto z^2 \exp{-(z/z_0)^{\alpha}},
\end{equation}
with $z_0=0.11$ and $\alpha=0.68$ and assuming a photometric redshift error $\sigma_z=0.05(1+z)$ as defined in the LSST DESC Science Requirements Document (SRD, \citet{mandelbaum2018lsst}).
The galaxy sources are divided into 5 tomographic bins, each containing an equal number of galaxies. 
For each redshift bin, we assume Gaussian noise with mean zero and variance given by
 \begin{equation}
     \sigma^2_n= \frac{\sigma_e^2}{A_{pix}n_{gal}},
 \end{equation}
where we set the shape noise $\sigma_e = 0.26$ and the galaxy number density $n_{gal}=27$ arcmin$^{-2}$. Both the shape noise and galaxy number density values are obtained from SRD. The pixel area is given by $A_{pix}\approx$. 
Figure  xx illustrates the resulting source redshift distribution. \autoref{tab:survey_spec} provides a summary of the survey specifications.
\begin{table}
	\centering
	\caption{ LSST Y10 source galaxy specifications in our analysis. All values are based on the LSST DESC SRD.}
	\begin{tabular}{lc} 
             \hline \hline
		Redshift binning & 5 bins \\
		Redshift distribution ($z_{0}, \alpha$) & (0.11, 0.68)  \\
		Number density $n_s$ & 27/arcmin$^2$ \\
		Shape noise $\sigma_e$ & 0.26 \\
		Redshift error $\sigma_z$ &0.05(1+z)  \\
		\hline
	\end{tabular}
	\label{tab:survey_spec}
\end{table}
%--------------------------------------------------------------------
\section{Inference over forward simulation models}
\subsection{Map sampling}

%--------------------------------------------------------------------
\subsection{Explicit Inference}
\subsubsection{Power spectrum}
\subsubsection{Full field and HMC}
%--------------------------------------------------------------------
\subsection{Implicit Inference}

\subsubsection{Benchmark compression scheme}
There are several approaches available for training the compressor, and this section aims to provide an overview of these different methods used in previous works. 
In \autoref{Sec:results}, we will compare the outcomes of these various strategies. Specifically, using the same prior $p(\bm {\theta})$ for the cosmological parameter $\bm{\theta}$ and the simulator described in \autoref{Sec:results}, which samples $\bm {x}$ from $p(\bm{x}| \bm{\theta})$ and produces a set of mock observations $\bm{x_0}$, we will compare the posterior distribution $p(\bm{\theta}|\bm{x})$ obtained form each inference pipeline. This analysis will help to determine the optimal compression strategy, enabling us to extract the maximum amount of information possible given the fixed cosmological parameters.

\paragraph{\textcolor{violet}{Mean Square Error (MSE)}}
One of the most straightforward approaches adopted to train a Neural Network consists of minimizing the $L_2$ norm, or Mean Square Error (MSE). In this section, we demonstrate that this approach is equivalent to training the model to estimate the mean of the posterior distribution, namely:
\begin{equation}\label{Eq:mean_mse}
\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}=\operatorname*{argmin}_{\mathcal{F}(\bm{d})}\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\left\Vert \bm {\theta}-\mathcal{F}(\bm{d})
 \right \Vert_{2}^{2}].
\end{equation}
To demonstrate \autoref{Eq:mean_mse}, we need to minimize the expected value of the $L_2$ norm with respect to $\mathcal{F}(\bm{d})$. Let us consider its derivative:
\begin{align}\label{Eq:moment_1}
   & \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm{d}))}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2] =  \\
    &
    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2+\mathcal{F}(\bm{d})^2-2\bm {\theta}\mathcal{F}(\bm{d})] = \nonumber \\
    &
    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  [
    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2]+\mathcal{F}(\bm{d})^2-2\mathcal{F}(\bm{d})\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]]= \nonumber \\
    &2\mathcal{F}(\bm{d})-2 \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. \nonumber
\end{align}
Setting it equal to zero, we obtain the critical value:
\begin{equation}
    \mathcal{F}(\bm{d})= \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. 
\end{equation}
Considering the second-order derivative:
\begin{equation}\label{Eq:minimum_mean}
   \frac{\partial^2}{\partial^2 \mathcal{F}(\bm{d})}  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2]=2, 
\end{equation}
we can assert that the critical value $\mathcal{F(\bm(\theta))}$ is also a minimum. 
Since 
\begin{equation}
    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}|\bm{d}]= \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}
\end{equation}
it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Maximum Absolute Error (MAE)}}
Instead of minimizing the $L_2$ norm, these approaches focus on the $L_1$ norm or Mean Absolute Error (MAE). In this section, we demonstrate that this is equivalent to training the model to estimate the median of the posterior distribution.
By definition, the median of a probability density function $p(x)$ is a real number $m$ that satisfies:
\begin{equation}\label{Eq:definition_median}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
\end{equation}. 
The expectation value of the mean absolute error is defined as:
\begin{equation}
    \mathbb{E}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
\end{equation}
which can be decomposed as
\begin{equation}
        \int_{\infty}^{m}p(x)|x-m|dx +\int_{m}^{\infty}p(x)|x-m|dx .
\end{equation}
To minimize this function with respect to $m$, we need to compute its derivative:
\begin{equation}\label{Eq:absolute_median}
    \frac{d\mathbb{E}[|x-m|]}{dm}=
    \frac{d}{dm}\int_{\infty}^{m}p(x)|x-m|dx +\frac{d}{dm}\int_{m}^{\infty}p(x)|x-m|dx. 
\end{equation}
Considering that $|x-m|=(x-m)$ for $m\le x$ and $|x-m|=(m-x)$ $m\ge x$, 
we can write \autoref{Eq:absolute_median} as:
\begin{equation}
    \frac{d\mathbb{E}[|x-m|]}{dm}=
    \frac{d}{dm}\int_{\infty}^{m}p(x)(m-x)dx +\frac{d}{dm}\int_{m}^{\infty}p(x)(x-m)dx .
\end{equation}
Using the Leibniz integral rule, we get:
\begin{align}
    &\frac{d\mathbb{E}[|x-m|]}{dm}= \\
    &
    p(x)(m-m)\frac{dm}{dm}+\int_{\infty}^{m}\frac{\partial}{\partial m}[p(x)(m-x)]dx + \nonumber \\
    & - p(x)(m-m)\frac{dm}{dm}+\int_{m}^{\infty}\frac{\partial}{\partial m}[p(x)(m-x)]dx \nonumber .
\end{align}
Setting the derivative to zero, we obtain:
\begin{equation}
    \frac{d\mathbb{E}[|x-m|]}{dm}= \int_{\infty}^{m} p(x)dx-\int_{m}^{\infty}p(x)dx =0.
\end{equation}
Thus,
\begin{equation}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx .
\end{equation}
Considering that
\begin{equation}
\int_{\infty}^{m} p(x)dx+\int_{m}^{\infty}p(x)dx=1,
\end{equation}
we obtain 
\begin{equation}\label{Eq:definition_median}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
\end{equation}
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Gaussian Negative Log-Likelihood (GNLL)}}
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Information Maximising Neural Networks (IMNNs)}}
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Variational Mutual Information Maximization (VMIM)}}
The Variational Mutual Information Maximization (VMIM) has been used for cosmological inference problems for the first time by \citet{jeffrey2021likelihood}. This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm t$ by training a neural network with parameters $\phi$. 
Mathematically, it is defined as follows:
\begin{align}\label{Eq:mutual_information}
    I(\bm{t}, \bm {\theta}) &= D_{KL}(p(\bm {t}, \bm {\theta})||p(\bm {t})p(\bm {\theta})) \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
        &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta}  d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- \mathbb{E}_{p(\bm {\theta})} [\log{p(\bm {\theta})}] \\ \nonumber
    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- H(\bm {\theta});
\end{align}

where $D_{KL}$ is the Kullback-Leibler divergence \citep{kullback1951information}, $p(\bm {t}, \bm {\theta})$ is the joint probability distribution of summary statistics and cosmological parameters, and $H(\bm {\theta})$ represents the \textit{entropy} of the distribution of cosmological parameters.  
In other words, the mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
Given the original high-dimensional data vector $\bm d$, the information $\bm t$ is extracted using $\bm {t}=F_{\bm {\phi}}(\bm {d})$. 
The goal is to find the parameters $\bm {\phi}$ that maximize the mutual information between the summary and cosmological parameters:
\begin{equation}
   \bm {\phi}^*= \operatorname*{argmax}_{\bm {\phi}} I(F_{\bm {\phi}}(\bm{d}), \bm {\theta}).
\end{equation}
However, the mutual information expressed in \autoref{Eq:mutual_information} is not tractable. To overcome this limitation,  several approaches have been developed, relying on tractable bounds that allow the training of a deep neural network to optimize the mutual information. In this work, we adopt the same strategy used by \citet{jeffrey2021likelihood} and use the variational lower bound \citep{barber2003information}:
\begin{equation}\label{Eq:variational_lower_bound}
    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\phi}')}]- H(\bm {\theta}).
\end{equation}
Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t} ; \bm{\phi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
Since the entropy of the cosmological parameters is constant for a fixed training set, the optimization problem based on the lower bound in  \autoref{Eq:variational_lower_bound} can be written as:
\begin{equation}
    \operatorname*{argmax}_{\bm {\phi}, \bm {\phi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\phi}}(\bm {d}) ; \bm{\phi}')}].
\end{equation}

\subsubsection{Inference strategy}
%--------------------------------------------------------------------

\begin{center}
\begin{table*}
\begin{tabular}{ |p{3.5cm}|p{2cm}|p{3cm}|p{2.5cm}|p{4cm}|  }
 \hline
Reference & \makecell{Architecture\\ compressor}  & \makecell{Loss function} & \makecell{Inference \\ strategy} & \makecell{Output compressor}   \\
 \hline
            \citet{2018PhRvD..97j3515G} & \makecell{CNN} & \makecell{MAE} &  \makecell{Likelihood \\ analysis} &\makecell{$\sigma_8, \Omega_m$}  \\
 \hline
            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
             \\
            $\tanh^{-1}{(\text{Corr}(\sigma_8,\Omega_m))}$,
             \\
            $\sigma_8,\Omega_m$} 
\\
\hline     
\rowcolor{lightgray}
            \citet{fluri2019cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m, A_{IA}/10, L$  \\ with  $L:\Sigma^{-1}=LL^{T}$}
\\
\hline            
            \citet{ribli2018improved} & \makecell{CNN} & \makecell{MSE}  &\makecell{RMSE for \\ evaluation}  & \makecell{$\sigma_8, \Omega_m$}   
\\
\hline            
            \citet{ribli2019weak} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
\\            
\hline             
            \citet{PhysRevD.102.123506} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
\\
\hline 
\rowcolor{lightgray}
            \citet{jeffrey2021likelihood} & \makecell{\makecell{CNN} \\\makecell{CNN}+NF} & 
            \makecell{MSE \\ VMIM}
            & \makecell{PyDelfi} & \makecell{$\phi: F_{\phi}(\bm{d})=\bm{\theta}$ \\ with $\bm{\theta}=\Omega_m, \sigma_8$}    
\\            
\hline             
            \citet{fluri2021cosmological} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
\\
\hline      
\rowcolor{lightgray}

            \citet{fluri2022full} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
\\            
\hline 
            \citet{lu2022simultaneously} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8, A_{IA}/10,$ \\ rescaled \\ baryonic parameters \\ $(M_c,M_{1,0}, \eta, \beta)$}   
\\           
\hline 
            \citet{kacprzak2022deeplss} & \makecell{CNN} & \makecell{GNLL}  & \makecell{Likelihood \\ analysis}  & \makecell{WL: $\Sigma, \Omega_m, \sigma_8, A_{IA}, \eta_{IA}$ \\  GC:$\Sigma, \Omega_m, \sigma_8, b_g, r_g,\eta_{b_{g}}$ }

\\
\hline 
\rowcolor{lightgray}
            \citet{lu2023cosmological} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8,A_{IA}/10,$\\ rescaled baryonic \\ parameters: $M_c,M_{1,0}, \eta, \beta$} 
\end{tabular}
\label{tab:biblio_survey}
\end{table*}
\end{center}

\citep{Campagne_2023}
%--------------------------------------------------------------------
\section{Results}\label{Sec:results}
%--------------------------------------------------------------------
\section{Conclusions}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\begin{acknowledgements}
This work was granted access to the HPC/AI resources of IDRIS under the allocation 2022-AD011013922 made by GENCI.
\end{acknowledgements}

\bibliographystyle{aa} % style aa.bst
\bibliography{paper/biblio} 
\end{document}

### 2023-06-23 paper/paper.tex
23/06

@@
 \title{Unveiling the Optimal Approach for Neural Network-Based Cosmological Parameter Inference through Likelihood Implicit Methods}
+
+\newcommand{\justine}[1]{{\color{cyan}JZ: #1}}
+\newcommand{\denise}[1]{{\color{red}DL: #1}}
 
@@
 \end{equation}
-
 The parameter $\lambda_i$, is also known as \textit{minimum convergence
@@
 The priors used in the simulations and in inference process are listed in \autoref{tab:prior}, according to \citet{zhang2022transitioning}.
-We compute the theoretical power and cross spectra using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023}.
+The simulation process proceeds as follows.
+Given the cosmological parameters, we compute the theoretical power- and cross- spectra using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023}.
 To calculate the
 lognormal shift parameter, we adopt the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which utilizes perturbation theory to compute the cosmology-dependent shift parameters.
-Therefore, the calculation of the shift parameters is based on a window function, while our pixels are rectangular.
-Following, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$  with $\Delta L$, the resolution of the pixels.
-It is important to note that, as pointed out by \citet{boruah2022map}, this perturbation theory-based approach may not provide accurate results at small scales.
+It should be noted that the calculation of the shift parameters is based on a window function, while our pixels are rectangular.
+Following \citet{boruah2022map}, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$  with $\Delta L$, the resolution of the pixels.
+Moreover, as pointed out by \citet{boruah2022map}, this perturbation theory-based approach may not provide accurate results at small scales.
 However, since the main objective of this paper is to compare different compression strategies, we are not concerned about the potential implications of this approximation.
@@
 \end{table}
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/redshift_distribution_light.pdf}
+    \caption{
+     Source sample redshift distributions for each tomographic bin for LSST Y10. The number density on the y-axis is shown in arcmin$^2$.
+    }
+     \label{fig:redshift_distribution}
+\end{figure}
 %--------------------------------------------------------------------
 \section{Inference over forward simulation models}
-\subsection{Map sampling}
+With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, enable to fully account for the non-Gaussian information in the lensing field at
+the scales that future surveys will be able to access.
+One of the goals of this paper is to present a forward model of the full observables to directly extract information from the raw pixel data, instead of attempting to evaluate analytically the likelihood of any summary statistics and in doing so, preserve all the available information.
+This approach has the advantage of easily incorporating systematic effects or easily combining multiple cosmological probes by joint simulations.  
+In this context, the simulator of the observables becomes our physical model, in which every single component is tractable.
+These models, also referred to as \textit{probabilistic program}, take as input the parameters $\bm{\theta}$, sample some \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, and generates the output $\bm x \sim p(\bm d|\bm \theta)$, with $\bm{d}$ observations.
+At this point, we can infer both the latent variables or the parameter $\bm{\theta}$.
+However,  the marginal likelihood is typically intractable:
+\begin{equation}
+    p(\bm{x}|\bm{\theta})=\int p(\bm{x},\bm{z}|\bm{\theta}) d\bm{z}=\int p(\bm{x}|\bm{z},\bm{\theta})p(\bm{z}|\bm{\theta}) d\bm{z}.
+\end{equation}
+To bypass this limitation, while keeping exploring the full information content of data, we have different approaches.
+Although in literature these approaches are usually referred to in different ways, in this work hereafter we will make the following distinction
+\paragraph{\textbf{Explicit inference}} referring to all Likelihood-based inference approaches. 
+In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Model (BHM).
+As a Bayesian forward approach it involves using a given model to predict observations and then comparing these predictions with real observations to infer the parameters of the model.
+The \textit{hierarchical} nature of this methodology comes from the fact that a complex inference task, such as the weak lensing inference task, can be broken down into several hierarchies of elements, which can be understood and from which we can sample from \citep{heavens2018bayesian}.
+However, performing a global analysis on complex and large datasets, such as what could be a lensing dataset, is not feasible.
+Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
+This approach involves treating the simulator as a probabilistic model and performing inference over the joint posterior:
+\begin{equation}
+        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x},\bm{z}|\bm{\theta}) p(\bm{z}|\bm{\theta}).
+\end{equation}
+\paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations.
+This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution
+\begin{equation}
+    (\bm{x}, \bm{\theta})\sim p(\bm{x}, \bm{\theta}).
+\end{equation}
+Within this class of methods, we can distinguish between more traditional methods such as the Approximate Bayesian Computation (ABC), a rejection-criteria based approach, which approximate the likelihood comparing the simulations with data, and the Density Estimation Likelihood Free Inference (DELFI), methods, where the inference task is approached as a density estimation problem.
 
+Ultimately, for a given simulation model, the two approaches should converge to the same posterior.
 %--------------------------------------------------------------------
@@
 \subsubsection{Power spectrum}
+Although this is not a full-field approach, we aim to compare the latest to a more standard approach. 
+To obtain the probability distribution of the cosmological parameters using a summary-statistics-methods, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
+The correlation between two generic tomographic bins $i$ and $j$ can be expressed as:
+\begin{equation}
+    \left \langle \kappa^{i}(\ell)\kappa^{j*}(\ell)\right \rangle =(2\pi)^2\delta_D(\ell-\ell')C_{\ell}^{ij},
+\end{equation}
+We assume a Gaussian likelihood with a cosmology-independent covariance matrix:
+\begin{equation}
+    \log{\mathcal{L}}(\bm{\theta})=-\frac{1}{2}[\bm{d}-\bm{\mu}(\bm{\theta})]^{T}\bm{C}^{-1}[\bm{d}-\bm{\mu}(\bm{\theta})].
+\end{equation}
+To compute the expected theoretical predictions $\bm{\mu}(\bm{\theta})$ we use the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}. 
+ The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, shown in \autoref{tab:prior}, using the same theoretical library.
+In {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
+\begin{equation}
+    \text{Cov}(C_{\ell},C_{\ell}')=\frac{1}{f_{sky}(2 \ell+1)}\left(C_{\ell}+\frac{\sigma_{\epsilon}^2}{2n_s}\right)
+\end{equation}
+where $f_{sky}$ is the fraction of sky observed by the survey, and $n_s$ is the number density of galaxies. 
+To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map.
 \subsubsection{Full field and HMC}
+The following steps outline our methods.
+We draw cosmological parameters from the priors illustrated in \autoref{tab:prior}.
+Given values for the cosmological parameters, we generate the log-normal field as described in the previous sections.
+So, in other words, the simulator becomes the physical model generating a non-linear representation of the convergence map.
+The measure of the convergence for each pixel and for each bin predicted from the physical model will differ from the ones from real observations due to the noise.
+This is taken into account in the likelihood.
+Specifically, for LSST Y10 the number of galaxies for each pixel should be sufficiently high that for the central limit theorem, we can assume that the observation is characterized by a Gaussian noise $\sigma_n^2=\sigma_e^2/N_s$, with $N_s$ the total number of source galaxies per bin and pixel.
+Given $\sigma_n^2$ the variance of this Gaussian likelihood, its form in the log form can be expressed as:
+\begin{equation}
+    \log{\mathcal{L}}(\bm{\theta})=
+    \sum_i^{N_{pix}} \sum_{j}^{N_{bins}} \log{P(\kappa^{obs}_{i,j}|\kappa_{i,j},\bm{\theta})}
+    =\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2}
+\end{equation}
+Not relying on any summary statistics, the full-field approach typically results in a high-dimensional problem, requiring more sophisticated statistical techniques.
+To sample the posterior distribution for $\bm{\theta}$, we use a Hamiltonian Monte Carlo (HMC) algorithm.
+The HMC algorithm is particularly helpful in high-dimensional spaces where a large number of steps are required to effectively explore the space.
+It improves the sampling process by using the information contained in the gradients to guide the sampling process.
+Since the code is implemented in Jax, the gradients are accessible via automatic differentiation.
+
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
+A third possible scenario consists in minimizing the Negative Log Likelihood.
+First of all, let us stating that minimizing the negative log likelihood is equivalent to the Maximum Likelihood Estimation (MLE).
+The proof of this statement is straightforward.
+As suggested by the name, this optimization problem involves maximizing the likelihood distribution $p(\bm{x}|\bm{\theta})$
+\begin{equation}
+    \bm{\theta}_{ML}= \operatorname*{argmax}_{\bm{\theta}}p(\bm{x}|\bm{\theta}).
+\end{equation}
+where
+\begin{equation}
+    p(\bm{x}|\bm{\theta})= \prod_{i-1}^n p(x_i|\bm{\theta}).
+\end{equation}
+Applying the log function in this context is done for several reasons, e.g.
+reducing the potential of underflow when the likelihoods are small or allowing the conversion of a product into a summation.
+Then, remembering that $-\log$ is strictly decreasing, the following are equivalent:
+\begin{equation}
+    \operatorname*{argmax}_{\bm{\theta}}\sum_{i-1}^n \log{p(x_i|\bm{\theta})}= \operatorname*{argmin}_{\bm{\theta}}-\sum_{i-1}^n \log{p(x_i|\bm{\theta})}.
+\end{equation}
+
 \paragraph{\textcolor{violet}{Information Maximising Neural Networks (IMNNs)}}
@@
 %--------------------------------------------------------------------
+To evaluate compression strategies, we propose a comparison of posteriors $p(\theta|s(x))$, where $s(x)$ is the summary statistic.
+Indeed, the more informative the summary statistics, the closer the posterior $p(\theta | s(x))$ is to the true posterior $p (\theta | x)$.
+To this end, we propose to do Neural Posterior Estimation (NPE) that aims to directly learn from simulations the posterior distribution
+
+\begin{equation}
+    p(\theta | x) \propto \int p(x | \theta, z) p(z|\theta) p(\theta) dz\,.
+\end{equation}
+
+To achieve this, we train a  conditional Normalizing Flow (NF) $p_{\varphi} (\theta | s(x)) $ by minimizing the negative log-probability in $(\theta, s(x))$ space:
+
+\begin{equation}
+    \mathcal{L}_{\rm NLL} = \mathbb{E}_{p(\theta, s(x))} \left[ - \log p_{\varphi} (\theta | s(x))  \right],
+    \label{eq:nll}
+\end{equation}
+
+By learning the posterior distribution for each statistic, we capture the significance of each statistic. This approach allows us to use the posterior distribution as a fair benchmark for comparison.
+In particular, for each statistic we use the same NF architecture which is a RealNVP (REF) with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ (REF).
+
+% talk about evaluation of approx posterior at a given obs
 
@@
 
-\citep{Campagne_2023}
 %--------------------------------------------------------------------

### 2023-06-25 paper/paper.tex
25/06

@@
 \newcommand{\denise}[1]{{\color{red}DL: #1}}
+\newcommand{\FL}[1]{{\color{magenta}FL: #1}}
 
@@
    \maketitle
-%
+
 %-------------------------------------------------------------------
-
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
 \section{Introduction}
-\section{Theory Background}
-\subsection{Weak lensing formalism}
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+\section{Inference over forward simulation models}
+With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, enable to fully account for the non-Gaussian information in the lensing field at
+the scales that future surveys will be able to access.
+One of the goals of this paper is to present a forward model of the full observables to directly extract information from the raw pixel data, instead of attempting to evaluate analytically the likelihood of any summary statistics and in doing so, preserve all the available information.
+This approach has the advantage of easily incorporating systematic effects or easily combining multiple cosmological probes by joint simulations.  
+In this context, the simulator of the observables becomes our physical model, in which every single component is tractable.
+These models, also referred to as \textit{probabilistic program}, take as input the parameters $\bm{\theta}$, sample some \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, and generates the output $\bm x \sim p(\bm d|\bm \theta)$, with $\bm{d}$ observations.
+At this point, we can infer both the latent variables or the parameter $\bm{\theta}$.
+However,  the marginal likelihood is typically intractable:
+\begin{equation}
+    p(\bm{x}|\bm{\theta})=\int p(\bm{x},\bm{z}|\bm{\theta}) d\bm{z}=\int p(\bm{x}|\bm{z},\bm{\theta})p(\bm{z}|\bm{\theta}) d\bm{z}.
+\end{equation}
+To bypass this limitation, while keeping exploring the full information content of data, we have different approaches.
+Although in literature these approaches are usually referred to in different ways, in this work hereafter we will make the following distinction
+%--------------------------------------------------------------------
+\paragraph{\textbf{Explicit inference}} referring to all Likelihood-based inference approaches. 
+In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Model (BHM).
+As a Bayesian forward approach it involves using a given model to predict observations and then comparing these predictions with real observations to infer the parameters of the model.
+The \textit{hierarchical} nature of this methodology comes from the fact that a complex inference task, such as the weak lensing inference task, can be broken down into several hierarchies of elements, which can be understood and from which we can sample from \citep{heavens2018bayesian}.
+However, performing a global analysis on complex and large datasets, such as what could be a lensing dataset, is not feasible.
+Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
+This approach involves treating the simulator as a probabilistic model and performing inference over the joint posterior:
+\begin{equation}
+        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x},\bm{z}|\bm{\theta}) p(\bm{z}|\bm{\theta}).
+\end{equation}
+%--------------------------------------------------------------------
+\paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations.
+This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution
+\begin{equation}
+    (\bm{x}, \bm{\theta})\sim p(\bm{x}, \bm{\theta}).
+\end{equation}
+Within this class of methods, we can distinguish between more traditional methods such as the Approximate Bayesian Computation (ABC), a rejection-criteria based approach, which approximate the likelihood comparing the simulations with data, and the Density Estimation Likelihood Free Inference (DELFI), methods, where the inference task is approached as a density estimation problem.
+Ultimately, for a given simulation model, the two approaches should converge to the same posterior.
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+\section{The SBILens framework}
 \subsection{Lognormal Modeling}
@@
 First, we need to compute the theoretical auto $C_{\ell}^{i}$ and cross-correlated $C_{\ell}^{ij}$ angular power spectrum for each tomographic bin.
-These theoretical predictions are computed using the public library  \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-mosmo}}.
+These theoretical predictions are computed using the public library  \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}.
 Then the 1-D $C_{\ell}$ are projected into 2-D grids of the same size as the desired final convergence maps and the Gaussian covariances $\xi^{ij}_y(\theta)$ are computed through \autoref{Eq:log_norm_corr}.
@@
 \end{equation}
+%------------------------------------------------------------------
+%------------------------------------------------------------------
 %--------------------------------------------------------------------
-\section{Simulations}
 \subsection{Data generation}
@@
 \end{table}
-%--------------------------------------------------------------------
-
+%------------------------------------------------------------------
+%------------------------------------------------------------------
+%------------------------------------------------------------------
 \subsection{Noise and survey setting}
@@
 %--------------------------------------------------------------------
-\section{Inference over forward simulation models}
-With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, enable to fully account for the non-Gaussian information in the lensing field at
-the scales that future surveys will be able to access.
-One of the goals of this paper is to present a forward model of the full observables to directly extract information from the raw pixel data, instead of attempting to evaluate analytically the likelihood of any summary statistics and in doing so, preserve all the available information.
-This approach has the advantage of easily incorporating systematic effects or easily combining multiple cosmological probes by joint simulations.  
-In this context, the simulator of the observables becomes our physical model, in which every single component is tractable.
-These models, also referred to as \textit{probabilistic program}, take as input the parameters $\bm{\theta}$, sample some \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, and generates the output $\bm x \sim p(\bm d|\bm \theta)$, with $\bm{d}$ observations.
-At this point, we can infer both the latent variables or the parameter $\bm{\theta}$.
-However,  the marginal likelihood is typically intractable:
-\begin{equation}
-    p(\bm{x}|\bm{\theta})=\int p(\bm{x},\bm{z}|\bm{\theta}) d\bm{z}=\int p(\bm{x}|\bm{z},\bm{\theta})p(\bm{z}|\bm{\theta}) d\bm{z}.
-\end{equation}
-To bypass this limitation, while keeping exploring the full information content of data, we have different approaches.
-Although in literature these approaches are usually referred to in different ways, in this work hereafter we will make the following distinction
-\paragraph{\textbf{Explicit inference}} referring to all Likelihood-based inference approaches. 
-In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Model (BHM).
-As a Bayesian forward approach it involves using a given model to predict observations and then comparing these predictions with real observations to infer the parameters of the model.
-The \textit{hierarchical} nature of this methodology comes from the fact that a complex inference task, such as the weak lensing inference task, can be broken down into several hierarchies of elements, which can be understood and from which we can sample from \citep{heavens2018bayesian}.
-However, performing a global analysis on complex and large datasets, such as what could be a lensing dataset, is not feasible.
-Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
-This approach involves treating the simulator as a probabilistic model and performing inference over the joint posterior:
-\begin{equation}
-        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x},\bm{z}|\bm{\theta}) p(\bm{z}|\bm{\theta}).
-\end{equation}
-\paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations.
-This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution
-\begin{equation}
-    (\bm{x}, \bm{\theta})\sim p(\bm{x}, \bm{\theta}).
-\end{equation}
-Within this class of methods, we can distinguish between more traditional methods such as the Approximate Bayesian Computation (ABC), a rejection-criteria based approach, which approximate the likelihood comparing the simulations with data, and the Density Estimation Likelihood Free Inference (DELFI), methods, where the inference task is approached as a density estimation problem.
-
-Ultimately, for a given simulation model, the two approaches should converge to the same posterior.
 %--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%------------------------------------------------------------------
+\section{Experiment}
 \subsection{Explicit Inference}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 \subsubsection{Power spectrum}
@@
 To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map.
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 \subsubsection{Full field and HMC}
@@
 Since the code is implemented in Jax, the gradients are accessible via automatic differentiation.
-
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
 \subsection{Implicit Inference}
-
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+\subsubsection{Need for data compression}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 \subsubsection{Benchmark compression scheme}
@@
 This analysis will help to determine the optimal compression strategy, enabling us to extract the maximum amount of information possible given the fixed cosmological parameters.
-
+%--------------------------------------------------------------------
 \paragraph{\textcolor{violet}{Mean Square Error (MSE)}}
@@
 %--------------------------------------------------------------------
-A third possible scenario consists in minimizing the Negative Log Likelihood.
+\FL{I think we should have this section after the VMIM}
+
+\FL{Recognizing that the uncertainty on different cosmological parameters will vary, a third
+class of inverse variance weighted MSE was proposed in \cite{fluri2018cosmological} with the idea of
+ensuring that each parameter contributes fairly to the overall loss by taking into account its uncertainty. The loss function typically takes the following form:
+\begin{equation}
+    \mathcal{L} = \frac{1}{2} \log(|\Sigma|) + \frac{1}{2}(t - \theta)^t \Sigma^{-1} (t - \theta) 
+\end{equation}
+where $t$ is the summary statistics and $\Sigma$ is the covariance matrix of representing the uncertainty on the cosmological parameters $\theta$.
+Both $t$ and $\Sigma$ can be outputs of the compression network, i.e.
+$f_\theta(x)=(t, \Sigma)$.
+
+One recognizes here the expression of a Gaussian probability function, and this expression can thus be straightforwardly related to the VMIM case by simply assuming a Gaussian distribution as the variational approximation for the posterior $q(\theta | x) = \mathcal{N}(\theta; t, \Sigma)$.
+
+This means that this loss function is actually sub-optimal for two reasons:
+\begin{itemize}
+    \item The summary extracted by the neural network $t$ is, similarly to the MSE case only an estimate of the mean of 
+    the posterior distribution, which is not guaranteed to be sufficient.
+    \item Because of the Gaussian variational assumption, the variational posterior can potentially be biased with respect to the true posterior, and thus the mean of this Gaussian approximation may be biased with respect to the true posterior mean.
+\end{itemize}
+ A simple MSE would thus seem to be preferable.
+}
+% A third possible scenario consists in minimizing the Negative Log Likelihood.
 First of all, let us stating that minimizing the negative log likelihood is equivalent to the Maximum Likelihood Estimation (MLE).
@@
 As suggested by the name, this optimization problem involves maximizing the likelihood distribution $p(\bm{x}|\bm{\theta})$
-\begin{equation}
-    \bm{\theta}_{ML}= \operatorname*{argmax}_{\bm{\theta}}p(\bm{x}|\bm{\theta}).
-\end{equation}
-where
-\begin{equation}
-    p(\bm{x}|\bm{\theta})= \prod_{i-1}^n p(x_i|\bm{\theta}).
-\end{equation}
-Applying the log function in this context is done for several reasons, e.g.
+% \begin{equation}
+%     \bm{\theta}_{ML}= \operatorname*{argmax}_{\bm{\theta}}p(\bm{x}|\bm{\theta}).
+% \end{equation}
+% where
+% \begin{equation}
+%     p(\bm{x}|\bm{\theta})= \prod_{i-1}^n p(x_i|\bm{\theta}).
+% \end{equation}
+% Applying the log function in this context is done for several reasons, e.g.
 reducing the potential of underflow when the likelihoods are small or allowing the conversion of a product into a summation.
-Then, remembering that $-\log$ is strictly decreasing, the following are equivalent:
-\begin{equation}
-    \operatorname*{argmax}_{\bm{\theta}}\sum_{i-1}^n \log{p(x_i|\bm{\theta})}= \operatorname*{argmin}_{\bm{\theta}}-\sum_{i-1}^n \log{p(x_i|\bm{\theta})}.
-\end{equation}
+% Then, remembering that $-\log$ is strictly decreasing, the following are equivalent:
+% \begin{equation}
+%     \operatorname*{argmax}_{\bm{\theta}}\sum_{i-1}^n \log{p(x_i|\bm{\theta})}= \operatorname*{argmin}_{\bm{\theta}}-\sum_{i-1}^n \log{p(x_i|\bm{\theta})}.
+% \end{equation}
 
+%--------------------------------------------------------------------
 \paragraph{\textcolor{violet}{Information Maximising Neural Networks (IMNNs)}}
@@
 \end{equation}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+\subsubsection{Inference strategy}
 
-\subsubsection{Inference strategy}
-%--------------------------------------------------------------------
 To evaluate compression strategies, we propose a comparison of posteriors $p(\theta|s(x))$, where $s(x)$ is the summary statistic.
 Indeed, the more informative the summary statistics, the closer the posterior $p(\theta | s(x))$ is to the true posterior $p (\theta | x)$.
-To this end, we propose to do Neural Posterior Estimation (NPE) that aims to directly learn from simulations the posterior distribution
-
+To this end, we propose to do Neural Posterior Estimation (NPE) \justine{describe a bit more what it is, sbi method etc} that aims to directly learn from simulations the posterior distribution
 \begin{equation}
@@
 \end{equation}
-
-To achieve this, we train a  conditional Normalizing Flow (NF) $p_{\varphi} (\theta | s(x)) $ by minimizing the negative log-probability in $(\theta, s(x))$ space:
-
+To achieve this, we train a  conditional Normalizing Flow (NF) \justine{as proposed by ...
+(REF) / neural simulation-based inference} $p_{\varphi} (\theta | s(x)) $ by minimizing the negative log-probability in $(\theta, s(x))$ space:
 \begin{equation}
-    \mathcal{L}_{\rm NLL} = \mathbb{E}_{p(\theta, s(x))} \left[ - \log p_{\varphi} (\theta | s(x))  \right],
+    \mathcal{L}_{\rm NLL} = \mathbb{E}_{p(\theta, s(x))} \left[ - \log p_{\varphi} (\theta | s(x))  \right].
     \label{eq:nll}
@@
 
+\justine{talk about evaluation of approx posterior at a given obs}
+
 By learning the posterior distribution for each statistic, we capture the significance of each statistic. This approach allows us to use the posterior distribution as a fair benchmark for comparison.
 In particular, for each statistic we use the same NF architecture which is a RealNVP (REF) with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ (REF).
-
-% talk about evaluation of approx posterior at a given obs
 

### 2023-06-25 paper/paper.tex
cropped plot

@@
 %------------------------------------------------------------------
-\section{Experiment}
+\section{Experiment}\label{Sec:experiment}
 \subsection{Explicit Inference}
@@
 \section{Results}\label{Sec:results}
+\begin{figure*}
+    \centering
+\includegraphics[width=\textwidth]{figures/contours_posterior_imp_ex_ps.pdf}
+    \caption{
+Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup.
+The constraints are obtained by applying the $C_{\ell}$ (light blue contours), the full field explicit inference (yellow contours), and the full field implicit inference strategy (violet contours), described in \autoref{Sec:experiment}.
+The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
+\label{fig:contours_posterior_imp_ex_ps}
+\end{figure*}
 %--------------------------------------------------------------------

### 2023-06-28 paper/paper.tex
28/06

@@
 \subsection{Lognormal Modeling}
-For several cosmological purposes, the non-Gaussian field can be modeled as a Lognormal field \citep{coles1991lognormal,bohm2017bayesian}.
-This model has the advantage of offering the rapid generation of the matter or the convergence field while allowing the extraction of information beyond the Gaussian one. 
-Although several studies demonstrated that this model fails in describing the 3D field, it properly describes the 2D convergence field.
-
-Assume a multivariate Gaussian distribution G described by the vector elements of mean $\mu_i$ and covariance matrix $\xi_g^{ij}$, we can define the density contrast $\delta(\bm{x})$ as a shifted lognormal random field :
-% The distribution of the convergence $\kappa_i$ in the $i-$th redshift bin can be modeled as:
-\begin{equation}
-    \delta(\bm{x})=\lambda[e^{g(\bm{x})}-1]. 
-\end{equation}\label{Eq:log_norm_kappa}
+For various cosmological applications, the non-Gaussian field can be modeled as a Lognormal field \citep{coles1991lognormal,bohm2017bayesian}.
+This model offers the advantage of generating the matter or convergence field rapidly while allowing the extraction of information beyond the two-point statistics. 
+Although studies demonstrated that this model fails in describing the 3D field \citep{klypin2018density}, it properly describes the 2D convergence field \citep{clerkin2017testing, xavier2016improving}.
+Assuming a simulated Gaussian convergence map $\kappa_g$, whose statistical properties are fully described by its power spectrum $C_{\ell}$ we know that this model is not a suitable representation of late-time and more evolved structures.
+One potential solution is to find a transformation $f(\kappa_g)$ of this map that captures the non-Gaussian features in the convergence field.
+In doing so, it is crucial to ensure that the transformed map maintains the correct mean and variance, effectively recovering the correct two-point statistics.
+Denoting $\mu$ and $\sigma_g^2$ the mean and covariance matrix of $\kappa_g$ respectively, we can define the transformed convergence $\kappa_{ln}$ as a shifted lognormal random field:
 \begin{equation}\label{Eq:log_norm_kappa}
-    \kappa_{i}=e^{y_i}-\lambda_i. 
+    \kappa_{ln}=e^{\kappa_{g}}-\lambda, 
 \end{equation}
-where $\lambda_i$ is a free parameter defining the shift of the lognormal distribution.
-The density contrast $\delta(\bm{x})$ is fully defined by the shift parameter $\lambda$, the mean of the associated Gaussian field $\mu_i$, and its variance $\sigma_g^2\equiv \xi_g$.
-That fact that we require $<\delta>=0$ implies that the Gaussian field satisfies the following relation:
-\begin{equation}
-    \mu_g=-\frac{\sigma_g^2}{2}.
-\end{equation}
-The parameter $\lambda_i$, is also known as \textit{minimum convergence
-parameter}, since defines the lowest values for all possible values $\kappa$ being the lognormal defined of for positive values.
-We can see from \autoref{Eq:log_norm_kappa} that th ereandom variables $\kappa_i$ is fully defined by the shift parameter $\lambda_i$, the mean of the associated gaussian variables $\mu_i$, and its variance $\sigma_i^2\equiv \xi_g^{ii}$.
-
- The modelling for the shift parameter can be approached in different ways.  determine
-For example, $\lambda$ can be determined following the approach of \citet{xavier2016improving} , i.e.
-matching moments of the distribution, or following the approach of \citet{hilbert2011cosmic}, i.e.
-fitting it as a free parameter.
-In general, it is a parameter dependent from the redshift, the cosmology, and the scale of the field at which the smoothing is applied.
-In this work, we choose the derived $\lambda$ using Cosmomentum.
-The lognormal $\xi^{ij}_{ln}$ and the associated Gaussian covariances $ \xi^{ij}_g$ are related via :
+where $\lambda$ is a free parameter that determines the shift of the lognormal distribution.
+The convergence $\kappa$ in a given redshift bin is fully determined by the shift parameter $\lambda$, the mean $\mu$ of the associated Gaussian field $\kappa_g$, and its variance $\sigma_{g}^2\equiv \xi_g$.
+The correlation of the lognormal field, denoted as $\xi_{ln}$, is also a function of these variables and is related to $\xi^{ij}_g$ through the following equations:
 \begin{align}
@@
 \end{align}
-%Note for Denise: The lambda_i and the \lambda_j in Eq:log_norm_corr are the same parameters defined in Eq:log_norm_kappa only for Gaussian distribution with mean zero.
-Otherwise, these parameters should be defined as alpha_i and alpha_j where alpha_i= <y>+lambda_i.
-See Xavier et al.
-2015 for a detailed description.
+Here $i$ and $j$ define a pair of redshift bins.
+The parameter $\lambda$, also known as \textit{minimum convergence
+parameter}, defines the lowest values for all possible values of $\kappa$.
+The modeling of the shift parameter can be approached in various ways.
+For example, it can be determined by matching moments of the distribution \citep{xavier2016improving} or by treating it as a free parameter \citep{hilbert2011cosmic}.
+In general, the value of $\lambda$ depends on the redshift, cosmology, and the scale of the field at which smoothing is applied.
 
-One of the advantages of modeling the convergence field through a lognormal model is the fact that it enables the constraining of the convergence field in different redshift bins simultaneously while taking into account the correlation between the bins.
-Indeed, in the Fourier space, the power spectrum of $y$ is defined as:
+While it is straightforward to simulate a single map, if we want to constrain the convergence map in different redshift bins, an additional condition must be met.
+The covariance of the map should recover the correct angular power spectrum:
+\begin{equation}\label{power_spectrum_definition}
+    \left \langle \tilde{\kappa}^{(i)}_{ln} (\ell)\tilde{\kappa}^{(j)}_{ln}(\ell')\right \rangle =C^{ij}_{ln}(\ell)\delta^{K}(\ell-\ell')
+\end{equation}
+where $ C^{ij}_{ln}(\ell)$ is the power spectrum of $\kappa_{ln}$ in Fourier space, defined as:
 \begin{equation}\label{Eq:log_norm_cls}
-    C^{ij}_y(\ell)=2\pi \int_0^{\pi} d\theta \sin{\theta}P_{\ell}(\cos{\theta})\xi^{ij}_{y}(\theta)
+    C^{ij}_{ln}(\ell)=2\pi \int_0^{\pi} d\theta \sin{\theta}P_{\ell}(\cos{\theta})\xi^{ij}_{ln}(\theta)
 \end{equation}
-with $P_{\ell}$ Legendre polynomial with order $\ell$.
+and $P_{\ell}$ is the Legendre polynomial of order $\ell$. 
+Using the lognormal model, we can simultaneously constrain the convergence field in different redshift bins while considering the correlation between the bins, as described by \autoref{Eq:log_norm_corr}.
 
-Having defined the survey to reproduce in terms of galaxy number density, redshifts, and shape noise, we can sample the Lognormal random map $\kappa$.
-using \autoref{Eq:log_norm_kappa} and \autoref{Eq:log_norm_corr}.
-First, we need to compute the theoretical auto $C_{\ell}^{i}$ and cross-correlated $C_{\ell}^{ij}$ angular power spectrum for each tomographic bin.
-These theoretical predictions are computed using the public library  \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}.
-Then the 1-D $C_{\ell}$ are projected into 2-D grids of the same size as the desired final convergence maps and the Gaussian covariances $\xi^{ij}_y(\theta)$ are computed through \autoref{Eq:log_norm_corr}.
-To take into account the cross-correlation between different redshift bins, we first stack together the vector $\xi^{ij}_y(\theta)$ to build a covariance matrix having the auto-correlation on the main diagonal and the cross-correlation off-diagonal.
-Then we perform an eigenvalues decomposition.
-This decomposition enables us to decorrelate the maps
-
+ In the SbiLens framework, the sampling of the convergence maps can be described as follows. 
+First, we define the survey in terms of galaxy number density, redshifts, and shape noise.
+Then, we compute the theoretical auto-angular power spectrum $C^{ii}(\ell)$ and cross-angular power spectrum $C^{ij}(\ell)$ for each tomographic bin.
+These theoretical predictions are calculated using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}.
+Next, we project the one-dimensional $C(\ell)$ onto two-dimensional grids with the desired final convergence map size.
+Afterwards, we compute the Gaussian correlation functions $\xi^{ij}_g(\theta)$ using \autoref{Eq:log_norm_corr}.
+ % To take into account the cross-correlation between different redshift bins, 
+ To sample the convergence field in a specific redshift bin while considering the correlation with other bins, we use \autoref{power_spectrum_definition}. 
+We construct the covariance matrix $\bm{\Sigma}$ of the random field $\bm{\kappa}$, where $\bm{\kappa}$ represents the vector of convergence maps at different redshifts as follows:
  \begin{equation}
@@
     C_{\ell}^{n1} & C_{\ell}^{n2} & \cdots & C_{\ell}^{nn} 
-    \end{pmatrix}
+    \end{pmatrix}.
 \end{equation}
+To sample more efficiently, we perform an eigenvalue decomposition of $\bm{\Sigma}$ to obtain a new matrix $\tilde{\bm{\Sigma}}$:
+\begin{equation}
+    \tilde{\bm{\Sigma} }=\bm{Q}\bm{\Lambda}^{1/2}\bm{Q}^{T}
+\end{equation}
+where $\bm{Q}$ and $\bm{\Lambda}$ are the eigenvectors and eigenvalues of $\bm{\Sigma}$, respectively.
+Next, we sample the Gaussian random maps $\bm{\kappa_g}$ using the equation:
+\begin{equation}
+     \bm{\kappa_g}=\bm{Z}*\tilde{\bm{\Sigma} }
+\end{equation}
+where $\bm{Z}$ represents the latent variables of the simulator.
+Finally, we transform the Gaussian map $\kappa_g$ into a LogNormal field using \autoref{Eq:log_norm_kappa}.
+
 %------------------------------------------------------------------
@@
 \subsection{Data generation}
-Our analysis is based on a standard flat $\Lambda$CDM cosmological model, with the following parameters: the baryonic density fraction $\Omega_b$, the total matter density fraction $\Omega_m$, the Hubble parameter $h_0$, the spectral index $n_s$, the amplitude of the primordial power spectrum $\sigma_8$ and the dark energy parameter $w_0$.
-The priors used in the simulations and in inference process are listed in \autoref{tab:prior}, according to \citet{zhang2022transitioning}.
-The simulation process proceeds as follows.
-Given the cosmological parameters, we compute the theoretical power- and cross- spectra using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023}.
-To calculate the
-lognormal shift parameter, we adopt the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which utilizes perturbation theory to compute the cosmology-dependent shift parameters.
-It should be noted that the calculation of the shift parameters is based on a window function, while our pixels are rectangular.
-Following \citet{boruah2022map}, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$  with $\Delta L$, the resolution of the pixels.
-Moreover, as pointed out by \citet{boruah2022map}, this perturbation theory-based approach may not provide accurate results at small scales.
-However, since the main objective of this paper is to compare different compression strategies, we are not concerned about the potential implications of this approximation.
+Our analysis is based on a standard flat $\Lambda$CDM cosmological model,  which includes the following parameters: the baryonic density fraction $\Omega_b$, the total matter density fraction $\Omega_m$, the Hubble parameter $h_0$, the spectral index $n_s$, the amplitude of the primordial power spectrum $\sigma_8$ and the dark energy parameter $w_0$.
+The priors used in the simulations and in the inference process are listed in \autoref{tab:prior}, following \citet{zhang2022transitioning}.
+To simulate our data, we develop the SbiLens package, which employs a Lognormal model to represent the convergence maps, as explained in the previous section.
+Specifically, the package uses the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023} to compute the theoretical power- and cross-spectra.
+The computation of the lognormal shift parameter is performed using the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which utilizes perturbation theory to compute the cosmology-dependent shift parameters.
+In Cosmomentum the calculation of the shift parameters assumes a cylindrical window function, while our pixels are rectangular.
+Following \citet{boruah2022map}, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$, where $\Delta L$ represents the pixel resolution.
+The shift parameters at each redshift are computed using the fiducial cosmology values for $\Omega_b$, $h_0$, and $n_s$.
+To incorporate the cosmology dependence of $\lambda$ from $\Omega_c$, $\sigma_8$ and $w_0$, we calculate the shift for different points in the cosmological parameter space and then interpolate the shift values for other points in the parameter space.
+
+As mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales.
+However, the main objective of this paper is to compare different inference strategies and we are not concerned with the potential implications of this approximation.
 In future applications of the presented inference pipeline, the simulation procedure will be based on N-body simulations.
 
-Simulating the log-normal field for one single convergence map is a straightforward process. 
-However, for a tomographic analysis, as described in the previous section, it is necessary to consider the cross-correlation between different tomographic bins.
-To account for this correlation, we employ the sampling strategy described in Section X.
-\\
-Each map is reproduced on a regular grid with dimensions of $256^2$ pixels and covers an area of $10\times 10$ deg$^2$.
+Each map is reproduced on a regular grid with dimensions of $256 \times 256$ pixels and covers an area of $10\times 10$ deg$^2$.
 
@@
 We conduct a tomographic study to reproduce the redshift distribution and the expected noise for the LSST-Y10 data release.
-Specifically, following \citet{zhang2022transitioning}, we model the underlying redshift distribution using the parametrized Smail distribution \citep{smail1995deep}:
+Following \citet{zhang2022transitioning}, we model the underlying redshift distribution using the parametrized Smail distribution \citep{smail1995deep}:
 \begin{equation}
@@
 \end{equation}
-with $z_0=0.11$ and $\alpha=0.68$ and assuming a photometric redshift error $\sigma_z=0.05(1+z)$ as defined in the LSST DESC Science Requirements Document (SRD, \citet{mandelbaum2018lsst}).
+with $z_0=0.11$ and $\alpha=0.68$. We also assume a photometric redshift error $\sigma_z=0.05(1+z)$ as defined in the LSST DESC Science Requirements Document (SRD, \citet{mandelbaum2018lsst}).
 The galaxy sources are divided into 5 tomographic bins, each containing an equal number of galaxies. 
@@
 The pixel area is given by $A_{pix}\approx$.
-Figure  xx illustrates the resulting source redshift distribution. \autoref{tab:survey_spec} provides a summary of the survey specifications.
+\autoref{fig:redshift_distribution} illustrates the resulting source redshift distribution, and \autoref{tab:survey_spec} provides a summary of the survey specifications.
 \begin{table}
@@
 \paragraph{\textcolor{violet}{Mean Square Error (MSE)}}
-One of the most straightforward approaches adopted to train a Neural Network consists of minimizing the $L_2$ norm, or Mean Square Error (MSE).
-In this section, we demonstrate that this approach is equivalent to training the model to estimate the mean of the posterior distribution, namely:
-\begin{equation}\label{Eq:mean_mse}
-\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}=\operatorname*{argmin}_{\mathcal{F}(\bm{d})}\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\left\Vert \bm {\theta}-\mathcal{F}(\bm{d})
- \right \Vert_{2}^{2}].
+One of the most common approaches used to train a Neural Network is minimizing the $L_2$ norm or Mean Square Error (MSE).
+This approach has been widely adopted in various previous studies  \citep{ribli2018improved, lu2022simultaneously, lu2023cosmological}, where the loss function is typically formulated as follows:
+\begin{equation}
+   \mathcal{L}=\frac{1}{N_{\theta}}(\bm{t}-\bm{\theta})^2.
 \end{equation}
-To demonstrate \autoref{Eq:mean_mse}, we need to minimize the expected value of the $L_2$ norm with respect to $\mathcal{F}(\bm{d})$. Let us consider its derivative:
-\begin{align}\label{Eq:moment_1}
-   & \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm{d}))}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2] =  \\
-    &
-    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2+\mathcal{F}(\bm{d})^2-2\bm {\theta}\mathcal{F}(\bm{d})] = \nonumber \\
-    &
-    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  [
-    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2]+\mathcal{F}(\bm{d})^2-2\mathcal{F}(\bm{d})\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]]= \nonumber \\
-    &2\mathcal{F}(\bm{d})-2 \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. \nonumber
-\end{align}
-Setting it equal to zero, we obtain the critical value:
-\begin{equation}
-    \mathcal{F}(\bm{d})= \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. 
-\end{equation}
-Considering the second-order derivative:
-\begin{equation}\label{Eq:minimum_mean}
-   \frac{\partial^2}{\partial^2 \mathcal{F}(\bm{d})}  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2]=2, 
-\end{equation}
-we can assert that the critical value $\mathcal{F(\bm(\theta))}$ is also a minimum. 
-Since 
-\begin{equation}
-    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}|\bm{d}]= \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}
-\end{equation}
-it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
+Here $N_{\theta}$ represents the number of cosmological parameters, $\bm{t}$ denotes the summary statistics, and $\bm{\theta}$ corresponds to the data vector of the cosmological parameters. 
+However, it is important to note that this approach does not guarantee to recover the maximal information summary statistics.
+Although this can be true for fixed cosmology and Gaussian posterior, this has to be proved for more generic cases.
+Indeed, minimizing the $L_{1}$ norm is equivalent to training the model to estimate the mean of the posterior distribution.
+We prove this statement in XXX of XX.
 %--------------------------------------------------------------------
@@
 The Variational Mutual Information Maximization (VMIM) has been used for cosmological inference problems for the first time by \citet{jeffrey2021likelihood}.
-This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm t$ by training a neural network with parameters $\phi$.
+This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm t$ by training a neural network with parameters $\varphi$.
 Mathematically, it is defined as follows:
@@
 In other words, the mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
-Given the original high-dimensional data vector $\bm d$, the information $\bm t$ is extracted using $\bm {t}=F_{\bm {\phi}}(\bm {d})$. 
-The goal is to find the parameters $\bm {\phi}$ that maximize the mutual information between the summary and cosmological parameters:
+Given the original high-dimensional data vector $\bm d$, the information $\bm t$ is extracted using $\bm {t}=F_{\bm {\varphi}}(\bm {d})$. 
+The goal is to find the parameters $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
 \begin{equation}
-   \bm {\phi}^*= \operatorname*{argmax}_{\bm {\phi}} I(F_{\bm {\phi}}(\bm{d}), \bm {\theta}).
+   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{d}), \bm {\theta}).
 \end{equation}
@@
 \begin{equation}\label{Eq:variational_lower_bound}
-    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\phi}')}]- H(\bm {\theta}).
+    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}]- H(\bm {\theta}).
 \end{equation}
-Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t} ; \bm{\phi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
+Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
 Since the entropy of the cosmological parameters is constant for a fixed training set, the optimization problem based on the lower bound in  \autoref{Eq:variational_lower_bound} can be written as:
 \begin{equation}
-    \operatorname*{argmax}_{\bm {\phi}, \bm {\phi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\phi}}(\bm {d}) ; \bm{\phi}')}].
+    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {d}) ; \bm{\varphi}')}].
 \end{equation}
@@
 
-To evaluate compression strategies, we propose a comparison of posteriors $p(\theta|s(x))$, where $s(x)$ is the summary statistic.
-Indeed, the more informative the summary statistics, the closer the posterior $p(\theta | s(x))$ is to the true posterior $p (\theta | x)$.
-To this end, we propose to do Neural Posterior Estimation (NPE) \justine{describe a bit more what it is, sbi method etc} that aims to directly learn from simulations the posterior distribution
+By comparing the posteriors $p(\theta| t)$ obtained with different compression strategies, we can assess the sensitivity of the results to the choice of statistics.
+In particular, the more informative the summary statistics, the “closer” the posterior $p(\theta | t)$ is to the true posterior $p(\theta | x)$.
+In this section, we describe how we approximate the posterior distribution for each summary statistic t, enable us to use it as a fair benchmark for comparison (we describe the performance metric in section ??).\\
+Since the likelihood $p(t|\theta)$ is intractable, we propose to do Neural Posterior Estimation \citep{npe1, npe2, npe3}, a Simulation-Based Inference (SBI) strategy that aims to directly approximate, from simulations, the posterior distribution
+
 \begin{equation}
-    p(\theta | x) \propto \int p(x | \theta, z) p(z|\theta) p(\theta) dz\,.
+    p(\theta | t) \propto \int p(t | \theta, z) p(z|\theta) p(\theta) dz\,.
 \end{equation}
-To achieve this, we train a  conditional Normalizing Flow (NF) \justine{as proposed by ...
-(REF) / neural simulation-based inference} $p_{\varphi} (\theta | s(x)) $ by minimizing the negative log-probability in $(\theta, s(x))$ space:
+
+Usually, NPE algorithm approximates the posterior distribution through the use of conditional Normalizing Flows $p_{\varphi} (\theta | t)$ \citep{nf1, nf2}.
+This is achieved by minimizing the negative log probability:
+
 \begin{equation}
-    \mathcal{L}_{\rm NLL} = \mathbb{E}_{p(\theta, s(x))} \left[ - \log p_{\varphi} (\theta | s(x))  \right].
+    \mathbb{E}_{p(\theta, t)} \left[ - \log p_{\varphi} (\theta | t)  \right].
     \label{eq:nll}
@@
 
-\justine{talk about evaluation of approx posterior at a given obs}
+Finally, the target posterior $p(\theta | t = t_0)$ is approximated by $p_{\varphi}(\theta | t = t_0)$ where $t_0 = t(x_0)$ with $x_0$ the observed convergence map.
 
-By learning the posterior distribution for each statistic, we capture the significance of each statistic. This approach allows us to use the posterior distribution as a fair benchmark for comparison.
-In particular, for each statistic we use the same NF architecture which is a RealNVP (REF) with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ (REF).
-
+For each summary statistic, we approximate it posterior distribution using the same NF architecture, namely a RealNVP \citep{realnvp} with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ \citep{silu}.
+And, for each summary statistic, the prior $p(\theta)$ used in the inference process remains the same.
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+\subsection{Result estimators}
+To evaluate the results obtained from different approaches, we use two commonly employed estimators found in the literature. Their definitions are provided in the following sections.
+%--------------------------------------------------------------------
+\subsubsection{Figure of Merit (FoM)}
+To quantify the constraining power of each inference strategy, we first employ the Figure of Merit (FoM), defined as follows:
+\begin{equation}\label{Eq:Figure_of_merit}
+    \text{FoM}_{\alpha \beta}=\sqrt{\text{det}(\Tilde{F}_{\alpha \beta})}.
+\end{equation}
+Here, $\alpha$ and $\beta$ represent a pair of cosmological parameters, and $\Tilde{F}_{\alpha \beta}$ refers to the marginalized Fisher matrix.
+We calculate $\Tilde{F}_{\alpha \beta}$ as the inverse of the parameter space covariance matrix $C_{\alpha \beta}$, which is estimated from the HMC chains and from the Normalizing flow sampling.
+Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalised parameter space of the $\alpha$ and $\beta$ pair.
+\subsubsection{KL}
+%--------------------------------------------------------------------
 \begin{center}
@@
             \makecell{MSE \\ VMIM}
-            & \makecell{PyDelfi} & \makecell{$\phi: F_{\phi}(\bm{d})=\bm{\theta}$ \\ with $\bm{\theta}=\Omega_m, \sigma_8$}    
+            & \makecell{PyDelfi} & \makecell{$\varphi: F_{\varphi}(\bm{d})=\bm{\theta}$ \\ with $\bm{\theta}=\Omega_m, \sigma_8$}    
 \\            
@@
 \section{Results}\label{Sec:results}
+We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10.
+The forecasted results are obtained using the simulation procedure outlined in and the parameter inference strategy described in .
+We begin by comparing the outcomes of three different inference procedures: Explicit 2-point statistics, Explicit full-field statistics, and Implicit statistics using CNN summaries.
+Subsequently, we analyze the impact of the different compression strategies on the final cosmological constraints.
+
+\subsection{Power spectrum and full-field statistics}
+We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard-two point statistics and two map-based approaches, the explicit inference using an HMC and the implicit inference using the Normalizing Flow as an estimator.
+As outlined before, our interest is to prove that the two map-based approaches lead to very comparable posterior distributions.
+We present the constraints in the full $\Lambda$CDM parameter space in \autoref{fig:contours_posterior_imp_ex_ps}.
+The contours obtained by the angular $C_{\ell}$ analysis are plotted in blue, the ones for the HMC in yellow, and those for the Normalizing Flow in violet.
+Confirming previous work, we find the two-point statistics are suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$.
+$h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches.
+ We quantify the results by computing the Figure of Merit as described in \autoref{Eq:Figure_of_merit}. The results are presented in \autoref{tab:f_o_m}. 
+A further investigation is represented by 
+-show the results for the KL divergence 
+-present a comparison with other works.
+
 \begin{figure*}
@@
 \end{figure*}
+
+\begin{center}
+\begin{table*}
+\caption{ Values of the Figure of Merit (FoM) as defined in \autoref{Eq:Figure_of_merit} for different inference strategies: the convergence power spectrum $C_{\ell}$, the HMC, the CNN map compressed statistics with the MSE, the VMIM, the MAE and the GNLL loss functions.
+The values of the figure of merit are inversely proportional to the area of the contours, larger the FoM, higher the constraining power.}
+\begin{tabular}{lcccccc} 
+ \hline
+    FoM &  & Explicit Inference  &  & & Implicit inference  \\
+     & $C_{\ell}$ & Full Field (HMC)&  VMIM & MSE & MAE & GNLL   \\
+  \hline
+ $\Omega_c-  \Omega_b$ & 5315 & 6951 & 6591 & 5651 & 5909\\
+ $\Omega_c-  \sigma_8$ & 1222 & 2520 & 2526 & 2043 & 2316\\
+ $\Omega_c-  h_0$      & 516  & 698  & 675  & 601  & 618\\
+ $\Omega_c-  n_s$      & 898  & 1243 & 1184 & 964  & 1030\\
+ $\Omega_c-  w_0$      & 100  & 198  & 190  & 152  & 162\\
+ $\Omega_b-  \sigma_8$ & 3997 & 5472 & 5514 & 5066 & 5398\\
+ $\Omega_b-  h_0$      & 2800 & 2994 & 2805 & 2923 & 2979\\
+ $\Omega_b-  n_s$      & 4035 & 4516 & 4174 & 4073 & 4259\\
+ $\Omega_b-  w_0$      & 523  & 846  & 791  & 736  & 771\\
+ $\sigma_8-  h_0$      & 402  & 566  & 577  & 544  & 571\\
+ $\sigma_8-  n_s$      & 583  & 889  & 891  & 787  & 854\\ 
+ $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
+ $h_0-       n_s$      & 425  & 502  & 458  & 455  & 469\\
+ $h_0-       w_0$      & 52   & 88   & 83   & 80   & 82\\
+ $n_s-       w_0$      & 80   & 133  & 126  & 112  & 119\\
+     \hline
+\end{tabular}
+\label{tab:f_o_m}
+\end{table*}
+\end{center}
 %--------------------------------------------------------------------
@@
 \bibliography{paper/biblio}
+
+\begin{appendix}
+
+\end{appendix}
+\section{Mean Square Error (MSE)}
+In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the mean of the posterior distribution, namely:
+\begin{equation}\label{Eq:mean_mse}
+\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}=\operatorname*{argmin}_{\mathcal{F}(\bm{d})}\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\left\Vert \bm {\theta}-\mathcal{F}(\bm{d})
+ \right \Vert_{2}^{2}].
+\end{equation}
+To demonstrate \autoref{Eq:mean_mse}, we need to minimize the expected value of the $L_2$ norm with respect to $\mathcal{F}(\bm{d})$. Let us consider its derivative:
+\begin{align}\label{Eq:moment_1}
+   & \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm{d}))}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2] =  \\
+    &
+    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2+\mathcal{F}(\bm{d})^2-2\bm {\theta}\mathcal{F}(\bm{d})] = \nonumber \\
+    &
+    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  [
+    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2]+\mathcal{F}(\bm{d})^2-2\mathcal{F}(\bm{d})\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]]= \nonumber \\
+    &2\mathcal{F}(\bm{d})-2 \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. \nonumber
+\end{align}
+Setting it equal to zero, we obtain the critical value:
+\begin{equation}
+    \mathcal{F}(\bm{d})= \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. 
+\end{equation}
+Considering the second-order derivative:
+\begin{equation}\label{Eq:minimum_mean}
+   \frac{\partial^2}{\partial^2 \mathcal{F}(\bm{d})}  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2]=2, 
+\end{equation}
+we can assert that the critical value $\mathcal{F(\bm(\theta))}$ is also a minimum. 
+Since 
+\begin{equation}
+    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}|\bm{d}]= \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}
+\end{equation}
+it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
 \end{document}

### 2023-06-28 paper/paper.tex
28/06_pt2

@@
 \and
+T. Lucas Makinen \inst{4}
+\and
 Jean-Luc Starck \inst{3}
@@
 Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France
+\and 
 }
@@
 
-   \keywords{giant planet formation --
-                $\kappa$-mechanism --
-                stability of gas spheres
+   \keywords{methods: statistical – gravitational lensing: weak – cosmology: large-scale structure of Universe
                }
@@
 %------------------------------------------------------------------
-\section{Inference over forward simulation models}
-With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, enable to fully account for the non-Gaussian information in the lensing field at
-the scales that future surveys will be able to access.
-One of the goals of this paper is to present a forward model of the full observables to directly extract information from the raw pixel data, instead of attempting to evaluate analytically the likelihood of any summary statistics and in doing so, preserve all the available information.
-This approach has the advantage of easily incorporating systematic effects or easily combining multiple cosmological probes by joint simulations.  
-In this context, the simulator of the observables becomes our physical model, in which every single component is tractable.
-These models, also referred to as \textit{probabilistic program}, take as input the parameters $\bm{\theta}$, sample some \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, and generates the output $\bm x \sim p(\bm d|\bm \theta)$, with $\bm{d}$ observations.
-At this point, we can infer both the latent variables or the parameter $\bm{\theta}$.
-However,  the marginal likelihood is typically intractable:
+\section{Motivation}
+With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, that may not fully capture the non-Gaussian information present in the lensing field at the scales accessible to future surveys.
+One of the main objectives of this paper is to introduce a forward model that directly extracts information from the raw pixel data, rather than relying on the analytical evaluation of summary statistics.
+By doing so, we aim to preserve all available information and facilitate the incorporation of systematic effects and the combination of multiple cosmological probes through joint simulations.
+In this context, the simulator of the observables serves as our physical model, where each component is tractable.
+These models, also knows as \textit{probabilistic program}, considering a prior distribution $p(\bm{\theta})$ for the parameter $\bm{\theta}$, sample \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, generating the output $\bm x \sim p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
+However, computing the marginal likelihood is typically intractable:
 \begin{equation}
@@
 \end{equation}
-To bypass this limitation, while keeping exploring the full information content of data, we have different approaches.
-Although in literature these approaches are usually referred to in different ways, in this work hereafter we will make the following distinction
+To overcome this limitation while still capturing the full information content of the data, different approaches have been proposed in the literature.
+Although these approaches are often referred to by different names, in this work, we will make the following distinction
 %--------------------------------------------------------------------
@@
 %------------------------------------------------------------------
-\section{The SBILens framework}
+\section{The SBILens framework}\label{Sec:the SBILens framework}
 \subsection{Lognormal Modeling}
@@
 %--------------------------------------------------------------------
-\subsubsection{Need for data compression}
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
 \subsubsection{Benchmark compression scheme}
-There are several approaches available for training the compressor, and this section aims to provide an overview of these different methods used in previous works. 
-In \autoref{Sec:results}, we will compare the outcomes of these various strategies.
-Specifically, using the same prior $p(\bm {\theta})$ for the cosmological parameter $\bm{\theta}$ and the simulator described in \autoref{Sec:results}, which samples $\bm {x}$ from $p(\bm{x}| \bm{\theta})$ and produces a set of mock observations $\bm{x_0}$, we will compare the posterior distribution $p(\bm{\theta}|\bm{x})$ obtained form each inference pipeline.
-This analysis will help to determine the optimal compression strategy, enabling us to extract the maximum amount of information possible given the fixed cosmological parameters.
+To ensure the scalability of density estimation LFI in cases where forward simulations are computationally expensive, it becomes necessary to employ compression techniques that reduce the dimensionality of the data space and extract summary statistics.
+Specifically, we try to find a function $\bm{t}=F(\bm{d})$, where $\bm{t}$ represents low-dimensional summaries of the original data vector $\bm{d}$.
+The objective is to achieve a compression function $F(\bm{d})$ that maximizes the information while minimizing dimensionality.
+Previous studies \citep{alsing2018generalized} have demonstrated that, under specific conditions, a compression scheme can be achieved where the dimension of the summaries dim($\bm{t}$) is equal to the dimension of the unknown parameters dim($\bm{\theta}$) without any loss of information at the Fisher level.
+Multiple approaches exist in an attempt to satisfy this condition.
+This section aims to provide an overview of the various neural compression-based methods employed in previous works.
+
 %--------------------------------------------------------------------
 \paragraph{\textcolor{violet}{Mean Square Error (MSE)}}
-One of the most common approaches used to train a Neural Network is minimizing the $L_2$ norm or Mean Square Error (MSE).
-This approach has been widely adopted in various previous studies  \citep{ribli2018improved, lu2022simultaneously, lu2023cosmological}, where the loss function is typically formulated as follows:
+%--------------------------------------------------------------------
+One of the commonly used techniques for training a Neural Network is by minimizing the $L_2$ norm or Mean Square Error (MSE).
+This methods has been widely adopted in various previous studies  \citep{ribli2018improved, lu2022simultaneously, lu2023cosmological}, where the loss function is typically formulated as follows:
 \begin{equation}
@@
 Here $N_{\theta}$ represents the number of cosmological parameters, $\bm{t}$ denotes the summary statistics, and $\bm{\theta}$ corresponds to the data vector of the cosmological parameters. 
-However, it is important to note that this approach does not guarantee to recover the maximal information summary statistics.
-Although this can be true for fixed cosmology and Gaussian posterior, this has to be proved for more generic cases.
-Indeed, minimizing the $L_{1}$ norm is equivalent to training the model to estimate the mean of the posterior distribution.
-We prove this statement in XXX of XX.
+However, it is important to note that this approach does not guarantee the recovery of maximally informative summary statistics.
+While this may hold true for fixed fiducial values of the cosmological parameters and assumptions of Gaussianity, its validation to more generic cases need to be proved.
+Indeed, minimizing the $L_{2}$ norm is equivalent to training the model to estimate the mean of the posterior distribution.
+We prove this statement in \autoref{Sec:appendix_Mean Square Error}.
 %--------------------------------------------------------------------
 \paragraph{\textcolor{violet}{Maximum Absolute Error (MAE)}}
-Instead of minimizing the $L_2$ norm, these approaches focus on the $L_1$ norm or Mean Absolute Error (MAE).
-In this section, we demonstrate that this is equivalent to training the model to estimate the median of the posterior distribution.
-By definition, the median of a probability density function $p(x)$ is a real number $m$ that satisfies:
-\begin{equation}\label{Eq:definition_median}
-\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
-\end{equation}. 
-The expectation value of the mean absolute error is defined as:
+%--------------------------------------------------------------------
+Another commonly used approach involves minimizing the $L_1$ norm or Mean Absolute Error (MAE). In this approach, the loss function is defined as:
 \begin{equation}
-    \mathbb{E}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
+    \mathcal{L}=|\bm{t}-\bm{\theta}|
 \end{equation}
-which can be decomposed as
+where $\bm{t}$ represents the summary statistics and $\bm{\theta}$ denotes the cosmological parameters.
+In \autoref{Sec:appendix_Maximum Absolute Error}, we demonstrate that minimizing this loss function is equivalent to training the model to estimate the median of the posterior distribution.
+While extensively employed in various previous studies \citep{2018PhRvD..97j3515G, fluri2018cosmological, ribli2019weak}, similar to the previous consideration for the MSE, it is important to note that the effectiveness of this loss function in extracting sufficient statistics needs to be demonstrated in more generic applications.
+%--------------------------------------------------------------------
+\paragraph{\textcolor{violet}{Variational Mutual Information Maximization (VMIM)}}
+%--------------------------------------------------------------------
+The Variational Mutual Information Maximization (VMIM) technique was first introduced for cosmological inference problems by \citet{jeffrey2021likelihood}.
+This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm{t}$.
+In the VMIM approach, the loss function is defined as:
+\begin{equation}\label{Eq:Loss_vmim}
+    \mathcal{L}=- \log{q(\bm{\theta}|\bm{t};\bm{\varphi'})}.
+\end{equation}
+Here, $q(\bm{\theta}|\bm{t};\bm{\varphi'})$ represents a variational conditional distribution, $\bm{t}$ denotes the summary statistics, $\bm{\theta}$ corresponds to the data vector of the cosmological parameters, and $\bm{\varphi'}$ are the parameters of the Neural Network.
+In order to understand the significance of this loss function, it is necessary to start by considering the mathematical definition of mutual information $I(\bm{t}, \bm {\theta})$:
+\begin{align}\label{Eq:mutual_information}
+    I(\bm{t}, \bm {\theta}) &= D_{KL}(p(\bm {t}, \bm {\theta})||p(\bm {t})p(\bm {\theta})) \\ \nonumber
+    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
+    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
+        &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta}  d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
+    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
+    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- \mathbb{E}_{p(\bm {\theta})} [\log{p(\bm {\theta})}] \\ \nonumber
+    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- H(\bm {\theta});
+\end{align}
+
+in the above equation, $D_{KL}$ is the Kullback-Leibler divergence \citep{kullback1951information}, $p(\bm {t}, \bm {\theta})$ is the joint probability distribution of summary statistics and cosmological parameters, and $H(\bm {\theta})$ represents the \textit{entropy} of the distribution of cosmological parameters.
+Essentially, mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
+The summary statistics $\bm{t}$ are extracted from the original high-dimensional data vector $\bm{d}$ using $\bm{t}=F_{\bm {\varphi}}(\bm{d})$.
+The goal is to find the parameters $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
 \begin{equation}
-        \int_{\infty}^{m}p(x)|x-m|dx +\int_{m}^{\infty}p(x)|x-m|dx .
+   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{d}), \bm {\theta}).
 \end{equation}
-To minimize this function with respect to $m$, we need to compute its derivative:
-\begin{equation}\label{Eq:absolute_median}
-    \frac{d\mathbb{E}[|x-m|]}{dm}=
-    \frac{d}{dm}\int_{\infty}^{m}p(x)|x-m|dx +\frac{d}{dm}\int_{m}^{\infty}p(x)|x-m|dx. 
+However, the mutual information expressed in \autoref{Eq:mutual_information} is not tractable.
+To overcome this limitation, various approaches have been developed that rely on tractable bounds, enabling the training of deep neural networks to optimize the mutual information.
+In this study, we adopt the same strategy used by \citet{jeffrey2021likelihood}, which involves using the variational lower bound \citep{barber2003information}:
+\begin{equation}\label{Eq:variational_lower_bound}
+    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}]- H(\bm {\theta}).
 \end{equation}
-Considering that $|x-m|=(x-m)$ for $m\le x$ and $|x-m|=(m-x)$ $m\ge x$, 
-we can write \autoref{Eq:absolute_median} as:
+Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t}; \bm{\varphi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
+As the entropy of the cosmological parameters remains constant for a fixed training set, the optimization problem based on the lower bound in \autoref{Eq:variational_lower_bound} can be formulated as:
 \begin{equation}
-    \frac{d\mathbb{E}[|x-m|]}{dm}=
-    \frac{d}{dm}\int_{\infty}^{m}p(x)(m-x)dx +\frac{d}{dm}\int_{m}^{\infty}p(x)(x-m)dx .
+    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {d}) ; \bm{\varphi}')}], 
 \end{equation}
-Using the Leibniz integral rule, we get:
-\begin{align}
-    &\frac{d\mathbb{E}[|x-m|]}{dm}= \\
-    &
-    p(x)(m-m)\frac{dm}{dm}+\int_{\infty}^{m}\frac{\partial}{\partial m}[p(x)(m-x)]dx + \nonumber \\
-    & - p(x)(m-m)\frac{dm}{dm}+\int_{m}^{\infty}\frac{\partial}{\partial m}[p(x)(m-x)]dx \nonumber .
-\end{align}
-Setting the derivative to zero, we obtain:
-\begin{equation}
-    \frac{d\mathbb{E}[|x-m|]}{dm}= \int_{\infty}^{m} p(x)dx-\int_{m}^{\infty}p(x)dx =0.
-\end{equation}
-Thus,
-\begin{equation}
-\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx .
-\end{equation}
-Considering that
-\begin{equation}
-\int_{\infty}^{m} p(x)dx+\int_{m}^{\infty}p(x)dx=1,
-\end{equation}
-we obtain 
-\begin{equation}\label{Eq:definition_median}
-\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
-\end{equation}
+yielding \autoref{Eq:Loss_vmim}.
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\FL{I think we should have this section after the VMIM}
-
 \FL{Recognizing that the uncertainty on different cosmological parameters will vary, a third
@@
 }
-% A third possible scenario consists in minimizing the Negative Log Likelihood.
-First of all, let us stating that minimizing the negative log likelihood is equivalent to the Maximum Likelihood Estimation (MLE).
-The proof of this statement is straightforward.
-As suggested by the name, this optimization problem involves maximizing the likelihood distribution $p(\bm{x}|\bm{\theta})$
-% \begin{equation}
-%     \bm{\theta}_{ML}= \operatorname*{argmax}_{\bm{\theta}}p(\bm{x}|\bm{\theta}).
-% \end{equation}
-% where
-% \begin{equation}
-%     p(\bm{x}|\bm{\theta})= \prod_{i-1}^n p(x_i|\bm{\theta}).
-% \end{equation}
-% Applying the log function in this context is done for several reasons, e.g.
-reducing the potential of underflow when the likelihoods are small or allowing the conversion of a product into a summation.
-% Then, remembering that $-\log$ is strictly decreasing, the following are equivalent:
-% \begin{equation}
-%     \operatorname*{argmax}_{\bm{\theta}}\sum_{i-1}^n \log{p(x_i|\bm{\theta})}= \operatorname*{argmin}_{\bm{\theta}}-\sum_{i-1}^n \log{p(x_i|\bm{\theta})}.
-% \end{equation}
-
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\paragraph{\textcolor{violet}{Variational Mutual Information Maximization (VMIM)}}
-The Variational Mutual Information Maximization (VMIM) has been used for cosmological inference problems for the first time by \citet{jeffrey2021likelihood}.
-This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm t$ by training a neural network with parameters $\varphi$.
-Mathematically, it is defined as follows:
-\begin{align}\label{Eq:mutual_information}
-    I(\bm{t}, \bm {\theta}) &= D_{KL}(p(\bm {t}, \bm {\theta})||p(\bm {t})p(\bm {\theta})) \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
-        &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta}  d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
-    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- \mathbb{E}_{p(\bm {\theta})} [\log{p(\bm {\theta})}] \\ \nonumber
-    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- H(\bm {\theta});
-\end{align}
 
-where $D_{KL}$ is the Kullback-Leibler divergence \citep{kullback1951information}, $p(\bm {t}, \bm {\theta})$ is the joint probability distribution of summary statistics and cosmological parameters, and $H(\bm {\theta})$ represents the \textit{entropy} of the distribution of cosmological parameters.
-In other words, the mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
-Given the original high-dimensional data vector $\bm d$, the information $\bm t$ is extracted using $\bm {t}=F_{\bm {\varphi}}(\bm {d})$. 
-The goal is to find the parameters $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
-\begin{equation}
-   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{d}), \bm {\theta}).
-\end{equation}
-However, the mutual information expressed in \autoref{Eq:mutual_information} is not tractable.
-To overcome this limitation,  several approaches have been developed, relying on tractable bounds that allow the training of a deep neural network to optimize the mutual information.
-In this work, we adopt the same strategy used by \citet{jeffrey2021likelihood} and use the variational lower bound \citep{barber2003information}:
-\begin{equation}\label{Eq:variational_lower_bound}
-    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}]- H(\bm {\theta}).
-\end{equation}
-Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
-Since the entropy of the cosmological parameters is constant for a fixed training set, the optimization problem based on the lower bound in  \autoref{Eq:variational_lower_bound} can be written as:
-\begin{equation}
-    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {d}) ; \bm{\varphi}')}].
-\end{equation}
 %--------------------------------------------------------------------
@@
 Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalised parameter space of the $\alpha$ and $\beta$ pair.
-\subsubsection{KL}
+\subsubsection{Kullback-Leibler divergence}
 %--------------------------------------------------------------------
@@
 \end{acknowledgements}
-
+%--------------------------------------------------------------------
+%              #########    START BIBLIO   #########
+%--------------------------------------------------------------------
 \bibliographystyle{aa} % style aa.bst
 \bibliography{paper/biblio} 
-
+%--------------------------------------------------------------------
+%              #########    START APPENDIX   #########
+%--------------------------------------------------------------------
 \begin{appendix}
-
-\end{appendix}
-\section{Mean Square Error (MSE)}
+\section{Mean Square Error (MSE)}\label{Sec:appendix_Mean Square Error}
 In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the mean of the posterior distribution, namely:
@@
 it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
+%--------------------------------------------------------------------
+\section{Maximum Absolute Error (MAE)}\label{Sec:appendix_Maximum Absolute Error}
+In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the median of the posterior distribution.
+By definition, the median of a probability density function $p(x)$ is a real number $m$ that satisfies:
+\begin{equation}\label{Eq:definition_median}
+\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
+\end{equation}. 
+The expectation value of the mean absolute error is defined as:
+\begin{equation}
+    \mathbb{E}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
+\end{equation}
+which can be decomposed as
+\begin{equation}
+        \int_{\infty}^{m}p(x)|x-m|dx +\int_{m}^{\infty}p(x)|x-m|dx .
+\end{equation}
+To minimize this function with respect to $m$, we need to compute its derivative:
+\begin{equation}\label{Eq:absolute_median}
+    \frac{d\mathbb{E}[|x-m|]}{dm}=
+    \frac{d}{dm}\int_{\infty}^{m}p(x)|x-m|dx +\frac{d}{dm}\int_{m}^{\infty}p(x)|x-m|dx. 
+\end{equation}
+Considering that $|x-m|=(x-m)$ for $m\le x$ and $|x-m|=(m-x)$ $m\ge x$, 
+we can write \autoref{Eq:absolute_median} as:
+\begin{equation}
+    \frac{d\mathbb{E}[|x-m|]}{dm}=
+    \frac{d}{dm}\int_{\infty}^{m}p(x)(m-x)dx +\frac{d}{dm}\int_{m}^{\infty}p(x)(x-m)dx .
+\end{equation}
+Using the Leibniz integral rule, we get:
+\begin{align}
+    &\frac{d\mathbb{E}[|x-m|]}{dm}= \\
+    &
+    p(x)(m-m)\frac{dm}{dm}+\int_{\infty}^{m}\frac{\partial}{\partial m}[p(x)(m-x)]dx + \nonumber \\
+    & - p(x)(m-m)\frac{dm}{dm}+\int_{m}^{\infty}\frac{\partial}{\partial m}[p(x)(m-x)]dx \nonumber .
+\end{align}
+Setting the derivative to zero, we obtain:
+\begin{equation}
+    \frac{d\mathbb{E}[|x-m|]}{dm}= \int_{\infty}^{m} p(x)dx-\int_{m}^{\infty}p(x)dx =0.
+\end{equation}
+Thus,
+\begin{equation}
+\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx .
+\end{equation}
+Considering that
+\begin{equation}
+\int_{\infty}^{m} p(x)dx+\int_{m}^{\infty}p(x)dx=1,
+\end{equation}
+we obtain 
+\begin{equation}\label{Eq:definition_median}
+\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
+\end{equation}
+%--------------------------------------------------------------------
+%              #########    END APPENDIX   #########
+%--------------------------------------------------------------------
+\end{appendix}
 \end{document}

### 2023-07-06 script/train_compressor_gnll.py
backup gnll

@@ -1,0 +1,217 @@
+
+from functools import partial
+from tqdm import tqdm
+import pickle
+from pathlib import Path
+import tensorboard
+import numpy as np
+import jax
+import jax.numpy as jnp
+
+import flax.linen as nn
+from flax.metrics import tensorboard
+
+from haiku._src.nets.resnet import ResNet18
+import optax
+import haiku as hk
+
+import tensorflow_datasets as tfds
+import tensorflow as tf
+from tensorflow_probability.substrates import jax as tfp
+
+tfd = tfp.distributions
+tfb = tfp.bijectors
+tfd = tfp.distributions
+
+from sbi_lens.gen_dataset.lensing_lognormal_dataset import LensingLogNormalDataset
+from sbi_lens.gen_dataset.utils import augmentation_noise, augmentation_flip
+from sbi_lens.config import config_lsst_y_10
+
+
+'unset XLA_FLAGS'
+
+import logging
+
+
+import tensorflow_probability as tfp; tfp = tfp.substrates.jax
+
+# this prints a WARNING
+
+tfp.distributions.TransformedDistribution(
+    tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity()
+)
+
+logger = logging.getLogger("root")
+
+
+class CheckTypesFilter(logging.Filter):
+    def filter(self, record):
+        return "check_types" not in record.getMessage()
+
+
+logger.addFilter(CheckTypesFilter())
+
+# this does not print a WARNING
+
+tfp.distributions.TransformedDistribution(tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity())
+
+
+
+N = config_lsst_y_10.N
+map_size = config_lsst_y_10.map_size
+sigma_e = config_lsst_y_10.sigma_e
+gals_per_arcmin2 = config_lsst_y_10.gals_per_arcmin2
+nbins = config_lsst_y_10.nbins
+a = config_lsst_y_10.a
+b = config_lsst_y_10.b
+z0 = config_lsst_y_10.z0
+
+truth = config_lsst_y_10.truth
+
+params_name = config_lsst_y_10.params_name
+
+tf.random.set_seed(1)
+
+augmentation = lambda example: augmentation_flip(
+    augmentation_noise(example=example,
+                       N=N,
+                       map_size=map_size,
+                       sigma_e=sigma_e,
[…]

### 2023-07-06 paper/paper.tex

@@
 
-\title{Unveiling the Optimal Approach for Neural Network-Based Cosmological Parameter Inference through Likelihood Implicit Methods}
+\title{Likelihood Implicit Methods}
 
@@
 \newcommand{\FL}[1]{{\color{magenta}FL: #1}}
+\newcommand{\LM}[1]{{\color{olive}LM: #1}}
 
@@
 \and
+T. Lucas Makinen \inst{3}
+\and
 Alexandre Boucaud \inst{2}
 \and
-Fran\c{c}ois Lanusse \inst{3}
+Fran\c{c}ois Lanusse \inst{4}
 \and
-T. Lucas Makinen \inst{4}
-\and
-Jean-Luc Starck \inst{3}
+Jean-Luc Starck \inst{4}
 }
@@
 Université Paris Cité, CNRS, Astroparticule et Cosmologie, F-75013 Paris, France
+\and Imperial Centre for Inference and Cosmology (ICIC) $\&$ Astrophysics Group, Imperial College London, Blackett Laboratory, Prince Consort Road, London SW7 2AZ, United Kingdom
 \and
 Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France
-\and 
 }
@@
 \section{Introduction}
+Weak gravitational lensing by the large-scale structure is caused by the presence of foreground matter bending the light emitted by background galaxies.
+Being sensitive to the large-scale structure of the universe, it is one of the most promising tools for investigating the nature of dark energy, the origin of the accelerating expansion of the universe and estimating cosmological parameters.
+Future cosmological surveys, such as the Legacy Survey of Space and Time (LSST) of the Vera C.
+Rubin Observatory \citep{ivezic2019lsst}, the Nancy Grace Roman Space Telescope \citep{spergel2015wide}, and  the Euclid Mission \citep{laureijs2011euclid}, will rely on weak gravitational lensing as one of the principal physical probes to address unresolved questions in current cosmology.
+ As these surveys become deeper, they will be able to access more non-Gaussian features of the matter fields.
+This makes the standard weak lensing analyses, which rely on two-point statistics such as the 2-point shear correlation or the angular power spectrum, sub-optimal.
+These analyses are unable to fully capture the non-Gaussian information imprinted in the lensing signal and can only access Gaussian information.
+
+To overcome this limitation, several higher-order statistics have been introduced.
+These include weak lensing peak counts \citep{liu2015cosmology,  liu2015cosmological, lin2015new, kacprzak2016cosmology, peel2017cosmological, shan2018kids, martinet2018kids, ajani2020constraining, harnois2021cosmic, zurcher2022dark}, wavelet and scattering transform \citep{ajani2021starlet, cheng2021weak}, the one point PDF \citep{liu2019constraining, uhlemann2020fisher, boyle2021nuw}, Minkowski functionals \citep{kratochvil2012probing, petri2013cosmology}, moments of mass maps \citep{gatti2021dark},  and 3 point statistics \citep{takada2004cosmological, semboloni2011weak, rizzato2019tomographic, halder2021integrated}.
+Although these methods have proven to improve cosmological constraints, they rely on summary statistics, for which an analytical description is often lacking.
+These methods frequently assume a Gaussian likelihood and require accurate computation of the covariance matrix, which can be challenging.
+
+One way to overcome the use of summary statistics and bypass the associated issues, such as the use of simplified and unjustified likelihood assumptions, is through the use of map-based approaches.
+Within this category, a further distinction can be made between methods that integrate observations into a forward model allowing the exact reconstruction of the likelihood, and those designed to directly reconstruct the likelihood from synthetic data as part of the inference pipeline.
+The former scenario, often referred to as Bayesian forward-modeling,  has gained popularity in the literature in the last decades \citep{schneider2015hierarchical, alsing2016hierarchical, alsing2017cosmological, bohm2017bayesian, porqueres2021bayesian, porqueres2022lifting, porqueres2023field}.
+Despite the differences among these works in terms of the physical models they assume or the specific quantities they aim to sample, they all demonstrate that conducting a field-level analysis results in more precise and accurate constraints compared to the two-point function analysis.
+Despite the immense potential of these approaches, there is a noteworthy limitation: they frequently result in high-dimensional problems, necessitating the use of sophisticated statistical sampling techniques.
+\\
+An alternative framework for inference is provided by Likelihood-Free Inference (LFI) approaches, which are sometimes referred to as Likelihood implicit or Simulation-Based Inference (SBI).
+\denise{TODO: cite some other work(?)}
+The potential of LFI methods lies in the fact that they primarily rely on our ability to forward-simulate synthetic data, without the need for an explicit likelihood function.
+As a result, these methods are free from assumptions and approximations.
+Furthermore, incorporating systematic effects and complex physical processes into forward simulations is more straightforward compared to including them when constructing the likelihood function.
+Given the potential of LFI methods, it is crucial to identify possible sources of bias and the associated challenges, as well as establish well-defined procedures to validate the results.
+For instance, it is widely recognized that even the most advanced LFI methods can be affected by the curse of dimensionality.
+This necessitates compression procedures that reduce the high-dimensional data set into a small number of summary statistics.
+A significant question is how to design data compression procedures, aiming to derive low-dimensional summary statistics while minimizing information loss.
+
+In this paper, we present a comparison of the performance of various neural network-based compression schemes within the Likelihood-Free Inference (LFI) framework.
+These methods differ in terms of the loss functions used to train the neural network, but they share the same inference strategy based on neural density estimation.
+Moreover, a crucial point is to validate the results obtained in the LFI context.
+To address this, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
+The forward model is based on \texttt{SbiLens}, a Python package designed for Weak Lensing Implicit Inference with Differentiable Simulator implemented in \texttt{Jax}.
+\texttt{SbiLens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
+
+The paper is structured as follows: in \autoref{Sec:Motivation}, we illustrate the motivation behind this work.
+In \autoref{Sec:the SBILens framework}, we introduce the \texttt{SbiLens} framework and describe the simulated data used in this work.
+In \autoref{Sec:experiment}, we detail the inference strategy and the three different approaches we used: the power spectrum, the map-based inference based on the Bayesian forward-modeling, and the map-based inference based on the LFI.
+In the same section, we also provide a detailed overview of different neural compression strategies.
+In \autoref{Sec:results}, we discuss the results and validate the LFI approaches.
+Finally, we conclude in \autoref{Sec:conclusion}.
+
 %------------------------------------------------------------------
@@
 %------------------------------------------------------------------
-\section{Motivation}
+\section{Motivation}\label{Sec:Motivation}
+%--------------------------------------------------------------------
 With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, that may not fully capture the non-Gaussian information present in the lensing field at the scales accessible to future surveys.
@@
 Afterwards, we compute the Gaussian correlation functions $\xi^{ij}_g(\theta)$ using \autoref{Eq:log_norm_corr}.
- % To take into account the cross-correlation between different redshift bins, 
  To sample the convergence field in a specific redshift bin while considering the correlation with other bins, we use \autoref{power_spectrum_definition}. 
@@
 Finally, we transform the Gaussian map $\kappa_g$ into a LogNormal field using \autoref{Eq:log_norm_kappa}.
-
 %------------------------------------------------------------------
@@
 \subsection{Data generation}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 Our analysis is based on a standard flat $\Lambda$CDM cosmological model,  which includes the following parameters: the baryonic density fraction $\Omega_b$, the total matter density fraction $\Omega_m$, the Hubble parameter $h_0$, the spectral index $n_s$, the amplitude of the primordial power spectrum $\sigma_8$ and the dark energy parameter $w_0$.
@@
 \subsection{Noise and survey setting}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 We conduct a tomographic study to reproduce the redshift distribution and the expected noise for the LSST-Y10 data release.
@@
 \section{Experiment}\label{Sec:experiment}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 \subsection{Explicit Inference}
@@
 \subsubsection{Power spectrum}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 Although this is not a full-field approach, we aim to compare the latest to a more standard approach. 
@@
 \subsubsection{Full field and HMC}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 The following steps outline our methods.
@@
 This section aims to provide an overview of the various neural compression-based methods employed in previous works.
-
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\paragraph{\textcolor{violet}{Information Maximising Neural Networks (IMNNs)}}
+\paragraph{\textcolor{violet}
+{Information Maximising Neural Networks (IMNNs)}}.
+\LM{TODO: condensed introduction; emphasize learning compression at single point; network learns score function which generalises smoothly away from fiducial \cite{Makinen_2021, Makinen_2022, Charnock_2018}, Wandelt (in prep.)
+}
+
 %--------------------------------------------------------------------
@@
 \subsubsection{Inference strategy}
-
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 By comparing the posteriors $p(\theta| t)$ obtained with different compression strategies, we can assess the sensitivity of the results to the choice of statistics.
@@
 \subsection{Result estimators}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 To evaluate the results obtained from different approaches, we use two commonly employed estimators found in the literature. Their definitions are provided in the following sections.
@@
 \end{center}
-
 %--------------------------------------------------------------------
 \section{Results}\label{Sec:results}
+%--------------------------------------------------------------------
 We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10.
@@
 Subsequently, we analyze the impact of the different compression strategies on the final cosmological constraints.
-
+%--------------------------------------------------------------------
 \subsection{Power spectrum and full-field statistics}
+%--------------------------------------------------------------------
 We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard-two point statistics and two map-based approaches, the explicit inference using an HMC and the implicit inference using the Normalizing Flow as an estimator.
@@
 -present a comparison with other works. 
-
+%--------------------------------------------------------------------
+% ############# PLOT CONTOURS HMC-SBI-PS  #############
+%--------------------------------------------------------------------
 \begin{figure*}
@@
 \end{figure*}
-
+%--------------------------------------------------------------------
+% ############# FIGURE OF MERIT TABLE  #############
+%--------------------------------------------------------------------
 \begin{center}
@@
  \hline
-    FoM &  & Explicit Inference  &  & & Implicit inference  \\
-     & $C_{\ell}$ & Full Field (HMC)&  VMIM & MSE & MAE & GNLL   \\
-  \hline
+    FoM & $C_{\ell}$ & Full Field (HMC)&  VMIM & MSE & MAE & GNLL   \\
+  \hline\hline
  $\Omega_c-  \Omega_b$ & 5315 & 6951 & 6591 & 5651 & 5909\\
@@
  $\Omega_b-  \sigma_8$ & 3997 & 5472 & 5514 & 5066 & 5398\\
- $\Omega_b-  h_0$      & 2800 & 2994 & 2805 & 2923 & 2979\\
- $\Omega_b-  n_s$      & 4035 & 4516 & 4174 & 4073 & 4259\\
  $\Omega_b-  w_0$      & 523  & 846  & 791  & 736  & 771\\
@@
  $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
- $h_0-       n_s$      & 425  & 502  & 458  & 455  & 469\\
  $h_0-       w_0$      & 52   & 88   & 83   & 80   & 82\\
- $n_s-       w_0$      & 80   & 133  & 126  & 112  & 119\\
+ $n_s-       w_0$      & 80   & 133  & 126  & 112  & 119 \\
      \hline
@@
 %--------------------------------------------------------------------
-\section{Conclusions}
+%  ############# SUMMARY TABLE  #############
+%--------------------------------------------------------------------
+\begin{table*}
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+    \hline
+    \multirow{2}{*}{} &
+      \multicolumn{2}{c|}{VMIM} &
+      \multicolumn{2}{c|}{MSE} &
+      \multicolumn{2}{c|}{MAE} &
+      \multicolumn{2}{c|}{GNLL} \\
+    & Mean & $\sigma$ & Mean & $\sigma$ & Mean & $\sigma$  & Mean & $\sigma$ \\
+    \hline
+    $\Omega_c$ & 0.2742 & 0.026 & 0.2830 & 0.030 & 0.2793 & 0.029  & 2.1\% & 2.1\%\\
+    \hline
+    $\Omega_b$ & 0.0499 & 0.006 & 0.0492 & 0.006 &  0.0501 & 0.006 & 2.1\% & 2.1\% \\
+    \hline
+    $\sigma_8$ & 0.819 & 0.03 &0.808 & 0.03 &  0.808 &0.03 & 2.1\% & 2.1\% \\
+    \hline
+    $w_0$      &-0.998 & 0.2 &-1.13 & 0.2 & -1.16& 0.2  & 2.1\% & 2.1\% \\
+    \hline
+    $h_0$      & 0.6658 & 0.058 &0.6620 & 0.057 & 0.6640 & 0.057&  2.1\% & 2.1\% \\
+    \hline
+    $n_s$      & 0.9635 & 0.04 & 0.9559 & 0.04 & 0.9547 & 0.04  & 2.1\% & 2.1\% \\
+    \hline
+  \end{tabular}
+\end{table*}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+\section{Conclusions}\label{Sec:conclusion}
 %--------------------------------------------------------------------
@@
 \end{equation}
-we obtain 
-\begin{equation}\label{Eq:definition_median}
-\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
-\end{equation}
+we obtain \autoref{Eq:definition_median}.
+% \begin{equation}
+% \int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
+% \end{equation}
 %--------------------------------------------------------------------

### 2023-07-10 paper/paper.tex

@@
 \usepackage{booktabs}
+\usepackage{fontawesome}
+\newcommand{\github}{\href{https://github.com/dlanzieri/WL_Implicit-Inference}{\faGithub}} %For Denise: Remember to label the final version of the Github repo with the notebook
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{Likelihood Implicit Methods}
+\title{Optimal Neural Summarisation for Full-Field Implicit Inference by Density Estimation }
 
@@
   % {} leave it empty if necessary  
-   {blabla}
+   {In many cosmological applications, the likelihood function is either unknown or intractable, limiting the use of traditional inference methods.
+Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data to approximate the posterior (or likelihood) distribution instead of evaluating an explicit likelihood function.
+However, these methods suffer from the curse of dimensionality, necessitating the development of compression schemes to reduce high-dimensional data into lower-dimensional summary statistics.
+
+   \justine{\textbf{I think the 2 main ideas for this part are: }
+
+    - classical way to overcome this issue: we compress the data into 2pt statistics  and use gaussian likelihood → pb: we miss a lot of information}
+    \denise{That's true. But adding it in the abstract would make the "Context" subsection (this division is mandatory for A$\&$A) too long. This is properly described in the introduction}
+
+    \justine{ - considering the previous pb: thanks to Implicit Inference methods we can consider any kind of summary stat (because we can learn the exact posterior p(theta | summary stat)), so to have the best contours (meaning that we exatrct all gaussian as well as non-gaussian signal) we only have to focus on how we compress our data.}
+    \denise{That's also true.
+But we don't use any summary statistics (except the power spectrum because it's easy), and I think that one of the points of the paper is to extract the information from the pixels, without using any summaries.
+So I personally would not talk about summary statistics }
+
+     \justine{\textbf{Note:}
+
+    - for me LFI = Implicit Inference, I think we should use only one
+   }
+   \denise{True again.
+But the definition "Likelihood-Free Inference" is more used in the community.
+Implicit Inference is something that we introduce in the motivation and introduction.
+I don't know exactly how much people are familiar with this concept.
+}
+    \justine{ - 'Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data instead of an explicit likelihood function.' → Implicit inference approaches offer a promising solution to approximate posterior distribution by using simulations from an implicit likelihood (a simulator) instead of evaluating explicit likelihood function ..?
+
+    - I'm probably wrong but not sure about the last sentence.
+}
+   }   
   % aims heading (mandatory)
-    {blabla}
+    {This paper presents a comparative analysis of various neural network-based compression schemes within the LFI framework for weak lensing analysis.
+Specifically, we focus on evaluating the performance of different loss functions employed during the training of the neural network and their impact on the forecasted constraints for the $\Lambda$CDM parameters expected from LSST-Y10.
+To validate our findings, we compare the posterior distributions obtained through LFI with those derived from a Bayesian forward modeling approach.
+\justine{instead of 'Bayesian forward modeling approach' maybe we can say field-based Bayesian Hierarchical Model analysis performed through HMC (considered as our ground truth)?}}
   % methods heading (mandatory)
-    {blabla}
+    {For both forward-modeling the convergence field and generating observed mock data, we use the \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}} package. 
+    This package enables posterior inference on cosmological parameters in a tomographic setting, accounting for cross-correlations between different redshift bins.
+\texttt{SbiLens} is a jax-based differentiable physical model specifically designed for Likelihood Free Inference and Bayesian inference algorithms that require access to likelihood derivatives.
+
+    \justine{\textbf{I think the main ideas for this part are: }
+
+    - first we describe sbi\_lens : Provides a differentiable log normal mass map simulator specifically designed for for weak lensing inference.
+Then you can describe a bit the simulations like tomographic bins etc I guess
+
+    - We construct our ground truth
+
+    - We use implicit inference techniques to go from the compressed stat to the posteriors}
+    }
   % results heading (mandatory)
-   {blabla}
+   {Our experiments validate the effectiveness of LFI methods,  as we demonstrate comparable posterior distributions between LFI and Bayesian forward modeling approaches \justine{field-based Bayesian Hierarchical Model analysis}.
+Additionally, our results provide insights into the design of data compression procedures to derive low-dimensional summary statistics while minimizing information loss.
+All codes associated with this paper can be found at this link \github.}
   % conclusions heading (optional), leave it empty if necessary 
    {}
-
    \keywords{methods: statistical – gravitational lensing: weak – cosmology: large-scale structure of Universe
@@
 \section{Introduction}
+\denise{PEOPLE CAN READ THIS!!}
 Weak gravitational lensing by the large-scale structure is caused by the presence of foreground matter bending the light emitted by background galaxies.
@@
 An alternative framework for inference is provided by Likelihood-Free Inference (LFI) approaches, which are sometimes referred to as Likelihood implicit or Simulation-Based Inference (SBI).
-\denise{TODO: cite some other work(?)}
-The potential of LFI methods lies in the fact that they primarily rely on our ability to forward-simulate synthetic data, without the need for an explicit likelihood function.
+Increasingly employed in astrophysics and cosmology \citep{lin2015new, alsing2018generalized, jeffrey2021likelihood, gerardi2021unbiased, tam2022likelihood}, the potential of LFI methods lies in the fact that they primarily rely on our ability to forward-simulate synthetic data, without the need for an explicit likelihood function.
 As a result, these methods are free from assumptions and approximations.
 Furthermore, incorporating systematic effects and complex physical processes into forward simulations is more straightforward compared to including them when constructing the likelihood function.
-Given the potential of LFI methods, it is crucial to identify possible sources of bias and the associated challenges, as well as establish well-defined procedures to validate the results.
+Given the potential of LFI methods, it is crucial to identify possible sources of bias and the associated challenges.
 For instance, it is widely recognized that even the most advanced LFI methods can be affected by the curse of dimensionality.
@@
 A significant question is how to design data compression procedures, aiming to derive low-dimensional summary statistics while minimizing information loss.
+Moreover, it is imperative to establish well-defined procedures to validate the results.
 
@@
 These methods differ in terms of the loss functions used to train the neural network, but they share the same inference strategy based on neural density estimation.
-Moreover, a crucial point is to validate the results obtained in the LFI context.
-To address this, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
-The forward model is based on \texttt{SbiLens}, a Python package designed for Weak Lensing Implicit Inference with Differentiable Simulator implemented in \texttt{Jax}.
+To address the need for result validation in the LFI context, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
+The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}}, a Python package designed for weak lensing implicit inference with differentiable simulator implemented in Jax.
 \texttt{SbiLens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
@@
 Finally, we conclude in \autoref{Sec:conclusion}.
-
 %------------------------------------------------------------------
@@
 With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, that may not fully capture the non-Gaussian information present in the lensing field at the scales accessible to future surveys.
-One of the main objectives of this paper is to introduce a forward model that directly extracts information from the raw pixel data, rather than relying on the analytical evaluation of summary statistics.
+In this paper we introduce a forward model that directly extracts information from the raw pixel data, rather than relying on the analytical evaluation of summary statistics.
 By doing so, we aim to preserve all available information and facilitate the incorporation of systematic effects and the combination of multiple cosmological probes through joint simulations.
 In this context, the simulator of the observables serves as our physical model, where each component is tractable.
-These models, also knows as \textit{probabilistic program}, considering a prior distribution $p(\bm{\theta})$ for the parameter $\bm{\theta}$, sample \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, generating the output $\bm x \sim p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
-However, computing the marginal likelihood is typically intractable:
+These models, often refereed to as \textit{probabilistic program}, can be described as follows.
+They take as input a vector parameter $\bm{\theta}$, which describes the underlying deterministic model.
+Then, they sample internal states, dubbed \textit{latent variables}, from the distribution $\bm{z}_i \sim p_i(\bm{z}_i|\bm \theta, \bm{z}_{<i})$, than can be directly or indirectly related to a physically meaningful state of the system.
+Finally, they generate the output $\bm x$z from the distribution $\ p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
+The ultimate goal of Bayesian inference in cosmology is to compute the posterior distribution:
+\begin{equation}\label{Eq:posterior}
+     p(\bm{\theta}|\bm{x})= 
+     \frac{p(\bm{x}|\bm{\theta})p(\bm{\theta})}
+     {\int d\bm{\theta'}p(\bm{x}|\bm{\theta}')p(\bm{\theta}')},
+\end{equation}
+However, a problems arises because the marginal likelihood $p(\bm{x}|\bm{\theta})$ is typically intractable:
 \begin{equation}
@@
 \end{equation}
-To overcome this limitation while still capturing the full information content of the data, different approaches have been proposed in the literature.
-Although these approaches are often referred to by different names, in this work, we will make the following distinction
+as it corresponds to an integral over all possible trajectories throughout the latent space.
+To overcome this limitation while still capturing the full information content of the data, two different approaches have been proposed in the literature.
+Although these approaches are often referred to by different names, in this work, we will make the following distinction:
 %--------------------------------------------------------------------
 \paragraph{\textbf{Explicit inference}} referring to all Likelihood-based inference approaches.  
-In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Model (BHM).
-As a Bayesian forward approach it involves using a given model to predict observations and then comparing these predictions with real observations to infer the parameters of the model.
-The \textit{hierarchical} nature of this methodology comes from the fact that a complex inference task, such as the weak lensing inference task, can be broken down into several hierarchies of elements, which can be understood and from which we can sample from \citep{heavens2018bayesian}.
-However, performing a global analysis on complex and large datasets, such as what could be a lensing dataset, is not feasible.
-Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
 This approach involves treating the simulator as a probabilistic model and performing inference over the joint posterior:
@@
 \end{equation}
+In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Models (BHMs).
+A Bayesian forward approach involves using a given physical model to predict observations and then comparing these predictions with observations to infer the parameters of the model.
+However, performing a global analysis on complex and large datasets, such as a lensing dataset, is not feasible.
+Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
+From here, the \textit{hierarchical} nature of this methodology
 %--------------------------------------------------------------------
 \paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations.
-This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution
+This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution:
 \begin{equation}
@@
 \end{equation}
-Within this class of methods, we can distinguish between more traditional methods such as the Approximate Bayesian Computation (ABC), a rejection-criteria based approach, which approximate the likelihood comparing the simulations with data, and the Density Estimation Likelihood Free Inference (DELFI), methods, where the inference task is approached as a density estimation problem.
-Ultimately, for a given simulation model, the two approaches should converge to the same posterior.
+Within this class of methods, we can differentiate between more traditional methods such as Approximate Bayesian Computation (ABC) and Density Estimation Likelihood-Free Inference (DELFI) methods.
+ABC employs rejection-criteria based approaches to approximate the likelihood by comparing simulations with data.
+In this work, our focus will be on DELFI methods, which approach the inference task as a density estimation problem.
+Specifically, we can divide Likelihood-Free Inference approaches into two separate steps:
+\begin{enumerate}
+    \item  Automatically learning an optimal low-dimensional summary statistic.
+    \item Using Neural Density Estimation in low dimensions to infer the target distributions.
+\end{enumerate}
+In the first step, we introduce a parametric function $f_{\varphi}$
+     \begin{equation}
+         \bm{y}=f_{\varphi}(\bm{\theta}),
+     \end{equation}
+which aims to reduce the dimensionality of the data while preserving information. Typically, $\bm{y}$ is assumed to have the same dimension of $\bm{\theta}$. 
+In the second step, Neural Density Estimation can target either building an estimate $p_{\phi}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (referred to as the Neural Likelihood Estimation (NLE) task), or targeting the posterior distribution $p(\bm{\theta}|\bm{x})$, (known as Neural Posterior Estimation (NPE) task).
+
+It is important to note that not all neural compression techniques are equivalent.
+Many papers that apply neural techniques for extracting summary statistics have used sub-optimal compression techniques, such as Mean Square Error or Maximum Absolute Error.
+However, it is also important to consider that in this case, the compression of summaries will be sub-optimal, resulting in a loss of information (larger contours), but it will not introduce bias in the final constraints.
+On the contrary, many other papers rely on assuming proxy Gaussian likelihoods and estimate the mean and covariance of these likelihoods from simulations.
+While likelihood-free inference using MSE and MAE summaries is robust in the sense that the approximations made in the compression steps will not introduce bias to the parameter inferences, the same cannot be said for summaries obtained from Gaussian likelihood approximations.
+We summarize the different neural compression strategies found in the literature in \autoref{tab:biblio_survey}.
+
+\vspace{1cm}
+Ultimately, for a given simulation model, if an optimal compression statistic is used, the two approaches should converge to the same posterior. Therefore, the goals of this paper will be:
+\begin{enumerate}
+    \item Find the optimal statistics.
+    \item Demonstrate that by using this optimal compression strategy, both the implicit and explicit methods yield comparable results.
+\end{enumerate}
+
 %------------------------------------------------------------------
@@
 \section{The SBILens framework}\label{Sec:the SBILens framework}
+\denise{PEOPLE CAN READ THIS!!}
 \subsection{Lognormal Modeling}
@@
 %--------------------------------------------------------------------
-Although this is not a full-field approach, we aim to compare the latest to a more standard approach. 
 To obtain the probability distribution of the cosmological parameters using a summary-statistics-methods, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
@@
 \subsubsection{Benchmark compression scheme}
+\denise{PEOPLE CAN READ THIS!!}
 To ensure the scalability of density estimation LFI in cases where forward simulations are computationally expensive, it becomes necessary to employ compression techniques that reduce the dimensionality of the data space and extract summary statistics.
@@
 \subsubsection{Figure of Merit (FoM)}
+%--------------------------------------------------------------------
 To quantify the constraining power of each inference strategy, we first employ the Figure of Merit (FoM), defined as follows:
@@
 Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalised parameter space of the $\alpha$ and $\beta$ pair.
+%--------------------------------------------------------------------
 \subsubsection{Kullback-Leibler divergence}
 %--------------------------------------------------------------------
+To measure the difference between the probability distribution obtained with different compressed summaries we compute the \textit{Kullback-Leibler divergence} or KL-divergence.
+As mentioned when we introduced the VMIM, the KL divergence is related to the relative entropy, and is a non-symmetric measure of the difference between two distributions.
+\denise{INCOMPLETE}
+%--------------------------------------------------------------------
+% ############# Begin Table BiblioSruvey  #############
+%--------------------------------------------------------------------
+
 \begin{center}
@@
  \hline
-            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
+            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
              \\
@@
 %--------------------------------------------------------------------
+% ############# End Table BiblioSruvey  #############
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
 \section{Results}\label{Sec:results}
@@
   \hline\hline
- $\Omega_c-  \Omega_b$ & 5315 & 6951 & 6591 & 5651 & 5909\\
  $\Omega_c-  \sigma_8$ & 1222 & 2520 & 2526 & 2043 & 2316\\
- $\Omega_c-  h_0$      & 516  & 698  & 675  & 601  & 618\\
- $\Omega_c-  n_s$      & 898  & 1243 & 1184 & 964  & 1030\\
  $\Omega_c-  w_0$      & 100  & 198  & 190  & 152  & 162\\
- $\Omega_b-  \sigma_8$ & 3997 & 5472 & 5514 & 5066 & 5398\\
- $\Omega_b-  w_0$      & 523  & 846  & 791  & 736  & 771\\
- $\sigma_8-  h_0$      & 402  & 566  & 577  & 544  & 571\\
- $\sigma_8-  n_s$      & 583  & 889  & 891  & 787  & 854\\ 
  $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
- $h_0-       w_0$      & 52   & 88   & 83   & 80   & 82\\
- $n_s-       w_0$      & 80   & 133  & 126  & 112  & 119 \\
+
      \hline
@@
     & Mean & $\sigma$ & Mean & $\sigma$ & Mean & $\sigma$  & Mean & $\sigma$ \\
+    \hline
     \hline
@@
 \bibliography{paper/biblio} 
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+
+\textbf{Comments:}
+\vspace{5mm}
+
+\begin{itemize}
+    \item \justine{Regarding the table with the differents losses - \citet{2018PhRvD..97j3515G} also used weighted MAE: 'Errors in predictions for maps from cosmologies in sparsely sampled regions are more severely penalized than those for maps from densely sampled re- gions.
+We show in §IV C that such a weighted loss func- tion reduces the bias in the predictions, at the cost of a longer network training, but has only a limited impact on the parameter constraints inferred from the predictions.'
+    I think it's because they don't have honmogeneous dataset so if needed we can added a note to say that we are in the case of homogeneous dataset and thus don't have bias coming from this etc.
+    + for inference startegy maybe we can say that It's gaussian likelihood with fixed cov matrix (evaluated at the fiducials) analysis.
+    \item \citet{fluri2018cosmological} I think they use gaussian log likelihood loss where they learn both the mean and cov matrix: 'Similar to [26], we chose a negative log-likelihood loss as cost function' page 5.
++ because of bias issues they add regularization term at some point in the training.
+    Then for the inference part they use a gaussian likelihood with mean and covariance matrix como dependent based on their CNN predictions.}
+    \denise{yes, the use gnll, it was just a typo. 
+However, they do not use the covariance matrix from the CNN predictions "
+We did not use the network predictions of the covariance matrices for our main analysis. One reason was that
+our examined test region was consisting out of four independent patches and the input of the network was only
+one patch at a tim"} 
+\justine{Yes you're right!
+They still present the case where you can use the covariance matrix, I mean in some way they try to pay attention to it.
+Idk if you plan to describe a bit this table, if yes I think you can say that they propose this but end up just using the mean which is the exact same as doing mse etc...
+?}
+\denise{yep yep, the table is only a draft, and I didn't mention it in the paper yet. I will definitely describe it in more detail. Ps, sorry for the re-writing, I have OCD with latex error message}
+\justine{yes ok no problem :) I was just checking the table since I'm currently reading these papers for our second project and did not know where to make these comments haha}
+\LM{Ok sweet I was confused about this -- Fluri et al talk about this loss function (which I'm using in another project ahaha) but if they don't make use of the network Covariance it's just a weirdly-weighted MSE (also I wasn't convinced about their training routine in that paper -- the losses didn't seem converged)}
+\item \justine{\cite{fluri2021cosmological} From my understanding, what they use is not exactly the imnn in the sense that they don't rely on gaussian assumptions for the summary distribution.}
+\item \LM{Nope they don't use IMNN.
+They're using a regression-gnll loss, but they don't show any optimality guarantees using this form.
+Plus, they don't make use of the covariance in a particularly smart way (e.g.
+to reweight summary distributions for posterior construction) so it's not super useful.
+I'll be sure to explicitly show the difference in Fluri et al and the IMNN in the corresponding section.}
+\justine{Awesome, thanks! :)}
+\end{itemize}
+
 %--------------------------------------------------------------------

### 2023-07-11 paper/paper.tex
11/07

@@
   % aims heading (mandatory)
-    {This paper presents a comparative analysis of various neural network-based compression schemes within the LFI framework for weak lensing analysis.
-Specifically, we focus on evaluating the performance of different loss functions employed during the training of the neural network and their impact on the forecasted constraints for the $\Lambda$CDM parameters expected from LSST-Y10.
-To validate our findings, we compare the posterior distributions obtained through LFI with those derived from a Bayesian forward modeling approach.
-\justine{instead of 'Bayesian forward modeling approach' maybe we can say field-based Bayesian Hierarchical Model analysis performed through HMC (considered as our ground truth)?}}
+    {We aim to derive the optimal neural-compression strategy for full field Likelihood free inference by density estimation.
+Moreover, by using the optimal compression strategy, we demonstrate that the posterior distribution is comparable to those derived from a Bayesian forward modeling approach.}
   % methods heading (mandatory)
-    {For both forward-modeling the convergence field and generating observed mock data, we use the \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}} package. 
-    This package enables posterior inference on cosmological parameters in a tomographic setting, accounting for cross-correlations between different redshift bins.
+    { We present a comparative analysis of different loss functions employed during the training of the neural network.
+We  evaluate their performance by measuring their impact on the constraints on the $\Lambda$CDM parameters expected from LSST-Y10.
+    For both forward-modeling the convergence field and generating observed mock data, we use the \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}} package. 
     \texttt{SbiLens} is a jax-based differentiable physical model specifically designed for Likelihood Free Inference and Bayesian inference algorithms that require access to likelihood derivatives.
@@
 Then, they sample internal states, dubbed \textit{latent variables}, from the distribution $\bm{z}_i \sim p_i(\bm{z}_i|\bm \theta, \bm{z}_{<i})$, than can be directly or indirectly related to a physically meaningful state of the system.
-Finally, they generate the output $\bm x$z from the distribution $\ p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
+Finally, they generate the output $\bm x$ from the distribution $\ p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
+
 The ultimate goal of Bayesian inference in cosmology is to compute the posterior distribution:
@@
 \end{equation}
-However, a problems arises because the marginal likelihood $p(\bm{x}|\bm{\theta})$ is typically intractable:
+however, a problems arises because the marginal likelihood $p(\bm{x}|\bm{\theta})$ is typically intractable:
 \begin{equation}
@@
 Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.
-From here, the \textit{hierarchical} nature of this methodology
+From here, the \textit{hierarchical} nature of this methodology.
 %--------------------------------------------------------------------
@@
 ABC employs rejection-criteria based approaches to approximate the likelihood by comparing simulations with data.
-In this work, our focus will be on DELFI methods, which approach the inference task as a density estimation problem.
-Specifically, we can divide Likelihood-Free Inference approaches into two separate steps:
+In this work, our focus will be on DELFI methods, which approach the inference task as a density estimation problem. \\
+We can divide Likelihood-Free Inference approaches into two separate steps:
 \begin{enumerate}
@@
 \end{enumerate}
-In the first step, we introduce a parametric function $f_{\varphi}$
+In the first step, we introduce a parametric function $f_{\varphi}$ such that:
      \begin{equation}
@@
      \end{equation}
-which aims to reduce the dimensionality of the data while preserving information. Typically, $\bm{y}$ is assumed to have the same dimension of $\bm{\theta}$. 
+which aims to reduce the dimensionality of the data while preserving information. Typically, the compressed statistics $\bm{y}$ is assumed to have the same dimension of $\bm{\theta}$. \\
 In the second step, Neural Density Estimation can target either building an estimate $p_{\phi}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (referred to as the Neural Likelihood Estimation (NLE) task), or targeting the posterior distribution $p(\bm{\theta}|\bm{x})$, (known as Neural Posterior Estimation (NPE) task).
@@
 \begin{enumerate}
-    \item Find the optimal statistics.
+    \item Find the optimal compression strategy.
     \item Demonstrate that by using this optimal compression strategy, both the implicit and explicit methods yield comparable results.
@@
 \denise{PEOPLE CAN READ THIS!!}
-\subsection{Lognormal Modeling}
+\subsection{Lognormal Modeling}\label{Sec:Lognormal Modeling}
 For various cosmological applications, the non-Gaussian field can be modeled as a Lognormal field \citep{coles1991lognormal,bohm2017bayesian}.
@@
 To obtain the probability distribution of the cosmological parameters using a summary-statistics-methods, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
-The correlation between two generic tomographic bins $i$ and $j$ can be expressed as:
-\begin{equation}
-    \left \langle \kappa^{i}(\ell)\kappa^{j*}(\ell)\right \rangle =(2\pi)^2\delta_D(\ell-\ell')C_{\ell}^{ij},
-\end{equation}
 We assume a Gaussian likelihood with a cosmology-independent covariance matrix:
@@
 To compute the expected theoretical predictions $\bm{\mu}(\bm{\theta})$ we use the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}. 
- The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, shown in \autoref{tab:prior}, using the same theoretical library.
-In {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
+ The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, presented in \autoref{tab:prior}, using the same theoretical library.
+Specifically, in {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
 \begin{equation}
@@
 where $f_{sky}$ is the fraction of sky observed by the survey, and $n_s$ is the number density of galaxies. 
-To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map.
+To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map (our fiducial).
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\subsubsection{Full field and HMC}
+\subsubsection{Full field with BHMs}
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-The following steps outline our methods.
-We draw cosmological parameters from the priors illustrated in \autoref{tab:prior}.
-Given values for the cosmological parameters, we generate the log-normal field as described in the previous sections.
-So, in other words, the simulator becomes the physical model generating a non-linear representation of the convergence map.
-The measure of the convergence for each pixel and for each bin predicted from the physical model will differ from the ones from real observations due to the noise.
-This is taken into account in the likelihood.
-Specifically, for LSST Y10 the number of galaxies for each pixel should be sufficiently high that for the central limit theorem, we can assume that the observation is characterized by a Gaussian noise $\sigma_n^2=\sigma_e^2/N_s$, with $N_s$ the total number of source galaxies per bin and pixel.
-Given $\sigma_n^2$ the variance of this Gaussian likelihood, its form in the log form can be expressed as:
+To construct the explicit map-based inference strategy in the Hierarchical Bayesian framework, we built a likelihood based on the data model described in \autoref{Sec:Lognormal Modeling}. 
+As mentioned before, this means that the simulator serves as the physical deterministic model capable of generating the non-linear representation of the convergence map.
+
+However, in practical terms, the measurement of convergence for each pixel and bins will differ from real observations due to noise.
+This is taken into consideration in the likelihood.
+Specifically, for LSST Y10, the number of galaxies for each pixel should be sufficiently high so that, according to the central limit theorem, we can assume the observation is characterized by Gaussian noise, with $\sigma_n^2=\sigma_e^2/N_s$, where $N_s$ represents the total number of source galaxies per bin and pixel.
+Given $\sigma_n^2$ the variance of this Gaussian likelihood, its log-form can be expressed as:
 \begin{equation}
-    \log{\mathcal{L}}(\bm{\theta})=
+    \mathcal{L}(\bm{\theta})=
     \sum_i^{N_{pix}} \sum_{j}^{N_{bins}} \log{P(\kappa^{obs}_{i,j}|\kappa_{i,j},\bm{\theta})}
-    =\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2}
+    =\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2},
 \end{equation}
-Not relying on any summary statistics, the full-field approach typically results in a high-dimensional problem, requiring more sophisticated statistical techniques.
+where $\kappa^{obs}$ refers to the values of convergence from noise maps. \\
+Since the full-field approach does not rely on any summary statistics, it typically leads to a high-dimensional problem, requiring more sophisticated statistical techniques.
 To sample the posterior distribution for $\bm{\theta}$, we use a Hamiltonian Monte Carlo (HMC) algorithm.
 The HMC algorithm is particularly helpful in high-dimensional spaces where a large number of steps are required to effectively explore the space.
-It improves the sampling process by using the information contained in the gradients to guide the sampling process.
-Since the code is implemented in Jax, the gradients are accessible via automatic differentiation.
+It improves the sampling process by leveraging the information contained in the gradients to guide the sampling process.
+As the code is implemented in Jax, the gradients are accessible via automatic differentiation.
 %--------------------------------------------------------------------

### 2023-07-24 paper/paper.tex
Updates from Overleaf

@@
 
-   \justine{\textbf{I think the 2 main ideas for this part are: }
+%    \justine{\textbf{I think the 2 main ideas for this part are: }
 
-    - classical way to overcome this issue: we compress the data into 2pt statistics  and use gaussian likelihood → pb: we miss a lot of information}
-    \denise{That's true. But adding it in the abstract would make the "Context" subsection (this division is mandatory for A$\&$A) too long. This is properly described in the introduction}
+%     - classical way to overcome this issue: we compress the data into 2pt statistics  and use gaussian likelihood → pb: we miss a lot of information}
+%     \denise{That's true. But adding it in the abstract would make the "Context" subsection (this division is mandatory for A$\&$A) too long. This is properly described in the introduction}
 
-    \justine{ - considering the previous pb: thanks to Implicit Inference methods we can consider any kind of summary stat (because we can learn the exact posterior p(theta | summary stat)), so to have the best contours (meaning that we exatrct all gaussian as well as non-gaussian signal) we only have to focus on how we compress our data.}
-    \denise{That's also true.
+%     \justine{ - considering the previous pb: thanks to Implicit Inference methods we can consider any kind of summary stat (because we can learn the exact posterior p(theta | summary stat)), so to have the best contours (meaning that we exatrct all gaussian as well as non-gaussian signal) we only have to focus on how we compress our data.}
+%     \denise{That's also true.
 But we don't use any summary statistics (except the power spectrum because it's easy), and I think that one of the points of the paper is to extract the information from the pixels, without using any summaries.
@@
 
-     \justine{\textbf{Note:}
+%      \justine{\textbf{Note:}
 
-    - for me LFI = Implicit Inference, I think we should use only one
-   }
-   \denise{True again.
+%     - for me LFI = Implicit Inference, I think we should use only one
+%    }
+%    \denise{True again.
 But the definition "Likelihood-Free Inference" is more used in the community.
@@
 }
-    \justine{ - 'Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data instead of an explicit likelihood function.' → Implicit inference approaches offer a promising solution to approximate posterior distribution by using simulations from an implicit likelihood (a simulator) instead of evaluating explicit likelihood function ..?
+%     \justine{ - 'Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data instead of an explicit likelihood function.' → Implicit inference approaches offer a promising solution to approximate posterior distribution by using simulations from an implicit likelihood (a simulator) instead of evaluating explicit likelihood function ..?
 
-    - I'm probably wrong but not sure about the last sentence.
-}
+%     - I'm probably wrong but not sure about the last sentence.
+% }
    }   
@@
 
-    \justine{\textbf{I think the main ideas for this part are: }
+    % \justine{\textbf{I think the main ideas for this part are: }
 
-    - first we describe sbi\_lens : Provides a differentiable log normal mass map simulator specifically designed for for weak lensing inference.
+    % - first we describe sbi\_lens : Provides a differentiable log normal mass map simulator specifically designed for for weak lensing inference.
 Then you can describe a bit the simulations like tomographic bins etc I guess
 
-    - We construct our ground truth
+    % - We construct our ground truth
 
-    - We use implicit inference techniques to go from the compressed stat to the posteriors}
+    % - We use implicit inference techniques to go from the compressed stat to the posteriors}
     }
   % results heading (mandatory)
-   {Our experiments validate the effectiveness of LFI methods,  as we demonstrate comparable posterior distributions between LFI and Bayesian forward modeling approaches \justine{field-based Bayesian Hierarchical Model analysis}.
+   {Our experiments validate the effectiveness of LFI methods,  as we demonstrate comparable posterior distributions between LFI and Bayesian forward modeling approaches.
+   % \justine{field-based Bayesian Hierarchical Model analysis}. 
    Additionally, our results provide insights into the design of data compression procedures to derive low-dimensional summary statistics while minimizing information loss.
@@
 \section{Introduction}
-\denise{PEOPLE CAN READ THIS!!}
 Weak gravitational lensing by the large-scale structure is caused by the presence of foreground matter bending the light emitted by background galaxies.
@@
 To address the need for result validation in the LFI context, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
-The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}}, a Python package designed for weak lensing implicit inference with differentiable simulator implemented in Jax.
+The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}}, a Python package designed for weak lensing inference with differentiable simulator implemented in Jax.
 \texttt{SbiLens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
@@
 They take as input a vector parameter $\bm{\theta}$, which describes the underlying deterministic model.
-Then, they sample internal states, dubbed \textit{latent variables}, from the distribution $\bm{z}_i \sim p_i(\bm{z}_i|\bm \theta, \bm{z}_{<i})$, than can be directly or indirectly related to a physically meaningful state of the system.
+Then, they sample internal states, dubbed \textit{latent variables}, from the distribution $\bm{z}_i \sim p_i(\bm{z}_i|\bm \theta, \bm{z}_{<i})$, that can be directly or indirectly related to a physically meaningful state of the system.
 Finally, they generate the output $\bm x$ from the distribution $\ p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
@@
      \begin{equation}
-         \bm{y}=f_{\varphi}(\bm{\theta}),
+         \bm{t}=F_{\varphi}(\bm{\theta}),
      \end{equation}
-which aims to reduce the dimensionality of the data while preserving information. Typically, the compressed statistics $\bm{y}$ is assumed to have the same dimension of $\bm{\theta}$. \\
+which aims to reduce the dimensionality of the data while preserving information. Typically, the compressed statistics $\bm{t}$ is assumed to have the same dimension of $\bm{\theta}$. \\
 In the second step, Neural Density Estimation can target either building an estimate $p_{\phi}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (referred to as the Neural Likelihood Estimation (NLE) task), or targeting the posterior distribution $p(\bm{\theta}|\bm{x})$, (known as Neural Posterior Estimation (NPE) task).
@@
 \section{The SBILens framework}\label{Sec:the SBILens framework}
-\denise{PEOPLE CAN READ THIS!!}
 \subsection{Lognormal Modeling}\label{Sec:Lognormal Modeling}
@@
 
-As mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales.
-However, the main objective of this paper is to compare different inference strategies and we are not concerned with the potential implications of this approximation.
-In future applications of the presented inference pipeline, the simulation procedure will be based on N-body simulations.
-
 Each map is reproduced on a regular grid with dimensions of $256 \times 256$ pixels and covers an area of $10\times 10$ deg$^2$.
-
+%--------------------------------------------------------------------
+% ############# PRIOR COSMO TABLE  #############
+%--------------------------------------------------------------------
 \begin{table}
-	\centering
-	\caption{ Prior and fiducial values used for the analyses. 
- The symbol $\mathcal{N}_T$ represents a Truncated Normal distribution.
-The lower bound of the support for the $\Omega_c$ distribution is set to zero, while the lower and upper bounds for the $w_0$ distribution are set to -2.0 and -0.3, respectively.}
+	\begin{center}
     	\begin{tabular}{lcc} 
@@
     	\end{tabular}
+        \caption{ Prior and fiducial values used for the analyses. 
+        The symbol $\mathcal{N}_T$ represents a Truncated Normal distribution.
+The lower bound of the support for the $\Omega_c$ distribution is set to zero, while the lower and upper bounds for the $w_0$ distribution are set to -2.0 and -0.3, respectively.}
 	    \label{tab:prior}
+    \end{center}
 \end{table}
@@
 The galaxy sources are divided into 5 tomographic bins, each containing an equal number of galaxies. 
-For each redshift bin, we assume Gaussian noise with mean zero and variance given by
+For each redshift bin, we assume Gaussian noise with mean zero and variance given by:
  \begin{equation}
@@
 \autoref{fig:redshift_distribution} illustrates the resulting source redshift distribution, and \autoref{tab:survey_spec} provides a summary of the survey specifications.
+%--------------------------------------------------------------------
+% ############# SURVEY SPECIFICATION TABLE  #############
+%--------------------------------------------------------------------
 \begin{table}
-	\centering
-	\caption{ LSST Y10 source galaxy specifications in our analysis. All values are based on the LSST DESC SRD.}
+	\begin{center}
     	\begin{tabular}{lc} 
@@
     	\end{tabular}
+     	\caption{ LSST Y10 source galaxy specifications in our analysis. All values are based on the LSST DESC SRD.}
 	    \label{tab:survey_spec}
+    \end{center}
 \end{table}
+%--------------------------------------------------------------------
+% ############# PLOT REDSHIFT DISTRIBUTION  #############
+%--------------------------------------------------------------------
 \begin{figure}
@@
 \subsubsection{Benchmark compression scheme}
-\denise{PEOPLE CAN READ THIS!!}
 To ensure the scalability of density estimation LFI in cases where forward simulations are computationally expensive, it becomes necessary to employ compression techniques that reduce the dimensionality of the data space and extract summary statistics.
@@
 }
-
 %--------------------------------------------------------------------
-
+\subsection{Compression strategy}
+%--------------------------------------------------------------------
+We aim to examine the impact of different loss functions used to train the compressor on the final cosmological constraints.
+We employ the same convolutional neural network architecture to build the compressor.
+Specifically, we utilize a deep ResNet architecture \citep{he2016deep}, the ResNet-18, implemented using Haiku \citep{haiku2020github}, a Python deep learning library built on top of JAX.
+While the different compression strategies share the same architecture, the training strategies for VMIM and GNLL differ significantly.
+To train the Neural compressor under VMIM and GNLL, we jointly optimize the weights of the Neural Network $F_{\varphi}$ (the compressor) and the parameters of the variational distribution $q_{\varphi'}$.
+For VMIM, the variational distribution $q(\bm{\theta}|\bm{t};\varphi')$ is modeled using a Normalizing Flow, while for GNLL, a Conditional Multivariate Gaussian estimator is employed.
+The Normalizing Flow used for the inference strategy is also adopted for the variational distribution required in the compression procedure.
+For the Multivariate Gaussian estimator, we implement a three-layer neural network.
+The first two layers consist of fully connected layers with a Leaky Rectified Linear Unit activation function and a hyperbolic tangent activation function, respectively.
+The number of channels is increased to 256 and 128 for the first two layers, respectively.
+The last layer is a fully connected layer without an activation function, responsible for mapping the output of the first two layers to the desired output size (mean and covariance of the Gaussian distribution).
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\subsubsection{Inference strategy}
+%--------------------------------------------------------------------
+\subsubsection{Inference strategy}\label{Sec:Inference_strategy}
 %--------------------------------------------------------------------
@@
 Since the likelihood $p(t|\theta)$ is intractable, we propose to do Neural Posterior Estimation \citep{npe1, npe2, npe3}, a Simulation-Based Inference (SBI) strategy that aims to directly approximate, from simulations, the posterior distribution
-
 \begin{equation}
@@
 %--------------------------------------------------------------------
-
+\begin{table*}
     \begin{center}
-\begin{table*}
-\begin{tabular}{ |p{3.5cm}|p{2cm}|p{3cm}|p{2.5cm}|p{4cm}|  }
+        \begin{tabular}{ |p{3.5cm}|p{3.5cm}|p{3.5cm}|p{4cm}|  }
             \hline
-Reference & \makecell{Architecture\\ compressor}  & \makecell{Loss function} & \makecell{Inference \\ strategy} & \makecell{Output compressor}   \\
+            Reference &  \makecell{Loss function} & \makecell{Inference \\ strategy} & \makecell{Output compressor}   \\
             \hline
-            \citet{2018PhRvD..97j3515G} & \makecell{CNN} & \makecell{MAE} &  \makecell{Likelihood \\ analysis} &\makecell{$\sigma_8, \Omega_m$}  \\
+            \citet{2018PhRvD..97j3515G} & \makecell{MAE} &  \makecell{Likelihood \\ analysis} &\makecell{$\sigma_8, \Omega_m$}  \\
             \hline
-            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
+            \citet{fluri2018cosmological} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
              \\
@@
             \rowcolor{lightgray}
-            \citet{fluri2019cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m, A_{IA}/10, L$  \\ with  $L:\Sigma^{-1}=LL^{T}$}
+                        \citet{fluri2019cosmological} &  \makecell{GNLL} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m, A_{IA}/10, L$  \\ with  $L:\Sigma^{-1}=LL^{T}$}
             \\
             \hline            
-            \citet{ribli2018improved} & \makecell{CNN} & \makecell{MSE}  &\makecell{RMSE for \\ evaluation}  & \makecell{$\sigma_8, \Omega_m$}   
+                        \citet{ribli2018improved} & \makecell{MSE}  &\makecell{RMSE for \\ evaluation}  & \makecell{$\sigma_8, \Omega_m$}   
             \\
             \hline            
-            \citet{ribli2019weak} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
+                        \citet{ribli2019weak} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
             \\            
             \hline             
-            \citet{PhysRevD.102.123506} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
+                        \citet{PhysRevD.102.123506} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
             \\
@@
             \rowcolor{lightgray}
-            \citet{jeffrey2021likelihood} & \makecell{\makecell{CNN} \\\makecell{CNN}+NF} & 
+                        \citet{jeffrey2021likelihood} & 
                         \makecell{MSE \\ VMIM}
@@
             \hline             
-            \citet{fluri2021cosmological} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
+                        \citet{fluri2021cosmological} &\makecell{FIM} & \makecell{GPABC} &  
             \\
@@
 
-            \citet{fluri2022full} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
+                        \citet{fluri2022full} &  \makecell{FIM} & \makecell{GPABC} &  
             \\            
             \hline 
-            \citet{lu2022simultaneously} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8, A_{IA}/10,$ \\ rescaled \\ baryonic parameters \\ $(M_c,M_{1,0}, \eta, \beta)$}
+                        \citet{lu2022simultaneously} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8, A_{IA}/10,$ \\ rescaled \\ baryonic parameters \\ $(M_c,M_{1,0}, \eta, \beta)$}
             \\           
             \hline 
-            \citet{kacprzak2022deeplss} & \makecell{CNN} & \makecell{GNLL}  & \makecell{Likelihood \\ analysis}  & \makecell{WL: $\Sigma, \Omega_m, \sigma_8, A_{IA}, \eta_{IA}$ \\  GC:$\Sigma, \Omega_m, \sigma_8, b_g, r_g,\eta_{b_{g}}$ }
+                        \citet{kacprzak2022deeplss} & \makecell{GNLL}  & \makecell{Likelihood \\ analysis}  & \makecell{WL: $\Sigma, \Omega_m, \sigma_8, A_{IA}, \eta_{IA}$ \\  GC:$\Sigma, \Omega_m, \sigma_8, b_g, r_g,\eta_{b_{g}}$ }
 
@@
             \rowcolor{lightgray}
-            \citet{lu2023cosmological} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8,A_{IA}/10,$\\ rescaled baryonic \\ parameters: $M_c,M_{1,0}, \eta, \beta$}
+                        \citet{lu2023cosmological} &  \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8,A_{IA}/10,$\\ rescaled baryonic \\ parameters: $M_c,M_{1,0}, \eta, \beta$}
         \end{tabular}
+        \caption{Table summarizing the current state of the art for neural compression schemes.
+The table includes all studies conducted within the context of Implicit analysis and standard Likelihood analysis.
+The shaded gray region indicates analyses performed on real observations.
+\\
+        Abbreviations used in the Table: MSE-Mean Square Error; MAE-Maximum Absolute Error; GNLL- Gaussian Negative Log Likelihood; VMIM- Variational Mutual Information Maximization; FIM- Fisher Information Matrix; GPABC-Gaussian Processes Approximate Bayesian Computation.
+WL- Weak Lensing; GC-Galaxy Cluster.}
         \label{tab:biblio_survey}
+    \end{center}
 \end{table*}
-\end{center}
 %--------------------------------------------------------------------
@@
 We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10.
-The forecasted results are obtained using the simulation procedure outlined in and the parameter inference strategy described in .
-We begin by comparing the outcomes of three different inference procedures: Explicit 2-point statistics, Explicit full-field statistics, and Implicit statistics using CNN summaries.
-Subsequently, we analyze the impact of the different compression strategies on the final cosmological constraints.
+The forecasted results are obtained using the simulation procedure outlined in \autoref{Sec:the SBILens framework} and the parameter inference strategy described in \autoref{Sec:Inference_strategy}.
+We begin by analyzing the impact of the different compression strategies on the final cosmological constraints.
+Subsequently, we compare the outcomes of three different inference procedures: the 2-point statistics, the explicit full-field statistics, and the implicit full-field statistics using CNN summaries.
+%--------------------------------------------------------------------
+\subsection{Optimal compression strategy}
+Figure X shows the posterior contours from one fiducial map using the compressed summaries.
 %--------------------------------------------------------------------
@@
 We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard-two point statistics and two map-based approaches, the explicit inference using an HMC and the implicit inference using the Normalizing Flow as an estimator.
-As outlined before, our interest is to prove that the two map-based approaches lead to very comparable posterior distributions.
-We present the constraints in the full $\Lambda$CDM parameter space in \autoref{fig:contours_posterior_imp_ex_ps}.
+As outlined before, our primary interest is to prove that the two map-based approaches lead to very comparable posterior distributions.
+We present the $68.3\%$ and $95.5\%$ confident regions for the full $\Lambda$CDM parameters in \autoref{fig:contours_posterior_imp_ex_ps}.
 The contours obtained by the angular $C_{\ell}$ analysis are plotted in blue, the ones for the HMC in yellow, and those for the Normalizing Flow in violet.
-Confirming previous work, we find the two-point statistics are suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$.
-$h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches.
- We quantify the results by computing the Figure of Merit as described in \autoref{Eq:Figure_of_merit}. The results are presented in \autoref{tab:f_o_m}. 
-A further investigation is represented by 
--show the results for the KL divergence 
--present a comparison with other works. 
+We can see from the figure that the contours from the HMC-based inference and the Normalizing flow-based inference are remarkably similar.
+We quantify the results by computing the Figure of Merit as described in \autoref{Eq:Figure_of_merit}.
+The results are presented in \autoref{tab:f_o_m}.
+The
+The remarkably strong agreement between the two posteriors confirms that the two map-based cosmology inference methods yield the same results.
+This establishes the validity of the implicit inference strategy.
+Moreover, we find that the two-point statistic is suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$, while the two map-based approaches yield much tighter constraints on these parameters.
+Additionally, for the full-field strategies we find that the size of the contours is significantly smaller than the size for the prior distributions adopted, suggesting that the
+posterior on these parameters are not affected by the specific choice of prior.
+Confirming previous work, we note that $h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches. 
+We find that the map-based explicit and implicit strategies lead to an improvement in the figure of merit of 2.06, 1.98, 2.28 and 2.06, 1.90, 2.22, respectively, in the $\Omega_c-\sigma_8; \Omega_c-w_0; \sigma_8-w_0$ plane.
+
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
+\begin{table*}
     \begin{center}
-\begin{table*}
-\caption{ Values of the Figure of Merit (FoM) as defined in \autoref{Eq:Figure_of_merit} for different inference strategies: the convergence power spectrum $C_{\ell}$, the HMC, the CNN map compressed statistics with the MSE, the VMIM, the MAE and the GNLL loss functions.
-The values of the figure of merit are inversely proportional to the area of the contours, larger the FoM, higher the constraining power.}
         \begin{tabular}{lcccccc} 
@@
             $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
-
             \hline
         \end{tabular}
+        \caption{ Values of the Figure of Merit (FoM) as defined in \autoref{Eq:Figure_of_merit} for different inference strategies: the convergence power spectrum $C_{\ell}$, the HMC, the CNN map compressed statistics with the MSE, the VMIM, the MAE and the GNLL loss functions.
+The values of the figure of merit are inversely proportional to the area of the contours, larger the FoM, higher the constraining power.}
         \label{tab:f_o_m}
+    \end{center}
 \end{table*}
-\end{center}
 %--------------------------------------------------------------------
@@
 \begin{table*}
+    \begin{center}
         \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
@@
         \end{tabular}
+        \caption{Marginalised mean and standard deviation of the parameter distributions.}
+  \end{center}
 \end{table*}
@@
 %--------------------------------------------------------------------
-\section{Conclusions}\label{Sec:conclusion}
+\section{Summary and conclusions}\label{Sec:conclusion}
+ In this work, we presented and compared two different map-based strategies for inferring cosmological parameters: the explicit full-field strategy, also known as Bayesian Hierarchical inference, based on an HMC sampler, and the implicit inference strategy, also known as Likelihood-Free Inference.
+\\Deep learning approaches for implicit inference typically involve two steps: the automatic learning of an optimal low-dimensional summary statistic and the use of a Neural Density Estimator in low dimensions to either build an estimate $P_{\varphi}$ of the likelihood function $p(x|\theta)$ (Neural Likelihood Estimation) or build an estimate $P_{\varphi}$ of the posterior distribution $p(\theta|x)$ (Neural Posterior Estimation).
+
+ The primary goal of this paper was to find an optimal compression strategy and demonstrate that, by using this strategy, both implicit and explicit methods yield the same posterior. \\
+To both forward model the convergence maps and simulate the mock data used to train the implicit model, we developed \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}}, a Jax-based weak lensing simulator optimized for inference applications requiring access to the derivatives of the model.
+Our analysis is based on synthetic weak lensing data with five tomographic bins for a survey like LSST-Y10.
+\\
+We begin by providing an overview of the different compression strategies adopted in the literature for both Likelihood-Free Inference and likelihood-based inference strategies.
+We then compare the impact of these different loss functions on the final constraints on the cosmological parameters for a $\Lambda$CDM model.
+We found the following results:
+\begin{enumerate}
+    \item The Mean Square Error and the Maximum Absolute Error lead to significantly similar results, while the VMIM (Variational-Mutual Information Maximization) leads to noticeably different contours, especially for the $w_0$ parameter.
+We quantified these differences by examining the figure of merit, standard deviation, and mean of the contours, and found that, although not substantial, the VMIM produces better results in terms of contour width and alignment with the fiducial value(??).
+    \item  When using the VMIM to compress the original high-dimensional data, we compared the posterior obtained in the implicit inference framework with those obtained from Bayesian hierarchical modeling and power spectrum.
+We demonstrate that both map-based approaches lead to a significant improvement in constraining $\Omega_c, w_0, \sigma_8$ compared to the 2-point statistics.
+However, $h, n_s,\Omega_b$ are not constrained by either and are prior dominated.
+    \item   When using the VMIM to compress the original high-dimensional data,  in the limit of our setting, the two methods, i.e.
+the Bayesian hierarchical inference and the Likelihood free inference, leads to the same posterior distributions.
+\end{enumerate}
+
+It is important to discuss the limitations of the methodology used in this paper and highlight particular strategies for future extensions and applications.
+In this work, we employed a physical model based on a lognormal prior, notoriously faster than simulation-based methods.
+Although we have shown that this description is sufficiently complex for our tasks, as evidenced by the different contours obtained from the full-field and power spectrum methods, it is important to note that this is a good approximation for the convergence at intermediate scales, but can not be appropriate to analyze small scales.
+For example, the lognormal shift parameters are computed using the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which employs perturbation theory.
+However, as mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales.
+However, as the main objective of this paper was to compare the different inference strategies, we are not concerned with the potential implications of this approximation.
+In future applications, the simulation procedure will be based on N-body simulations.
+
+Additionally, we did not include any systemic in the current application, although previous studies demonstrated that the map-based approaches help to dramatically improve the constraints on systematic and cosmological parameters in the presence of these last.
+The main reason for this absence is mainly related to the difficulty of modeling systematic effects, like for example intrinsic alignment, in the lognormal description.
+
+Hence, the natural next step will be to use N-body simulations as the physical model for the  \texttt{SBILens} package.
+Specifically, we aim to start from the model presented in [reference], and further develop it by including additional systematics such as the redshift uncertainties, the baryonic feedback, and a more complicated intrinsic alignment model.
+With these improvements, our model will be better suited to handle real cosmic shear data, allowing us to fully maximize the information gained from next-generation surveys.
+
 %--------------------------------------------------------------------
@@
 to reweight summary distributions for posterior construction) so it's not super useful.
-I'll be sure to explicitly show the difference in Fluri et al and the IMNN in the corresponding section.}
+I'll be sure to explicitly show the difference in Fluri et al and the IMNN in the corresponding section.} \\
 \justine{Awesome, thanks! :)}
 \end{itemize}
+\denise{I wrote IMNN for \cite{fluri2021cosmological} because their optimization, as for the IMNN, is based on the CramerRao bound, so on the Fisher Information.
+The difference is that in the IMNN there is an assumption of Gaussian distribution, which allows you to write the Fisher matrix as:
+\begin{equation}
+      F_{\alpha \beta}= \sum_{ij}\frac{\partial \mu}{\partial \theta_{\alpha}} C_{ij}^{-1}
+      \frac{\partial \mu}{\partial \theta_{\beta}}
+      \label{Fisher_matrix}
+\end{equation}
+and then the determinant of the fisher matrix is optimized.
+In \cite{fluri2021cosmological} they minimize the determinant of the jacobian of the summaries and the determinant of the covariance. 
+Both methods come from the CramerRao bound, but as Lucas said, this is something that has to be explained in the section.
+I could not write all of this in the Table.
+We can think of an acronym (FIM maybe?
+Fisher Information Matrix ) and introduce it in the proper section.
+}
 

### 2023-09-08 paper/paper.tex
text 08/09/2023

--- 2023-07-11 paper/paper.tex
@@
    {In many cosmological applications, the likelihood function is either unknown or intractable, limiting the use of traditional inference methods.
-Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data to approximate the posterior (or likelihood) distribution instead of evaluating an explicit likelihood function.
+Likelihood-free inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data to approximate the posterior (or likelihood) distribution instead of evaluating an explicit likelihood function.
 However, these methods suffer from the curse of dimensionality, necessitating the development of compression schemes to reduce high-dimensional data into lower-dimensional summary statistics.
-
-   \justine{\textbf{I think the 2 main ideas for this part are: }
-
-    - classical way to overcome this issue: we compress the data into 2pt statistics  and use gaussian likelihood → pb: we miss a lot of information}
-    \denise{That's true. But adding it in the abstract would make the "Context" subsection (this division is mandatory for A$\&$A) too long. This is properly described in the introduction}
-
-    \justine{ - considering the previous pb: thanks to Implicit Inference methods we can consider any kind of summary stat (because we can learn the exact posterior p(theta | summary stat)), so to have the best contours (meaning that we exatrct all gaussian as well as non-gaussian signal) we only have to focus on how we compress our data.}
-    \denise{That's also true.
-But we don't use any summary statistics (except the power spectrum because it's easy), and I think that one of the points of the paper is to extract the information from the pixels, without using any summaries.
-So I personally would not talk about summary statistics }
-
-     \justine{\textbf{Note:}
-
-    - for me LFI = Implicit Inference, I think we should use only one
-   }
-   \denise{True again.
-But the definition "Likelihood-Free Inference" is more used in the community.
-Implicit Inference is something that we introduce in the motivation and introduction.
-I don't know exactly how much people are familiar with this concept.
-}
-    \justine{ - 'Likelihood-Free Inference (LFI) approaches offer a promising solution by using forward-simulated synthetic data instead of an explicit likelihood function.' → Implicit inference approaches offer a promising solution to approximate posterior distribution by using simulations from an implicit likelihood (a simulator) instead of evaluating explicit likelihood function ..?
-
-    - I'm probably wrong but not sure about the last sentence.
-}
    }   
   % aims heading (mandatory)
-    {We aim to derive the optimal neural-compression strategy for full field Likelihood free inference by density estimation.
-Moreover, by using the optimal compression strategy, we demonstrate that the posterior distribution is comparable to those derived from a Bayesian forward modeling approach.}
+    {The goal of this paper is to investigate the performance of several neural compression strategies for full field Likelihood free inference by density estimation.
+Moreover, we aim to demonstrate that, by using an optimal compression strategy, the posterior distribution is comparable to those derived from a Bayesian forward modeling approach.}
   % methods heading (mandatory)
@@
 We evaluate their performance by measuring their impact on the constraints on the $\Lambda$CDM parameters expected from LSST-Y10.
-    For both forward-modeling the convergence field and generating observed mock data, we use the \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}} package. 
-    \texttt{SbiLens} is a jax-based differentiable physical model specifically designed for Likelihood Free Inference and Bayesian inference algorithms that require access to likelihood derivatives.
+    For both forward-modeling the convergence field and generating observed mock data, we develop the \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}} package. 
+    \texttt{SBILens} is a Jax-based differentiable physical model used to generate convergence maps as lognormal random fields.
+It is specifically designed for likelihood-free inference and Bayesian inference algorithms that require access to likelihood derivatives.}
 
-    \justine{\textbf{I think the main ideas for this part are: }
-
-    - first we describe sbi\_lens : Provides a differentiable log normal mass map simulator specifically designed for for weak lensing inference.
-Then you can describe a bit the simulations like tomographic bins etc I guess
-
-    - We construct our ground truth
-
-    - We use implicit inference techniques to go from the compressed stat to the posteriors}
-    }
-  % results heading (mandatory)
-   {Our experiments validate the effectiveness of LFI methods,  as we demonstrate comparable posterior distributions between LFI and Bayesian forward modeling approaches \justine{field-based Bayesian Hierarchical Model analysis}.
-Additionally, our results provide insights into the design of data compression procedures to derive low-dimensional summary statistics while minimizing information loss.
+   {Our experiments validate the effectiveness of LFI methods, as we demonstrate comparable posterior distributions between LFI and Bayesian forward modeling approaches.
    All codes associated with this paper can be found at this link \github.}
-  % conclusions heading (optional), leave it empty if necessary 
    {}
@@
 To address the need for result validation in the LFI context, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
-The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SbiLens}}, a Python package designed for weak lensing implicit inference with differentiable simulator implemented in Jax.
-\texttt{SbiLens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
+The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Python package designed for weak lensing inference with differentiable simulator implemented in Jax.
+\texttt{SBILens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
 
 The paper is structured as follows: in \autoref{Sec:Motivation}, we illustrate the motivation behind this work.
-In \autoref{Sec:the SBILens framework}, we introduce the \texttt{SbiLens} framework and describe the simulated data used in this work.
+In \autoref{Sec:the SBILens framework}, we introduce the \texttt{SBILens} framework and describe the simulated data used in this work.
 In \autoref{Sec:experiment}, we detail the inference strategy and the three different approaches we used: the power spectrum, the map-based inference based on the Bayesian forward-modeling, and the map-based inference based on the LFI.
@@
 %--------------------------------------------------------------------
-With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, that may not fully capture the non-Gaussian information present in the lensing field at the scales accessible to future surveys.
-In this paper we introduce a forward model that directly extracts information from the raw pixel data, rather than relying on the analytical evaluation of summary statistics.
+With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, which may not fully capture the non-Gaussian information present in the lensing field at the scales accessible to future surveys.
+In this paper, we introduce a forward model that directly extracts information from the convergence maps, rather than relying on the analytical evaluation of summary statistics.
 By doing so, we aim to preserve all available information and facilitate the incorporation of systematic effects and the combination of multiple cosmological probes through joint simulations.
-In this context, the simulator of the observables serves as our physical model, where each component is tractable.
-These models, often refereed to as \textit{probabilistic program}, can be described as follows.
-They take as input a vector parameter $\bm{\theta}$, which describes the underlying deterministic model.
-Then, they sample internal states, dubbed \textit{latent variables}, from the distribution $\bm{z}_i \sim p_i(\bm{z}_i|\bm \theta, \bm{z}_{<i})$, than can be directly or indirectly related to a physically meaningful state of the system.
-Finally, they generate the output $\bm x$ from the distribution $\ p(\bm d|\bm \theta)$, where $\bm{d}$ represents the observations.
+In a forward model context, the simulator of the observables serves as our physical model, where each component is tractable.
+These models, often referred to as \textit{probabilistic program}, as illustrated by \citet{cranmer2020frontier}, can be described as follows: The models take as input a vector parameter $\bm{\theta}$, which describes the underlying deterministic model.
+Then, they sample internal states $\bm{z}$, dubbed \textit{latent variables}, from the distribution $p(\bm{z}|\bm \theta)$.
+These states can be directly or indirectly related to a physically meaningful state of the system.
+Finally, the models generate the output $\bm x$ from the distribution $p(\bm x|\bm \theta,\bm{z})$, where $\bm{x}$ represents the observations.
 
@@
 \end{equation}
-as it corresponds to an integral over all possible trajectories throughout the latent space.
+since it involves integrating over all potential paths through the latent space.
 To overcome this limitation while still capturing the full information content of the data, two different approaches have been proposed in the literature.
-Although these approaches are often referred to by different names, in this work, we will make the following distinction:
+Although these approaches are often referred to by different names, hereinafter we will make the following distinction:
 %--------------------------------------------------------------------
@@
 \end{equation}
-In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Models (BHMs).
+In the context of the forward model and full-field analysis, Bayesian Hierarchical Models (BHMs) fall into this category.
 A Bayesian forward approach involves using a given physical model to predict observations and then comparing these predictions with observations to infer the parameters of the model.
@@
 %--------------------------------------------------------------------
-\paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations.
+\paragraph{\textbf{Implicit inference}} referring to all the approaches that try to infer the marginal likelihood from simulations.
 This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution:
@@
 In this work, our focus will be on DELFI methods, which approach the inference task as a density estimation problem. \\
-We can divide Likelihood-Free Inference approaches into two separate steps:
+In particular, the standard deep learning approach for Likelihood-Free Inference can be described as separated into two distinct steps:
 \begin{enumerate}
@@
 \end{enumerate}
-In the first step, we introduce a parametric function $f_{\varphi}$ such that:
+In the first step, we introduce a parametric function $F_{\varphi}$ such that:
      \begin{equation}
-         \bm{y}=f_{\varphi}(\bm{\theta}),
+         \bm{t}=F_{\varphi}(\bm{\theta}),
      \end{equation}
-which aims to reduce the dimensionality of the data while preserving information. Typically, the compressed statistics $\bm{y}$ is assumed to have the same dimension of $\bm{\theta}$. \\
-In the second step, Neural Density Estimation can target either building an estimate $p_{\phi}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (referred to as the Neural Likelihood Estimation (NLE) task), or targeting the posterior distribution $p(\bm{\theta}|\bm{x})$, (known as Neural Posterior Estimation (NPE) task).
+which aims to reduce the dimensionality of the data while preserving information. Typically, the compressed statistics $\bm{t}$ is assumed to have the same dimension of $\bm{\theta}$. \\
+In the second step, Neural Density Estimation can target either building an estimate $p_{\varphi}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (referred to as the Neural Likelihood Estimation (NLE) task), or targeting the posterior distribution $p(\bm{\theta}|\bm{x})$, (known as Neural Posterior Estimation (NPE) task).
 
-It is important to note that not all neural compression techniques are equivalent.
-Many papers that apply neural techniques for extracting summary statistics have used sub-optimal compression techniques, such as Mean Square Error or Maximum Absolute Error.
-However, it is also important to consider that in this case, the compression of summaries will be sub-optimal, resulting in a loss of information (larger contours), but it will not introduce bias in the final constraints.
-On the contrary, many other papers rely on assuming proxy Gaussian likelihoods and estimate the mean and covariance of these likelihoods from simulations.
-While likelihood-free inference using MSE and MAE summaries is robust in the sense that the approximations made in the compression steps will not introduce bias to the parameter inferences, the same cannot be said for summaries obtained from Gaussian likelihood approximations.
-We summarize the different neural compression strategies found in the literature in \autoref{tab:biblio_survey}.
+The main motivation for this work is to evaluate the impact of a given compression strategy on the final posterior distribution and determine whether an optimal compression strategy exists.
+Indeed, it is important to consider that different neural compression techniques may not be equivalent.
+We summarize the various neural compression strategies found in the literature in \autoref{tab:biblio_survey}.
+Many of these papers have used neural compression techniques that rely on optimizing the Mean Square Error or the Maximum Absolute Error. \\
+As we will demonstrate in the following sections, this corresponds to training the model to estimate the mean and the median of the posterior distribution.
+Other papers rely on assuming proxy Gaussian likelihoods and estimate the mean and covariance of these likelihoods from simulations.
+Such compression of summaries could be sub-optimal in certain applications, resulting in a loss of information and hence larger contours.
+
+Therefore, our primary focus will be to investigate the effectiveness and validity of these compression techniques in the context of weak lensing applications (within the limits of our setting).
 
 \vspace{1cm}
-Ultimately, for a given simulation model, if an optimal compression statistic is used, the two approaches should converge to the same posterior. Therefore, the goals of this paper will be:
+Ultimately, we should keep in mind that, for a given simulation model, if an optimal compression statistic is used, the two approaches should converge to the same posterior.
+Therefore, the goals of this work will be:
 \begin{enumerate}
-    \item Find the optimal compression strategy.
-    \item Demonstrate that by using this optimal compression strategy, both the implicit and explicit methods yield comparable results.
+    \item Find a near-optimal compression strategy.
+    \item Demonstrate that by using this near-optimal compression strategy, both the implicit and explicit methods yield comparable results.
 \end{enumerate}
@@
 \section{The SBILens framework}\label{Sec:the SBILens framework}
-\denise{PEOPLE CAN READ THIS!!}
+To investigate the questions above, we have developed the Python package \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\texttt{SBILens}}, a weak lensing differentiable simulator based on a log-normal model.
+\texttt{SBILens} enables the sampling of convergence maps in a tomographic setting while considering the cross-correlation between different redshift bins.
 \subsection{Lognormal Modeling}\label{Sec:Lognormal Modeling}
@@
 where $\lambda$ is a free parameter that determines the shift of the lognormal distribution.
-The convergence $\kappa$ in a given redshift bin is fully determined by the shift parameter $\lambda$, the mean $\mu$ of the associated Gaussian field $\kappa_g$, and its variance $\sigma_{g}^2\equiv \xi_g$.
+The convergence $\kappa$ in a given redshift bin is fully determined by the shift parameter $\lambda$, the mean $\mu$ of the associated Gaussian field $\kappa_g$, and its variance $\sigma_{g}^2$.
 The correlation of the lognormal field, denoted as $\xi_{ln}$, is also a function of these variables and is related to $\xi^{ij}_g$ through the following equations:
@@
 
- In the SbiLens framework, the sampling of the convergence maps can be described as follows. 
+ In the SBILens framework, the sampling of the convergence maps can be described as follows. 
 First, we define the survey in terms of galaxy number density, redshifts, and shape noise.
@@
 \begin{equation}
-     \bm{\kappa_g}=\bm{Z}*\tilde{\bm{\Sigma} }
+     \bm{\kappa_g}=\hat{\bm{Z}}\cdot\tilde{\bm{\Sigma} }
 \end{equation}
-where $\bm{Z}$ represents the latent variables of the simulator.
+where $\hat{\bm{Z}}$ represents the Fourier transform of the latent variables of the simulator.
 Finally, we transform the Gaussian map $\kappa_g$ into a LogNormal field using \autoref{Eq:log_norm_kappa}.
+ \\
+To ensure that we recover the correct auto- and cross-power spectra, we compare the results from our simulations to theoretical predictions for different tomographic bin combinations.
+We show the results in \autoref{fig:psconvergence_maps}.
+\denise{Not sure we want this}
 %------------------------------------------------------------------
@@
 The priors used in the simulations and in the inference process are listed in \autoref{tab:prior}, following \citet{zhang2022transitioning}.
-To simulate our data, we develop the SbiLens package, which employs a Lognormal model to represent the convergence maps, as explained in the previous section.
+To simulate our data, we develop the SBILens package, which employs a Lognormal model to represent the convergence maps, as explained in the previous section.
 Specifically, the package uses the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023} to compute the theoretical power- and cross-spectra.
@@
 Following \citet{boruah2022map}, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$, where $\Delta L$ represents the pixel resolution.
-The shift parameters at each redshift are computed using the fiducial cosmology values for $\Omega_b$, $h_0$, and $n_s$.
-To incorporate the cosmology dependence of $\lambda$ from $\Omega_c$, $\sigma_8$ and $w_0$, we calculate the shift for different points in the cosmological parameter space and then interpolate the shift values for other points in the parameter space.
-
-As mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales.
-However, the main objective of this paper is to compare different inference strategies and we are not concerned with the potential implications of this approximation.
-In future applications of the presented inference pipeline, the simulation procedure will be based on N-body simulations.
-
+\\
+For each redshift bin, we tested the dependency of the shift parameter $\lambda$ on various cosmological parameters.
+Specifically, we investigated how the value of $\lambda$ changed when varying a specific cosmological parameter while keeping the others fixed.
+Our findings revealed that the parameters $\Omega_b$, $h_0$, and $n_s$ had almost no significant impact on $\lambda$.
+As a result, we computed the shift parameters for each redshift using the fiducial cosmology values of $\Omega_b$, $h_0$, and $n_s$.
+To account for the cosmology dependence of $\lambda$ on $\Omega_c$, $\sigma_8$, and $w_0$, we calculated the shift for various points in the cosmological parameter space and then interpolated the shift values for other points in the parameter space.
 Each map is reproduced on a regular grid with dimensions of $256 \times 256$ pixels and covers an area of $10\times 10$ deg$^2$.
-
+An example of a tomographic convergence map simulated using the \texttt{SBILens} package is shown in \autoref{fig:convergence_maps}.
+\denise{TODO: change disp images}
+%--------------------------------------------------------------------
+% ############# PRIOR COSMO TABLE  #############
+%--------------------------------------------------------------------
 \begin{table}
-	\centering
-	\caption{ Prior and fiducial values used for the analyses. 
- The symbol $\mathcal{N}_T$ represents a Truncated Normal distribution.
-The lower bound of the support for the $\Omega_c$ distribution is set to zero, while the lower and upper bounds for the $w_0$ distribution are set to -2.0 and -0.3, respectively.}
+	\begin{center}
     	\begin{tabular}{lcc} 
@@
 Both the shape noise and galaxy number density values are obtained from SRD.
-The pixel area is given by $A_{pix}\approx$.
+The pixel area is given by $A_{pix}\approx 5.49$ arcmin$^2$.
 \autoref{fig:redshift_distribution} illustrates the resulting source redshift distribution, and \autoref{tab:survey_spec} provides a summary of the survey specifications.
+%--------------------------------------------------------------------
+% ############# SURVEY SPECIFICATION TABLE  #############
+%--------------------------------------------------------------------
 \begin{table}
-	\centering
-	\caption{ LSST Y10 source galaxy specifications in our analysis. All values are based on the LSST DESC SRD.}
+	\begin{center}
     	\begin{tabular}{lc} 
@@
 %--------------------------------------------------------------------
+% ############# PLOT CONVERGENCE MAPS  #############
+%--------------------------------------------------------------------
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{figures/bin1and2.pdf}
+    \includegraphics[width=\textwidth]{figures/bin3.pdf}
+    \includegraphics[width=0.5\textwidth]{figures/bin5.pdf}
+    \caption{
+     Example of convergence map simulated using the \texttt{SBILens} package.
+    }
+     \label{fig:convergence_maps}
+     \end{center}
+\end{figure}
+%--------------------------------------------------------------------
+% ############# PLOT CONVERGENCE MAPS  #############
+%--------------------------------------------------------------------
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/psconvergencemaps.pdf}
+    \caption{
+    Convergence power spectra for different tomographic bin combinations.
+The solid yellow line shows the measurement from 20 simulated maps using the survey setting described in \autoref{Sec:the SBILens framework}, while the black dashed line shows the theoretical predictions computed using jax-cosmo.
+In this figure, the shaded regions represent the standard deviation from 20 independent map realizations.
+    }
+     \label{fig:psconvergence_maps}
+\end{figure*}
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
@@
 \section{Experiment}\label{Sec:experiment}
+In the following section, we illustrate the inference strategy for the two approaches under investigation: the Bayesian forward modeling and the map-based inference based on the LFI.
+Additionally, we conduct a power spectrum study.
+Indeed, as discussed in \autoref{Sec:Lognormal Modeling}, log-normal fields offer the advantage of rapidly generating convergent fields while accounting for non-Gaussianities.
+To emphasize this claim further, along with the full-field analysis, we include a power spectrum analysis.
+This analysis demonstrates that there is indeed a gain of information when adopting a full-field approach.
+\\
+Finally, we provide a detailed overview of various neural compression strategies that differ in terms of the loss functions used to train the neural network.
 %--------------------------------------------------------------------
@@
 \subsection{Explicit Inference}
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-\subsubsection{Power spectrum}
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-To obtain the probability distribution of the cosmological parameters using a summary-statistics-methods, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
-We assume a Gaussian likelihood with a cosmology-independent covariance matrix:
-\begin{equation}
-    \log{\mathcal{L}}(\bm{\theta})=-\frac{1}{2}[\bm{d}-\bm{\mu}(\bm{\theta})]^{T}\bm{C}^{-1}[\bm{d}-\bm{\mu}(\bm{\theta})].
-\end{equation}
-To compute the expected theoretical predictions $\bm{\mu}(\bm{\theta})$ we use the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}. 
- The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, presented in \autoref{tab:prior}, using the same theoretical library.
-Specifically, in {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
-\begin{equation}
-    \text{Cov}(C_{\ell},C_{\ell}')=\frac{1}{f_{sky}(2 \ell+1)}\left(C_{\ell}+\frac{\sigma_{\epsilon}^2}{2n_s}\right)
-\end{equation}
-where $f_{sky}$ is the fraction of sky observed by the survey, and $n_s$ is the number density of galaxies. 
-To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map (our fiducial).
 %--------------------------------------------------------------------
@@
 To construct the explicit map-based inference strategy in the Hierarchical Bayesian framework, we built a likelihood based on the data model described in \autoref{Sec:Lognormal Modeling}. 
-As mentioned before, this means that the simulator serves as the physical deterministic model capable of generating the non-linear representation of the convergence map.
+As mentioned before, this means that the simulator serves as the physical model capable of generating the non-linear representation of the convergence map.
 
-However, in practical terms, the measurement of convergence for each pixel and bins will differ from real observations due to noise.
+However, in practical terms, the measurement of convergence for each pixel and bin will differ from real observations due to noise.
 This is taken into consideration in the likelihood.
 Specifically, for LSST Y10, the number of galaxies for each pixel should be sufficiently high so that, according to the central limit theorem, we can assume the observation is characterized by Gaussian noise, with $\sigma_n^2=\sigma_e^2/N_s$, where $N_s$ represents the total number of source galaxies per bin and pixel.
-Given $\sigma_n^2$ the variance of this Gaussian likelihood, its log-form can be expressed as:
+Given $\sigma_n^2$ the variance of this Gaussian likelihood, its negative log-form can be expressed as:
 \begin{equation}
@@
     \sum_i^{N_{pix}} \sum_{j}^{N_{bins}} \log{P(\kappa^{obs}_{i,j}|\kappa_{i,j},\bm{\theta})}
-    =\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2},
+    =-\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2},
 \end{equation}
@@
 To sample the posterior distribution for $\bm{\theta}$, we use a Hamiltonian Monte Carlo (HMC) algorithm.
+Specifically, we employ the \texttt{NUTS} algorithm, an adaptive variant of HMC implemented in \href{https://github.com/pyro-ppl/numpyro}{\texttt{NumPyro}} \citep{phan2019composable, bingham2019pyro}, which uses the No U-Turn Sampler (NUTS, \citet{hoffman2014no}).
+\\
 The HMC algorithm is particularly helpful in high-dimensional spaces where a large number of steps are required to effectively explore the space.
@@
 As the code is implemented in Jax, the gradients are accessible via automatic differentiation.
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+\subsubsection{Power spectrum}
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+To obtain the probability distribution of the cosmological parameters using a summary-statistics method, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
+We assume a Gaussian likelihood with a cosmology-independent covariance matrix:
+\begin{equation}
+    \mathcal{L}(\bm{\theta})=-\frac{1}{2}[\bm{d}-\bm{\mu}(\bm{\theta})]^{T}\bm{C}^{-1}[\bm{d}-\bm{\mu}(\bm{\theta})].
+\end{equation}
+To compute the expected theoretical predictions $\bm{\mu}(\bm{\theta})$ we use the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023}.
+ The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, presented in \autoref{tab:prior}, using the same theoretical library.
+Specifically, in {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
+\begin{equation}
+    \text{Cov}(C_{\ell},C_{\ell'})=\frac{1}{f_{sky}(2 \ell+1)}\left(C_{\ell}+\frac{\sigma_{\epsilon}^2}{2n_s}\right)
+\end{equation}
+where $f_{sky}$ is the fraction of sky observed by the survey, and $n_s$ is the number density of galaxies. 
+To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map (our fiducial).
+To constrain the cosmological parameters, we sample from the posterior distribution using an HMC algorithm with adaptive step size, implemented in the TensorFlow Probability MCMC python package.
 %--------------------------------------------------------------------
@@
 \subsubsection{Benchmark compression scheme}
-\denise{PEOPLE CAN READ THIS!!}
 To ensure the scalability of density estimation LFI in cases where forward simulations are computationally expensive, it becomes necessary to employ compression techniques that reduce the dimensionality of the data space and extract summary statistics.
-Specifically, we try to find a function $\bm{t}=F(\bm{d})$, where $\bm{t}$ represents low-dimensional summaries of the original data vector $\bm{d}$.
-The objective is to achieve a compression function $F(\bm{d})$ that maximizes the information while minimizing dimensionality.
-Previous studies \citep{alsing2018generalized} have demonstrated that, under specific conditions, a compression scheme can be achieved where the dimension of the summaries dim($\bm{t}$) is equal to the dimension of the unknown parameters dim($\bm{\theta}$) without any loss of information at the Fisher level.
+Specifically, we try to find a function $\bm{t}=F(\bm{x})$, where $\bm{t}$ represents low-dimensional summaries of the original data vector $\bm{x}$.
+The objective is to achieve a compression function $F(\bm{x})$ that maximizes the information while minimizing dimensionality.
+Previous studies \citep{heavens2000massive} have demonstrated that a compression scheme can be achieved where the dimension of the summaries dim($\bm{t}$) is equal to the dimension of the unknown parameters dim($\bm{\theta}$) without any loss of information at the Fisher level.
 Multiple approaches exist in an attempt to satisfy this condition.
@@
 However, it is important to note that this approach does not guarantee the recovery of maximally informative summary statistics.
-While this may hold true for fixed fiducial values of the cosmological parameters and assumptions of Gaussianity, its validation to more generic cases need to be proved.
 Indeed, minimizing the $L_{2}$ norm is equivalent to training the model to estimate the mean of the posterior distribution.
@@
 In \autoref{Sec:appendix_Maximum Absolute Error}, we demonstrate that minimizing this loss function is equivalent to training the model to estimate the median of the posterior distribution.
-While extensively employed in various previous studies \citep{2018PhRvD..97j3515G, fluri2018cosmological, ribli2019weak}, similar to the previous consideration for the MSE, it is important to note that the effectiveness of this loss function in extracting sufficient statistics needs to be demonstrated in more generic applications.
+While extensively employed in various previous studies \citep{2018PhRvD..97j3515G, fluri2018cosmological, ribli2019weak}, similar to the previous consideration for the MSE, it is important to note that the effectiveness of this loss function in extracting sufficient statistics needs to be demonstrated.
 %--------------------------------------------------------------------
@@
 \begin{equation}\label{Eq:Loss_vmim}
-    \mathcal{L}=- \log{q(\bm{\theta}|\bm{t};\bm{\varphi'})}.
+    \mathcal{L}=- \log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {x}) ; \bm{\varphi}')}.
 \end{equation}
-Here, $q(\bm{\theta}|\bm{t};\bm{\varphi'})$ represents a variational conditional distribution, $\bm{t}$ denotes the summary statistics, $\bm{\theta}$ corresponds to the data vector of the cosmological parameters, and $\bm{\varphi'}$ are the parameters of the Neural Network.
+Here, $q(\bm {\theta} |F_{\bm {\varphi}}(\bm {x});\bm{\varphi'})$ represents a variational conditional distribution, where $\bm{\theta}$ corresponds to the data vector of the cosmological parameters, and $\bm{\varphi'}$ to the parameters characterizing the variational conditional distribution itself.
+$F_{\bm {\varphi}}$ denotes the compression network of parameter $\bm {\varphi}$, used to extract the summary statistics $\bm{t}$ from the original high-dimensional data vector $\bm{x}$, such that  $\bm{t}=F_{\bm {\varphi}}(\bm{x})$.
 In order to understand the significance of this loss function, it is necessary to start by considering the mathematical definition of mutual information $I(\bm{t}, \bm {\theta})$:
@@
     I(\bm{t}, \bm {\theta}) &= D_{KL}(p(\bm {t}, \bm {\theta})||p(\bm {t})p(\bm {\theta})) \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
-        &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta}  d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
-    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
+    &= \int d\bm{\theta} d\bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
+    &= \int d\bm{\theta} d\bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
+        &= \int d\bm{\theta} d\bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d\bm{\theta}  d\bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
+    &= \int d\bm{\theta} d\bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d\bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
     &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- \mathbb{E}_{p(\bm {\theta})} [\log{p(\bm {\theta})}] \\ \nonumber
@@
 \end{align}
-
 in the above equation, $D_{KL}$ is the Kullback-Leibler divergence \citep{kullback1951information}, $p(\bm {t}, \bm {\theta})$ is the joint probability distribution of summary statistics and cosmological parameters, and $H(\bm {\theta})$ represents the \textit{entropy} of the distribution of cosmological parameters.
 Essentially, mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
-The summary statistics $\bm{t}$ are extracted from the original high-dimensional data vector $\bm{d}$ using $\bm{t}=F_{\bm {\varphi}}(\bm{d})$.
-The goal is to find the parameters $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
+The goal is to find the parameters of the network $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
 \begin{equation}
-   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{d}), \bm {\theta}).
+   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{x}), \bm {\theta}).
 \end{equation}
 However, the mutual information expressed in \autoref{Eq:mutual_information} is not tractable.
-To overcome this limitation, various approaches have been developed that rely on tractable bounds, enabling the training of deep neural networks to optimize the mutual information.
+To overcome this limitation, various approaches have been developed that rely on tractable bounds, enabling the training of deep neural networks to optimize mutual information.
 In this study, we adopt the same strategy used by \citet{jeffrey2021likelihood}, which involves using the variational lower bound \citep{barber2003information}:
@@
 \begin{equation}
-    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {d}) ; \bm{\varphi}')}], 
+    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {x}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {x}) ; \bm{\varphi}')}], 
 \end{equation}
@@
 %--------------------------------------------------------------------
-\FL{Recognizing that the uncertainty on different cosmological parameters will vary, a third
+Recognizing that the uncertainty on different cosmological parameters will vary, a third
 class of inverse variance weighted MSE was proposed in \cite{fluri2018cosmological} with the idea of
@@
 \begin{equation}
-    \mathcal{L} = \frac{1}{2} \log(|\Sigma|) + \frac{1}{2}(t - \theta)^t \Sigma^{-1} (t - \theta) 
+    \mathcal{L} = \frac{1}{2} \log(|\bm{\Sigma}|) + \frac{1}{2}(\bm{t} - \bm{\theta})^{\top} \Sigma^{-1} (\bm{t} - \bm{\theta}) 
 \end{equation}
-where $t$ is the summary statistics and $\Sigma$ is the covariance matrix of representing the uncertainty on the cosmological parameters $\theta$.
-Both $t$ and $\Sigma$ can be outputs of the compression network, i.e.
-$f_\theta(x)=(t, \Sigma)$.
+where $\bm{t}$ is the summary statistics and $\bm{\Sigma}$ is the covariance matrix representing the uncertainty on the cosmological parameters $\bm{\theta}$.
+Both $\bm{t}$ and $\bm{\Sigma}$ can be outputs of the compression network, i.e.
+$F_{\bm{\varphi}}(\bm{x})=(\bm{t}, \bm{\Sigma})$.
 
-One recognizes here the expression of a Gaussian probability function, and this expression can thus be straightforwardly related to the VMIM case by simply assuming a Gaussian distribution as the variational approximation for the posterior $q(\theta | x) = \mathcal{N}(\theta; t, \Sigma)$.
+One recognizes here the expression of a Gaussian probability function, and this expression can thus be straightforwardly related to the VMIM case by simply assuming a Gaussian distribution as the variational approximation for the posterior $q(\bm{\theta} | \bm{x}) = \mathcal{N}(\bm{\theta}; \bm{t}, \bm{\Sigma})$.
 
-This means that this loss function is actually sub-optimal for two reasons:
+This means that for this loss function:
 \begin{itemize}
-    \item The summary extracted by the neural network $t$ is, similarly to the MSE case only an estimate of the mean of 
+    \item The summary $\bm{t}$ extracted by the neural network is, similarly to the MSE case, only an estimate of the mean of 
     the posterior distribution, which is not guaranteed to be sufficient.
@@
 \end{itemize}
- A simple MSE would thus seem to be preferable.
-}
 %--------------------------------------------------------------------
@@
 {Information Maximising Neural Networks (IMNNs)}}. 
-\LM{TODO: condensed introduction; emphasize learning compression at single point; network learns score function which generalises smoothly away from fiducial \cite{Makinen_2021, Makinen_2022, Charnock_2018}, Wandelt (in prep.)
-}
+A different approach has been proposed by \cite{charnock2018automatic} and further explored in \citet{Makinen_2021, Makinen_2022}.
+They implemented the Information Maximizing Neural Networks (IMNN), a neural network trained on forward simulations designed to learn optimal compression schemes, even in circumstances where the likelihood function is intractable or unknown.
+Specifically, they proposed a new scheme to find optimal non-linear data summaries by using the Fisher information to train a neural network. 
+Inspired by the MOPED algorithm \citep{heavens2000massive}, they aimed to find a transformation $f$ that maps the data to compressed summaries: $f: \textbf{x} \to \textbf{t}$ while conserving the Fisher information.
+The loss function takes the following form:
+\begin{equation}\label{eq:loss_imnn}
+    \mathcal{L} = -  |\text{det}(\textbf{F})|+ r_{\Sigma},
+\end{equation}
+where $\textbf{F}$ is the Fisher matrix, and $r_{\Sigma}$ is a regularization term typically dependent on the covariance matrix,  introduced to control the magnitude of the summaries.
+Since computing the Fisher matrix requires a large number of simulations, they proceed as follows: a large number of simulations with the same fiducial cosmology but different initial random conditions are fed forwards through identical networks.
+The summaries from these simulations are combined to compute the covariance matrix.
+Additionally, the summaries from simulations created with different fixed cosmologies are used to calculate the derivative of the mean of the summary with respect to the parameter.\footnote{The method of finite differences is necessary when the framework in which the code is implemented does not support automatic differentiation.} Finally, the covariance and the mean derivatives are combined to obtain the Fisher matrix.
 
+An application of this method in the context of weak lensing can be traced back to the work of \citet{fluri2021cosmological, fluri2022full}.
+Following \citet{charnock2018automatic}, they presented a compression approach that relies on the optimization of the Cramér-Rao bound.
+However, their approach does not assume that the summary follows a Gaussian likelihood.
+The loss function takes the form:
+\begin{equation}\label{eq:loss_gfim}
+\mathcal{L}=\log{\text{det}(\bm{\Sigma}_{\bm{\theta}}(\bm{t})})-2\log{\left|\text{det}\left(\frac{\partial \Psi_{\bm{\theta}}(\bm{t})}{\partial \bm{\theta}}\right)\right|},
+\end{equation}
+where $\bm{\Sigma}$ is the covariance matrix, and $\Psi_{\bm{\theta}}(\bm{t})=\mathbb{E}_{p(\bm {x}|\bm {\theta})}[\bm{t}]$.
+This loss function is equivalent to the log-determinant of the inverse Fisher matrix used in \autoref{eq:loss_imnn}. \\
+% \LM{TODO: condensed introduction; emphasize learning compression at single point; network learns score function which generalises smoothly away from fiducial \cite{Makinen_2021, Makinen_2022, Charnock_2018}, Wandelt (in prep.)
+% }
 %--------------------------------------------------------------------
+\subsection{Compression strategy}
+%--------------------------------------------------------------------
+The different compression strategies differ in the loss function employed during the training phase.
+For all of them, we use the same convolutional neural network architecture to build the compressor, specifically the ResNet-18, a deep ResNet architecture \citep{he2016deep}.
+The ResNet-18 is implemented using Haiku \citep{haiku2020github}, a Python deep learning library built on top of Jax.
 
+While the different compression strategies share the same architecture, the training strategies for VMIM and GNLL differ significantly.
+To train the Neural compressor under VMIM and GNLL, we jointly optimize the weights $\bm{\varphi}$ of the Neural Network $F_{\bm{\varphi}}$ (the compressor) and the parameters $\bm{\varphi}'$ of the variational distribution $q_{\bm{\varphi}'}$.
+For VMIM, the variational distribution $q(\bm{\theta}|\bm{t};\bm{\varphi}')$ is modeled using a Normalizing Flow, while for GNLL, a Conditional Multivariate Gaussian estimator is employed.
+The Normalizing Flow used for the inference strategy is also adopted for the variational distribution required in the compression procedure.
+For the Multivariate Gaussian estimator, we implement a three-layer neural network.
+The first two layers consist of fully connected layers with a Leaky Rectified Linear Unit activation function and a hyperbolic tangent activation function, respectively.
+The number of channels is increased to 256 and 128 for the first two layers, respectively.
+The last layer is a fully connected layer without an activation function, responsible for mapping the output of the first two layers to the desired output size (mean and covariance of the Gaussian distribution).
+\\
+After training, we export the results of the neural compression $F_{\bm{\varphi}}$ but discard the results from the density estimator.
+Indeed, as mentioned before, we perform the Likelihood-Free Inference (LFI) as a two-step procedure.
+This choice is motivated by the fact that it is difficult to train a very precise conditional density estimator when the compressor can still change from iteration to iteration.
+Hence, the choice to split the problem into two steps: first, the dimensionality reduction, where the density estimation part does not need to be perfect; second, the density estimation itself, which needs to be done very carefully now, but it is much easier because we are in low dimension.
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\subsubsection{Inference strategy}
+%--------------------------------------------------------------------
+\subsubsection{Inference strategy}\label{Sec:Inference_strategy}
 %--------------------------------------------------------------------
 %--------------------------------------------------------------------
-By comparing the posteriors $p(\theta| t)$ obtained with different compression strategies, we can assess the sensitivity of the results to the choice of statistics.
-In particular, the more informative the summary statistics, the “closer” the posterior $p(\theta | t)$ is to the true posterior $p(\theta | x)$.
-In this section, we describe how we approximate the posterior distribution for each summary statistic t, enable us to use it as a fair benchmark for comparison (we describe the performance metric in section ??).\\
-Since the likelihood $p(t|\theta)$ is intractable, we propose to do Neural Posterior Estimation \citep{npe1, npe2, npe3}, a Simulation-Based Inference (SBI) strategy that aims to directly approximate, from simulations, the posterior distribution
+By comparing the posteriors $p(\bm{\theta}| \bm{t})$ obtained with different compression strategies, we can assess the sensitivity of the results to the choice of the compression method.
+In particular, the more informative the summary statistics, the closer the posterior $p(\bm{\theta} | \bm{t})$ is to the true posterior $p(\bm{\theta} | \bm{x})$.
+\\
+As we have mentioned before, in the LFI context, the simulator is a black box that defines the likelihood distribution $p(\bm{x}|\bm{\theta})$ as an implicit distribution.
+We can sample from it, but we cannot directly evaluate it.
+The idea of Neural density Estimation is to introduce a parametric distribution model $\mathcal{P}_{\bm{\varphi}'}$ to approximate the implicit distribution $\mathcal{P}$.
+\\
+In particular, we focus on Neural Posterior Estimation \citep{npe1, npe2, npe3}, i.e., we aim to directly approximate the posterior distribution.
+Specifically, we use a conditional Normalizing Flows \citep{nf1, nf2} to model the parametric conditional distribution $q_{\bm{\varphi}'} (\bm{\theta} | \bm{t})$, and we optimize the parameters ${\bm{\varphi}'}$ according to the following negative log-likelihood (NLL) loss function:
+\begin{equation}\label{eq:nll}
+    \mathcal{L}=- \log{q_{\bm{\varphi}'}(\bm{\theta}|\bm{t})}.
+\end{equation}
+In the limit of a large number of samples and sufficient flexibility, we obtain:
+\begin{equation}
+    q_{\bm{\varphi}'^\ast}(\bm{\theta} | \bm{t}) \approx p(\bm{\theta} | \bm{t})
+\end{equation}
+where we indicate with $\bm{\varphi}'^\ast$ the values of $\bm{\varphi}'$ minimizing \autoref{eq:nll}. \\
+Finally, the target posterior $p(\bm{\theta} | \bm{t} = \bm{t_0})$ is approximated by $q_{\bm{\varphi}'^\ast}(\bm{\theta} | \bm{t} = \bm{t_0})$, where $\bm{t_0} = F_{\bm{\varphi}}(\bm{x_0})$, i.e., the compressed statistics from the fiducial convergence map $\bm{x_0}$ for a given compression strategy.
 
-\begin{equation}
-    p(\theta | t) \propto \int p(t | \theta, z) p(z|\theta) p(\theta) dz\,.
-\end{equation}
-
-Usually, NPE algorithm approximates the posterior distribution through the use of conditional Normalizing Flows $p_{\varphi} (\theta | t)$ \citep{nf1, nf2}.
-This is achieved by minimizing the negative log probability:
-
-\begin{equation}
-    \mathbb{E}_{p(\theta, t)} \left[ - \log p_{\varphi} (\theta | t)  \right].
-    \label{eq:nll}
-\end{equation}
-
-Finally, the target posterior $p(\theta | t = t_0)$ is approximated by $p_{\varphi}(\theta | t = t_0)$ where $t_0 = t(x_0)$ with $x_0$ the observed convergence map.
-
-For each summary statistic, we approximate it posterior distribution using the same NF architecture, namely a RealNVP \citep{realnvp} with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ \citep{silu}.
-And, for each summary statistic, the prior $p(\theta)$ used in the inference process remains the same.
+For each summary statistic, we approximate the posterior distribution using the same NF architecture, namely a RealNVP \citep{realnvp} with 4 coupling layers.
+The shift and the scale parameters are learned using a neural network with 2 layers of 128 neurons with Sigmoid-weighted Linear Units (SiLU) activation functions \citep{silu}.
 %--------------------------------------------------------------------
 %--------------------------------------------------------------------
-\subsection{Result estimators}
-%--------------------------------------------------------------------
-%--------------------------------------------------------------------
-To evaluate the results obtained from different approaches, we use two commonly employed estimators found in the literature. Their definitions are provided in the following sections.
-%--------------------------------------------------------------------
-\subsubsection{Figure of Merit (FoM)}
-%--------------------------------------------------------------------
-To quantify the constraining power of each inference strategy, we first employ the Figure of Merit (FoM), defined as follows:
-\begin{equation}\label{Eq:Figure_of_merit}
-    \text{FoM}_{\alpha \beta}=\sqrt{\text{det}(\Tilde{F}_{\alpha \beta})}.
-\end{equation}
-Here, $\alpha$ and $\beta$ represent a pair of cosmological parameters, and $\Tilde{F}_{\alpha \beta}$ refers to the marginalized Fisher matrix.
-We calculate $\Tilde{F}_{\alpha \beta}$ as the inverse of the parameter space covariance matrix $C_{\alpha \beta}$, which is estimated from the HMC chains and from the Normalizing flow sampling.
-Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalised parameter space of the $\alpha$ and $\beta$ pair.
-%--------------------------------------------------------------------
-\subsubsection{Kullback-Leibler divergence}
-%--------------------------------------------------------------------
-To measure the difference between the probability distribution obtained with different compressed summaries we compute the \textit{Kullback-Leibler divergence} or KL-divergence.
-As mentioned when we introduced the VMIM, the KL divergence is related to the relative entropy, and is a non-symmetric measure of the difference between two distributions.
-\denise{INCOMPLETE}
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-
+\begin{table*}
     \begin{center}
-\begin{table*}
-\begin{tabular}{ |p{3.5cm}|p{2cm}|p{3cm}|p{2.5cm}|p{4cm}|  }
+        \begin{tabular}{ |p{5cm}|p{3cm}|p{5cm}| }
             \hline
-Reference & \makecell{Architecture\\ compressor}  & \makecell{Loss function} & \makecell{Inference \\ strategy} & \makecell{Output compressor}   \\
+            \makecell{\textbf{Reference}} &  \makecell{\textbf{Loss function}} &  \makecell{\textbf{Inference strategy}}  \\
             \hline
-            \citet{2018PhRvD..97j3515G} & \makecell{CNN} & \makecell{MAE} &  \makecell{Likelihood \\ analysis} &\makecell{$\sigma_8, \Omega_m$}  \\
+            \citet{2018PhRvD..97j3515G} & \makecell{MAE} &  \makecell{Likelihood-based analysis} \\
             \hline
-            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
-             \\
-            $\tanh^{-1}{(\text{Corr}(\sigma_8,\Omega_m))}$,
-             \\
-            $\sigma_8,\Omega_m$} 
+            \citet{fluri2018cosmological} & \makecell{GNLL} & \makecell{Likelihood-based analysis}
             \\
@@
              \rowcolor{lightgray}
-            \citet{fluri2019cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m, A_{IA}/10, L$  \\ with  $L:\Sigma^{-1}=LL^{T}$}
+                        \citet{fluri2019cosmological} &  \makecell{GNLL} &  \makecell{Likelihood-based analysis}
             \\
             \hline                    
-            \citet{ribli2018improved} & \makecell{CNN} & \makecell{MSE}  &\makecell{RMSE for \\ evaluation}  & \makecell{$\sigma_8, \Omega_m$}   
+                        \citet{ribli2019weak} & \makecell{MAE} & \makecell{Likelihood-based analysis}
             \\            
             \hline             
-            \citet{ribli2019weak} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
-\\            
-\hline             
-            \citet{PhysRevD.102.123506} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
+                        \citet{PhysRevD.102.123506} & \makecell{MAE} & \makecell{Likelihood-based analysis}
             \\
@@
              \rowcolor{lightgray}
-            \citet{jeffrey2021likelihood} & \makecell{\makecell{CNN} \\\makecell{CNN}+NF} & 
+                        \citet{jeffrey2021likelihood} & 
                         \makecell{MSE \\ VMIM}
-            & \makecell{PyDelfi} & \makecell{$\varphi: F_{\varphi}(\bm{d})=\bm{\theta}$ \\ with $\bm{\theta}=\Omega_m, \sigma_8$}    
+                        &  \makecell{Likelihood Free Inference \\ (Py-Delfi)} 
             \\            
             \hline             
-            \citet{fluri2021cosmological} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
+                        \citet{fluri2021cosmological} &\makecell{IMNN} &   \makecell{Likelihood Free Inference \\ (GPABC)}  
             \\
@@
              \rowcolor{lightgray}
-
-            \citet{fluri2022full} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
+                        \citet{fluri2022full} &  \makecell{IMNN} & \makecell{Likelihood Free Inference \\ (GPABC)} 
             \\            
             \hline 
-            \citet{lu2022simultaneously} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8, A_{IA}/10,$ \\ rescaled \\ baryonic parameters \\ $(M_c,M_{1,0}, \eta, \beta)$}
+                        \citet{lu2022simultaneously} & \makecell{MSE} & \makecell{Likelihood-based analysis}   
             \\           
             \hline 
-            \citet{kacprzak2022deeplss} & \makecell{CNN} & \makecell{GNLL}  & \makecell{Likelihood \\ analysis}  & \makecell{WL: $\Sigma, \Omega_m, \sigma_8, A_{IA}, \eta_{IA}$ \\  GC:$\Sigma, \Omega_m, \sigma_8, b_g, r_g,\eta_{b_{g}}$ }
+                        \citet{kacprzak2022deeplss} & \makecell{GNLL}  & \makecell{Likelihood-based analysis}
 
@@
              \rowcolor{lightgray}
-            \citet{lu2023cosmological} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8,A_{IA}/10,$\\ rescaled baryonic \\ parameters: $M_c,M_{1,0}, \eta, \beta$}
+                        \citet{lu2023cosmological} &  \makecell{MSE} & \makecell{Likelihood-based analysis} \\
+
+            \hline      
         \end{tabular}
+        \caption{Table summarizing the different neural compression schemes used for weak lensing applications.
+The table includes all studies conducted within the context of implicit analysis (Likelihood-free) and standard Likelihood-based analysis.
+\\
+        Abbreviations used in the Table: MSE-Mean Square Error; MAE-Maximum Absolute Error; GNLL- Gaussian Negative Log Likelihood; VMIM- Variational Mutual Information Maximization; IMNN- Information Maximizing Neural Networks; GPABC-Gaussian Processes Approximate Bayesian Computation.}
         \label{tab:biblio_survey}
+    \end{center}
 \end{table*}
-\end{center}
 %--------------------------------------------------------------------
 % ############# End Table BiblioSruvey  #############
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
@@
 We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10.
-The forecasted results are obtained using the simulation procedure outlined in and the parameter inference strategy described in .
-We begin by comparing the outcomes of three different inference procedures: Explicit 2-point statistics, Explicit full-field statistics, and Implicit statistics using CNN summaries.
-Subsequently, we analyze the impact of the different compression strategies on the final cosmological constraints.
+The results are obtained using the simulation procedure outlined in \autoref{Sec:the SBILens framework} and the parameter inference strategy described in \autoref{Sec:Inference_strategy}.
+We begin by analyzing the impact of the different compression strategies on the final cosmological constraints.
+Subsequently, we compare the outcomes of three different inference procedures: the 2-point statistics, the explicit full-field statistics, and the implicit full-field statistics using CNN summaries.
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard-two point statistics and two map-based approaches, the explicit inference using an HMC and the implicit inference using the Normalizing Flow as an estimator.
+We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard two-point statistics and two map-based approaches, the explicit inference and the implicit inference strategy.
 As outlined before, our interest is to prove that the two map-based approaches lead to very comparable posterior distributions.
-We present the constraints in the full $\Lambda$CDM parameter space in \autoref{fig:contours_posterior_imp_ex_ps}.
-The contours obtained by the angular $C_{\ell}$ analysis are plotted in blue, the ones for the HMC in yellow, and those for the Normalizing Flow in violet.
-Confirming previous work, we find the two-point statistics are suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$.
-$h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches.
- We quantify the results by computing the Figure of Merit as described in \autoref{Eq:Figure_of_merit}. The results are presented in \autoref{tab:f_o_m}. 
-A further investigation is represented by 
--show the results for the KL divergence 
--present a comparison with other works. 
+We present the $68.3\%$ and $95.5\%$ confident regions from one fiducial map for the full $\Lambda$CDM parameters in \autoref{fig:contours_posterior_imp_ex_ps}.
+The contours obtained by the angular $C_{\ell}$ analysis are plotted in violet, the ones for the explicit inference (HMC) in orange, and those for the implicit inference (Normalizing Flow) in blue.
+We can see from the figure that the contours from the HMC-based inference and the Normalizing flow-based inference are remarkably similar.
+We quantify the results by computing the Figure of Merit defined as follows:
+\begin{equation}\label{Eq:Figure_of_merit}
+    \text{FoM}_{\alpha \beta}=\sqrt{\text{det}(\Tilde{F}_{\alpha \beta})}.
+\end{equation}
+Here, $\alpha$ and $\beta$ represent a pair of cosmological parameters, and $\Tilde{F}_{\alpha \beta}$ refers to the marginalized Fisher matrix.
+We calculate $\Tilde{F}_{\alpha \beta}$ as the inverse of the parameter space covariance matrix $C_{\alpha \beta}$, which is estimated from the Normalizing flow sampling.
+Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalized parameter space of the $\alpha$ and $\beta$ pair.
+The results are presented in \autoref{tab:f_o_m}.
+The remarkably strong agreement between the two posteriors confirms that the two map-based cosmology inference methods yield the same results.
+The ratio of their FoM, corresponding to 1.00, 1.04, 1.03, in the $(\Omega_c-\sigma_8); (\Omega_c-w_0); (\sigma_8-w_0$) plane, establishes the validity of the implicit inference strategy.
+\\
+Confirming previous work, we note that $h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches. 
+Additionally, for the full-field strategies we find that the size of the contours is significantly smaller than the size for the prior distributions adopted, suggesting that the 
+posterior on these parameters are not affected by the specific choice of prior.
+Moreover, we find that the two-point statistic is suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$, while the two map-based approaches yield much tighter constraints on these parameters. 
+We find that the map-based explicit and implicit strategies lead to an improvement in the figure of merit of $2.06\times$, 1.98$\times$, 2.28$\times$ and 2.06$\times$, 1.90$\times$, 2.22$\times$, respectively, in the $(\Omega_c-\sigma_8); (\Omega_c-w_0); (\sigma_8-w_0$) plane.
+%--------------------------------------------------------------------
+\subsection{Optimal compression strategy}
+\autoref{fig:contours_posterior_diff_loss} shows our $68.3\%$ and $95.5\%$ constraints from one fiducial map using different compressed summaries.
+We compare the constraints obtained from MSE (black contours), MAE (magenta contours), and VMIM (blue contours).
+We note that the results obtained using different summaries are generally in agreement with each other, and there are no tensions present for any of the cosmological parameters.
+We report the marginalized summary constraints in \autoref{tab:summaries}.
+The results concern the cosmological parameters that are better constrained from weak lensing: $\Omega_c$, $\sigma_8$, $w_0$.
+We note that the VMIM compressed summary statistics prefer values of $\Omega_c$, $\sigma_8$, and $w_0$ that are closer to our fiducial cosmology than those inferred by MSE and MAE. \\
+To further quantify these outcomes, we consider the Figure of Merit described in \autoref{Eq:Figure_of_merit}. 
+The results are presented in \autoref{tab:f_o_m}.
+We can see that VMIM yields more precise measurements than MSE and MAE for all considered parameters.
+In particular, the FoM of ($\Omega_c$, $\sigma_8$) is improved by $1.24\times$ and $1.1\times$ compared to MSE and MAE, respectively;  the FoM of ($\Omega_c$, $w_0$) is improved by $1.25\times$ and $1.17\times$ from the MSE and MAE; the FoM of ($\sigma_8$, $w_0$) is improved by $1.23\times$ and $1.15\times$ from the MSE and MAE.
+%----------------------
+%--------------------------------------------------------------------
 %--------------------------------------------------------------------
@@
     \centering
-\includegraphics[width=\textwidth]{figures/contours_posterior_imp_ex_ps.pdf}
+        \includegraphics[width=\textwidth]{figures/contours_posterior_imp_ex_ps_dashed.pdf}
         \caption{
         Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup.
-The constraints are obtained by applying the $C_{\ell}$ (light blue contours), the full field explicit inference (yellow contours), and the full field implicit inference strategy (violet contours), described in \autoref{Sec:experiment}.
+The constraints are obtained by applying the $C_{\ell}$ (violet contours), the full field explicit inference (orange contours), and the full field implicit inference strategy (blue dashed contours), described in \autoref{Sec:experiment}.
         The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
         \label{fig:contours_posterior_imp_ex_ps}
+\end{figure*}
+%-----------------------------------------------------------------
+%--------------------------------------------------------------------
+%--------------------------------------------------------------------
+% ############# PLOT CONTOURS DIFFERENT LOSS FUNCTIONS  #############
+%--------------------------------------------------------------------
+\begin{figure*}
+    \centering
+        \includegraphics[width=\textwidth]{figures/contours_posterior_loss_dashed.pdf}
+        \caption{
+        Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup.
+The constraints are obtained from three CNN map compressed statistics: the MSE (black dashed contours), the MAE (magenta dashed contours), and VMIM (blue contours), described in \autoref{Sec:experiment}.
+        The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
+        \label{fig:contours_posterior_diff_loss}
 \end{figure*}
@@
             $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
-
             \hline
         \end{tabular}
+        \caption{ Figure of Merit (FoM) for different inference strategies: the convergence power spectrum $C_{\ell}$, the HMC, the CNN map compressed statistics with the MSE, the VMIM, the MAE and the GNLL loss functions.
+The values of the figure of merit are inversely proportional to the area of the contours; the larger the FoM, the higher the constraining power.}
         \label{tab:f_o_m}
+    \end{center}
 \end{table*}
-\end{center}
 %--------------------------------------------------------------------
@@
 \begin{table*}
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+    \begin{center}
+        \begin{tabular}{|l|l|l|l|l|}
             \hline
-    \multirow{2}{*}{} &
-      \multicolumn{2}{c|}{VMIM} &
-      \multicolumn{2}{c|}{MSE} &
-      \multicolumn{2}{c|}{MAE} &
-      \multicolumn{2}{c|}{GNLL} \\
-    & Mean & $\sigma$ & Mean & $\sigma$ & Mean & $\sigma$  & Mean & $\sigma$ \\
+             & VMIM & MSE & MAE &GNLL \\
             \hline
             \hline
-    $\Omega_c$ & 0.2742 & 0.026 & 0.2830 & 0.030 & 0.2793 & 0.029  & 2.1\% & 2.1\%\\
+            $\Omega_c$ & $0.274^{+0.026}_{-0.025}$  & $0.283^{+0.031}_{-0.029}$  & $0.279^{+0.030}_{-0.028}$ & \\
             \hline
-    $\Omega_b$ & 0.0499 & 0.006 & 0.0492 & 0.006 &  0.0501 & 0.006 & 2.1\% & 2.1\% \\
+            $\Omega_b$ & $ \left( 49.9^{+6.0}_{-6.4} \right) \times 10^{-3}$ &  $\left( 49.2\pm 6.1 \right) \times 10^{-3}$ &  $\left( 50.1^{+5.9}_{-6.1} \right) \times 10^{-3}$ &  \\
             \hline
-    $\sigma_8$ & 0.819 & 0.03 &0.808 & 0.03 &  0.808 &0.03 & 2.1\% & 2.1\% \\
+            $\sigma_8$ & $0.819^{+0.031}_{-0.029}$  & $0.808^{+0.034}_{-0.032}$  &  $0.808^{+0.032}_{-0.030}$ &   \\
             \hline
-    $w_0$      &-0.998 & 0.2 &-1.13 & 0.2 & -1.16& 0.2  & 2.1\% & 2.1\% \\
+            $w_0$      & $-1.00^{+0.20}_{-0.22}$  & $-1.13^{+0.23}_{-0.22}$ &   $-1.16\pm 0.22$  &  \\
             \hline
-    $h_0$      & 0.6658 & 0.058 &0.6620 & 0.057 & 0.6640 & 0.057&  2.1\% & 2.1\% \\
+            $h_0$      & $0.666^{+0.061}_{-0.057}$ & $0.662^{+0.056}_{-0.057}$  & $0.664^{+0.056}_{-0.058}$  & \\
             \hline
-    $n_s$      & 0.9635 & 0.04 & 0.9559 & 0.04 & 0.9547 & 0.04  & 2.1\% & 2.1\% \\
+            $n_s$      &  $0.963^{+0.041}_{-0.038}$ &  $0.956 \pm 0.041$ &  $0.955^{+0.039}_{-0.041}$  & \\
             \hline
         \end{tabular}
+        \caption{Summary of the marginalized parameter distributions.}
+        \label{tab:summaries}
+  \end{center}
 \end{table*}
@@
 %--------------------------------------------------------------------
-\section{Conclusions}\label{Sec:conclusion}
+\section{Summary and conclusions}\label{Sec:conclusion}
+ In this work, we presented and compared two different map-based strategies for inferring cosmological parameters: the explicit full-field strategy, also known as Bayesian Hierarchical inference, based on an HMC sampler, and the implicit inference strategy, also known as Likelihood-Free Inference, based on a neural density estimator.
+
+We start with a consideration: Deep learning approaches for implicit inference typically involve two steps.
+The first step is the automatic learning of an optimal low-dimensional summary statistic, and the second step is the use of a Neural Density Estimator in low dimensions to either build an estimate $P_{\bm{\varphi}}$ of the likelihood function $p(\bm{x}|\bm{\theta})$ (Neural Likelihood Estimation) or build an estimate $P_{\bm{\varphi}}$ of the posterior distribution $p(\bm{\theta}|\bm{x})$ (Neural Posterior Estimation).
+\\
+Now, one can understand that both of these steps may potentially impact the final constraints on the parameters of interest. \\
+Having said that, the main motivation for this work is to evaluate the impact of a given compression strategy on the final posterior distribution and, consequently, determine whether an optimal compression strategy exists.
+Furthermore, the aim is to demonstrate that by using this strategy, both implicit and explicit methods yield the same posterior.
+\\
+To construct the forward model for explicit inference and simulate the mock data required to train the implicit model, we developed \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Jax-based weak lensing simulator optimized for inference applications that need access to the model's derivatives.
+Our analysis is based on synthetic weak lensing data with five tomographic bins, mimicking a survey like LSST-Y10.
+
+After providing an overview of the different compression strategies adopted in the literature for both Likelihood-Free Inference and likelihood-based inference strategies, we compared the impact of some of those on the final constraints on the cosmological parameters for a $\Lambda$CDM model.
+We found the following results:
+\begin{enumerate}
+    \item The marginalized summary statistics indicate that VMIM produces better results for $\Omega_c$, $w_0$, and $\sigma_8$ in terms of agreement with the fiducial value.
+However, it is important to note that the results from MSE and MAE are not in tension with the fiducial parameters.
+    Furthermore, we quantified the outcomes by examining the figure of merit and found that VMIM provides more precise measurements compared to MSE and MAE.
+    \item When using the VMIM to compress the original high-dimensional data, we compared the posterior obtained in the implicit inference framework with those obtained from Bayesian hierarchical modeling and the power spectrum.
+We demonstrate that both map-based approaches lead to a significant improvement in constraining $\Omega_c, w_0, \sigma_8$ compared to the 2-point statistics.
+However, $h, n_s, \Omega_b$ are not constrained by either and are prior-dominated.
+    \item When using the VMIM to compress the original high-dimensional data the two methods, i.e., Bayesian hierarchical inference and Likelihood-free inference, lead to the same posterior distributions.
+\end{enumerate}
+
+It is important to consider the potential limitations of the current implementation and highlight particular strategies for future extensions and applications of this project.
+In this work, we employed a physical model based on a lognormal prior, which is notably faster than simulation-based methods.
+Although we have shown that this description accounts for additional non-Gaussian information, as evidenced by the fact that we obtain different posteriors from the full-field and power spectrum methods, it is important to note that this is a good approximation for the convergence at intermediate scales but may not be appropriate for analyzing small scales.
+Furthermore, the lognormal shift parameters are computed using the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which employs perturbation theory.
+However, as mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales.\footnote{However, as the main objective of this project is to compare the different inference strategies, we are not very concerned with the potential implications of this approximation at this stage.}
+
+Additionally, we did not include any systemic in the current application, although previous studies demonstrated that the map-based approaches help to dramatically improve the constraints on systematic and cosmological parameters in the presence of these last.
+The main reason for this absence is mainly related to the difficulty of modeling systematic effects, like for example intrinsic alignment, in the lognormal description.
+\\
+Hence, the natural next step will be to implement an N-body model as the physical model for the \texttt{SBILens} package. 
+Specifically, we aim to work with the differentiable simulation we have presented in \citet{}.
+As mentioned in our previous work, the simulations can be improved and further developed by including additional systematics such as redshift uncertainties, baryonic feedback, and a more complicated intrinsic alignment model.
+With these improvements, our model will be better suited to handle real cosmic shear data, allowing us to fully maximize the information gained from next-generation surveys.
+
 %--------------------------------------------------------------------
@@
 
-\textbf{Comments:}
-\vspace{5mm}
-
-\begin{itemize}
-    \item \justine{Regarding the table with the differents losses - \citet{2018PhRvD..97j3515G} also used weighted MAE: 'Errors in predictions for maps from cosmologies in sparsely sampled regions are more severely penalized than those for maps from densely sampled re- gions.
-We show in §IV C that such a weighted loss func- tion reduces the bias in the predictions, at the cost of a longer network training, but has only a limited impact on the parameter constraints inferred from the predictions.'
-    I think it's because they don't have honmogeneous dataset so if needed we can added a note to say that we are in the case of homogeneous dataset and thus don't have bias coming from this etc.
-    + for inference startegy maybe we can say that It's gaussian likelihood with fixed cov matrix (evaluated at the fiducials) analysis.
-    \item \citet{fluri2018cosmological} I think they use gaussian log likelihood loss where they learn both the mean and cov matrix: 'Similar to [26], we chose a negative log-likelihood loss as cost function' page 5.
-+ because of bias issues they add regularization term at some point in the training.
-    Then for the inference part they use a gaussian likelihood with mean and covariance matrix como dependent based on their CNN predictions.}
-    \denise{yes, the use gnll, it was just a typo. 
-However, they do not use the covariance matrix from the CNN predictions "
-We did not use the network predictions of the covariance matrices for our main analysis. One reason was that
-our examined test region was consisting out of four independent patches and the input of the network was only
-one patch at a tim"} 
-\justine{Yes you're right!
-They still present the case where you can use the covariance matrix, I mean in some way they try to pay attention to it.
-Idk if you plan to describe a bit this table, if yes I think you can say that they propose this but end up just using the mean which is the exact same as doing mse etc...
-?}
-\denise{yep yep, the table is only a draft, and I didn't mention it in the paper yet. I will definitely describe it in more detail. Ps, sorry for the re-writing, I have OCD with latex error message}
-\justine{yes ok no problem :) I was just checking the table since I'm currently reading these papers for our second project and did not know where to make these comments haha}
-\LM{Ok sweet I was confused about this -- Fluri et al talk about this loss function (which I'm using in another project ahaha) but if they don't make use of the network Covariance it's just a weirdly-weighted MSE (also I wasn't convinced about their training routine in that paper -- the losses didn't seem converged)}
-\item \justine{\cite{fluri2021cosmological} From my understanding, what they use is not exactly the imnn in the sense that they don't rely on gaussian assumptions for the summary distribution.}
-\item \LM{Nope they don't use IMNN.
-They're using a regression-gnll loss, but they don't show any optimality guarantees using this form.
-Plus, they don't make use of the covariance in a particularly smart way (e.g.
-to reweight summary distributions for posterior construction) so it's not super useful.
-I'll be sure to explicitly show the difference in Fluri et al and the IMNN in the corresponding section.}
-\justine{Awesome, thanks! :)}
-\end{itemize}
-
 %--------------------------------------------------------------------
@@
 \section{Mean Square Error (MSE)}\label{Sec:appendix_Mean Square Error}
-In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the mean of the posterior distribution, namely:
+In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the mean of the posterior distribution. Namely:
 \begin{equation}\label{Eq:mean_mse}
-\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}=\operatorname*{argmin}_{\mathcal{F}(\bm{d})}\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\left\Vert \bm {\theta}-\mathcal{F}(\bm{d})
- \right \Vert_{2}^{2}].
+\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{x})}=\operatorname*{argmin}_{F(\bm{x})}\mathbb{E}_{p(\bm {\theta}|\bm {x})}[\left\Vert \bm {\theta}-F(\bm{x})
+ \right \Vert_{2}^{2}],
 \end{equation}
-To demonstrate \autoref{Eq:mean_mse}, we need to minimize the expected value of the $L_2$ norm with respect to $\mathcal{F}(\bm{d})$. Let us consider its derivative:
+where the posterior mean $\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{x})}$, is calculated as follows:
+\begin{equation}\label{Eq:mean_pos}
+     \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{x})}= \mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}].
+\end{equation}
+To demonstrate this statement, we need to minimize the expected value of the $L_2$ norm with respect to $F(\bm{x})$. Let us consider its derivative:
 \begin{align}\label{Eq:moment_1}
-   & \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm{d}))}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2] =  \\
+   & \frac{\partial}{\partial F(\bm{x}) }  \mathbb{E}_{p(\bm {\theta}|\bm{x}))}[(\bm {\theta}-F(\bm{x}))^2] =  \\
     &
-    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2+\mathcal{F}(\bm{d})^2-2\bm {\theta}\mathcal{F}(\bm{d})] = \nonumber \\
+    \frac{\partial}{\partial F(\bm{x}) }  \mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}^2+F(\bm{x})^2-2\bm {\theta}F(\bm{x})] = \nonumber \\
     &
-    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  [
-    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2]+\mathcal{F}(\bm{d})^2-2\mathcal{F}(\bm{d})\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]]= \nonumber \\
-    &2\mathcal{F}(\bm{d})-2 \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. \nonumber
+    \frac{\partial}{\partial F(\bm{x}) }  [
+    \mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}^2]+F(\bm{x})^2-2F(\bm{x})\mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}]]= \nonumber \\
+    &2F(\bm{x})-2 \mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}]. \nonumber
 \end{align}
 Setting it equal to zero, we obtain the critical value:
-\begin{equation}
-    \mathcal{F}(\bm{d})= \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. 
+\begin{equation}\label{Eq:minimum_mean}
+    F(\bm{x})= \mathbb{E}_{p(\bm {\theta}|\bm {x})}[\bm {\theta}]. 
 \end{equation}
 Considering the second-order derivative:
-\begin{equation}\label{Eq:minimum_mean}
-   \frac{\partial^2}{\partial^2 \mathcal{F}(\bm{d})}  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2]=2, 
+\begin{equation}
+   \frac{\partial^2}{\partial^2 F(\bm{x})}  \mathbb{E}_{p(\bm {\theta}|\bm {x})}[(\bm {\theta}-F(\bm{x}))^2]=2>0, 
 \end{equation}
-we can assert that the critical value $\mathcal{F(\bm(\theta))}$ is also a minimum. 
-Since 
-\begin{equation}
-    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}|\bm{d}]= \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}
-\end{equation}
-it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
+we can assert that this critical value is also a minimum. \\
+From \autoref{Eq:minimum_mean} and \autoref{Eq:mean_pos}, we obtain \autoref{Eq:mean_mse}.
 %--------------------------------------------------------------------
 \section{Maximum Absolute Error (MAE)}\label{Sec:appendix_Maximum Absolute Error}
-In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the median of the posterior distribution.
-By definition, the median of a probability density function $p(x)$ is a real number $m$ that satisfies:
+In this section, we demonstrate that minimizing the $L_1$ norm is equivalent to training the model to estimate the median of the posterior distribution.
+Namely:
+\begin{equation}\label{Eq:median_mae}
+ \bm {\theta}^M_{p(\bm {\theta}|\bm{x})}=\operatorname*{argmin}_{F(\bm{x})}\mathbb{E}_{p(\bm {\theta}|\bm {x})}[| \bm {\theta}-F(\bm{x})|].
+\end{equation}
+By definition, the median of a one-dimensional\footnote{For notational simplicity, we demonstrate this statement in one dimension; the generalization to N-dimensions is straightforward.} probability density function $p(x)$ is a real number $m$ that satisfies:
 \begin{equation}\label{Eq:definition_median}
 \int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
-\end{equation}. 
-The expectation value of the mean absolute error is defined as:
+\end{equation}
+The expectation value of the maximum absolute error is defined as:
 \begin{equation}
-    \mathbb{E}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
+    \mathbb{E}_{p(x)}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
 \end{equation}
@@
 we obtain \autoref{Eq:definition_median}.
-% \begin{equation}
-% \int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
-% \end{equation}
 %--------------------------------------------------------------------

### 2023-09-08 script/job_hmc_bhm.sh
rearrangement script folder

@@ -1,12 +1,13 @@
 #!/bin/bash
 #SBATCH -A ykz@v100
-#SBATCH --job-name=name_of_the_job         
+#SBATCH --job-name=name_of_the_job  
 #SBATCH --output=name_of_the_job%j.out     
 #SBATCH --error=name_of_the_job%j.out     
+#SBATCH --array=21-31
 #SBATCH --constraint v100-32g
-#SBATCH --qos=qos_gpu-t4
+#SBATCH --qos=qos_gpu-t3
 #SBATCH --hint=nomultithread        
-#SBATCH -t 30:00:00
+#SBATCH -t 20:00:00
 #SBATCH --nodes=1
 #SBATCH --gpus-per-task=1
 #SBATCH --ntasks-per-node=1
@@ -16,4 +17,4 @@
 module purge
 module load tensorflow-gpu/py3/2.7.0
 
-python mcmc_ps.py  
+python hmc_bhm.py  --seed=$SLURM_ARRAY_TASK_ID --filename=job_$SLURM_ARRAY_TASK_ID

### 2023-09-08 script/job_hmc_ps.sh

@@ -1,13 +1,12 @@
 #!/bin/bash
 #SBATCH -A ykz@v100
-#SBATCH --job-name=name_of_the_job  
+#SBATCH --job-name=name_of_the_job         
 #SBATCH --output=name_of_the_job%j.out     
 #SBATCH --error=name_of_the_job%j.out     
-#SBATCH --array=21-31
 #SBATCH --constraint v100-32g
-#SBATCH --qos=qos_gpu-t3
+#SBATCH --qos=qos_gpu-t4
 #SBATCH --hint=nomultithread        
-#SBATCH -t 20:00:00
+#SBATCH -t 30:00:00
 #SBATCH --nodes=1
 #SBATCH --gpus-per-task=1
 #SBATCH --ntasks-per-node=1
@@ -17,4 +16,4 @@
 module purge
 module load tensorflow-gpu/py3/2.7.0
 
-python hmc_ff.py  --seed=$SLURM_ARRAY_TASK_ID --filename=job_$SLURM_ARRAY_TASK_ID
+python hmc_ps.py  

### 2023-09-08 script/train_estimator.py

@@ -1,301 +1,0 @@
-from functools import partial
-from tqdm import tqdm
-import pickle
-import argparse
-from pathlib import Path
-
-import numpy as np
-import jax
-import jax.numpy as jnp
-
-from haiku._src.nets.resnet import ResNet34, ResNet18
-import optax
-import haiku as hk
-
-import tensorflow_datasets as tfds
-import tensorflow as tf
-
-import tensorflow_probability as tfp; tfp = tfp.experimental.substrates.jax
-tfb = tfp.bijectors
-tfd = tfp.distributions
-
-from sbi_lens.gen_dataset.lensing_lognormal_dataset import LensingLogNormalDataset
-from sbi_lens.gen_dataset.utils import augmentation_noise, augmentation_flip
-from sbi_lens.normflow.models import AffineSigmoidCoupling
-from sbi_lens.normflow.models import (
-  ConditionalRealNVP,
-  AffineCoupling
-)
-from sbi_lens.normflow.train_model import TrainModel
-from sbi_lens.config import config_lsst_y_10
-
-###################
-import logging
-import tensorflow_probability as tfp; tfp = tfp.substrates.jax
-
-tfp.distributions.TransformedDistribution(
-    tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity()
-)
-
-logger = logging.getLogger("root")
-
-
-class CheckTypesFilter(logging.Filter):
-    def filter(self, record):
-        return "check_types" not in record.getMessage()
-
-
-logger.addFilter(CheckTypesFilter())
-###################
-
-
-# script arguments
-parser = argparse.ArgumentParser()
-parser.add_argument("--total_steps", type=int, default=9000)
-parser.add_argument("--resnet", type=str, default='resnet18')
-parser.add_argument("--loss", type=str, default='train_compressor_vmim')
-
-args = parser.parse_args()
-
-SOURCE_FILE = Path(__file__)
-SOURCE_DIR = SOURCE_FILE.parent
-ROOT_DIR = SOURCE_DIR.parent.resolve()
-DATA_DIR = ROOT_DIR / "data"
-
-print('######## Data direct ########', ROOT_DIR /'notebook/tensorflow_dataset' )
-print('######## CONFIG LSST Y 10 ########')
-
[…]

### 2023-09-08 script/train_estimator_array.py

@@ -1,301 +1,0 @@
-from functools import partial
-from tqdm import tqdm
-import pickle
-import argparse
-from pathlib import Path
-
-import numpy as np
-import jax
-import jax.numpy as jnp
-
-from haiku._src.nets.resnet import ResNet34, ResNet18
-import optax
-import haiku as hk
-from numpyro.handlers import seed, condition, trace
-
-import tensorflow_datasets as tfds
-import tensorflow as tf
-
-import tensorflow_probability as tfp; tfp = tfp.experimental.substrates.jax
-tfb = tfp.bijectors
-tfd = tfp.distributions
-
-from sbi_lens.gen_dataset.lensing_lognormal_dataset import LensingLogNormalDataset
-from sbi_lens.gen_dataset.utils import augmentation_noise, augmentation_flip
-
-from sbi_lens.normflow.models import AffineSigmoidCoupling
-from sbi_lens.normflow.models import (
-  ConditionalRealNVP,
-  AffineCoupling
-)
-from sbi_lens.normflow.train_model import TrainModel
-from sbi_lens.config import config_lsst_y_10
-from sbi_lens.simulator import lensingLogNormal
-
-###################
-import logging
-import tensorflow_probability as tfp; tfp = tfp.substrates.jax
-
-tfp.distributions.TransformedDistribution(
-    tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity()
-)
-
-logger = logging.getLogger("root")
-
-
-class CheckTypesFilter(logging.Filter):
-    def filter(self, record):
-        return "check_types" not in record.getMessage()
-
-
-logger.addFilter(CheckTypesFilter())
-###################
-
-
-# script arguments
-parser = argparse.ArgumentParser()
-parser.add_argument("--total_steps", type=int, default=9000)
-parser.add_argument("--resnet", type=str, default='resnet18')
-parser.add_argument("--loss", type=str, default='train_compressor_vmim')
-parser.add_argument("--seed", type=int, default=4)
-parser.add_argument("--filename", type=str, default='res' )
-
-args = parser.parse_args()
-
-SOURCE_FILE = Path(__file__)
-SOURCE_DIR = SOURCE_FILE.parent
[…]

### 2023-09-08 script/job_array_estimetor.sh
script-folder useless job rm

@@ -1,20 +1,0 @@
-#!/bin/bash
-#SBATCH -A ykz@v100
-#SBATCH --array=0-10
-#SBATCH --job-name=name_of_the_job         
-#SBATCH --output=name_of_the_job%j.out     
-#SBATCH --error=name_of_the_job%j.out     
-#SBATCH --constraint v100-32g
-#SBATCH --qos=qos_gpu-t3
-#SBATCH --hint=nomultithread        
-#SBATCH -t 08:00:00
-#SBATCH --nodes=1
-#SBATCH --gpus-per-task=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --gres=gpu:1 
-#SBATCH --cpus-per-task=20
-
-module purge
-module load python cudnn/8.5.0.96-11.7-cuda
-
-python /gpfsdswork/projects/rech/ykz/ulm75uc/VMIM-vs-MSE-/script/train_estimator_array.py --total_steps=9000  --resnet='resnet18' --loss='train_compressor_mse' --seed=$SLURM_ARRAY_TASK_ID --filename=job_$SLURM_ARRAY_TASK_ID

### 2023-09-08 script/job_estimetor.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-#SBATCH -A ykz@v100
-#SBATCH --job-name=name_of_the_job         
-#SBATCH --output=name_of_the_job%j.out     
-#SBATCH --error=name_of_the_job%j.out     
-#SBATCH --constraint v100-32g
-#SBATCH --qos=qos_gpu-t3
-#SBATCH --hint=nomultithread        
-#SBATCH -t 08:00:00
-#SBATCH --nodes=1
-#SBATCH --gpus-per-task=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --gres=gpu:1 
-#SBATCH --cpus-per-task=20
-
-module purge
-module load python cudnn/8.5.0.96-11.7-cuda
-
-python /gpfsdswork/projects/rech/ykz/ulm75uc/VMIM-vs-MSE-/script/train_estimator.py --total_steps=9000  --resnet='resnet18' --loss='train_compressor_vmim'

### 2023-10-26 script/dev/job_array_estimetor.sh
backup of jz folder

@@ -1,0 +1,20 @@
+#!/bin/bash
+#SBATCH -A ykz@v100
+#SBATCH --array=0-10
+#SBATCH --job-name=name_of_the_job         
+#SBATCH --output=name_of_the_job%j.out     
+#SBATCH --error=name_of_the_job%j.out     
+#SBATCH --constraint v100-32g
+#SBATCH --qos=qos_gpu-t3
+#SBATCH --hint=nomultithread        
+#SBATCH -t 08:00:00
+#SBATCH --nodes=1
+#SBATCH --gpus-per-task=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --gres=gpu:1 
+#SBATCH --cpus-per-task=20
+
+module purge
+module load python cudnn/8.5.0.96-11.7-cuda
+
+python /gpfsdswork/projects/rech/ykz/ulm75uc/VMIM-vs-MSE-/script/train_estimator_array.py --total_steps=9000  --resnet='resnet18' --loss='train_compressor_mse' --seed=$SLURM_ARRAY_TASK_ID --filename=job_$SLURM_ARRAY_TASK_ID

### 2023-10-26 script/dev/job_estimetor.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+#SBATCH -A ykz@v100
+#SBATCH --job-name=name_of_the_job         
+#SBATCH --output=name_of_the_job%j.out     
+#SBATCH --error=name_of_the_job%j.out     
+#SBATCH --constraint v100-32g
+#SBATCH --qos=qos_gpu-t3
+#SBATCH --hint=nomultithread        
+#SBATCH -t 08:00:00
+#SBATCH --nodes=1
+#SBATCH --gpus-per-task=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --gres=gpu:1 
+#SBATCH --cpus-per-task=20
+
+module purge
+module load python cudnn/8.5.0.96-11.7-cuda
+
+python /gpfsdswork/projects/rech/ykz/ulm75uc/VMIM-vs-MSE-/script/train_estimator.py --total_steps=9000  --resnet='resnet18' --loss='train_compressor_vmim'

### 2023-10-26 script/dev/train_estimator.py

@@ -1,0 +1,301 @@
+from functools import partial
+from tqdm import tqdm
+import pickle
+import argparse
+from pathlib import Path
+
+import numpy as np
+import jax
+import jax.numpy as jnp
+
+from haiku._src.nets.resnet import ResNet34, ResNet18
+import optax
+import haiku as hk
+
+import tensorflow_datasets as tfds
+import tensorflow as tf
+
+import tensorflow_probability as tfp; tfp = tfp.experimental.substrates.jax
+tfb = tfp.bijectors
+tfd = tfp.distributions
+
+from sbi_lens.gen_dataset.lensing_lognormal_dataset import LensingLogNormalDataset
+from sbi_lens.gen_dataset.utils import augmentation_noise, augmentation_flip
+from sbi_lens.normflow.models import AffineSigmoidCoupling
+from sbi_lens.normflow.models import (
+  ConditionalRealNVP,
+  AffineCoupling
+)
+from sbi_lens.normflow.train_model import TrainModel
+from sbi_lens.config import config_lsst_y_10
+
+###################
+import logging
+import tensorflow_probability as tfp; tfp = tfp.substrates.jax
+
+tfp.distributions.TransformedDistribution(
+    tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity()
+)
+
+logger = logging.getLogger("root")
+
+
+class CheckTypesFilter(logging.Filter):
+    def filter(self, record):
+        return "check_types" not in record.getMessage()
+
+
+logger.addFilter(CheckTypesFilter())
+###################
+
+
+# script arguments
+parser = argparse.ArgumentParser()
+parser.add_argument("--total_steps", type=int, default=9000)
+parser.add_argument("--resnet", type=str, default='resnet18')
+parser.add_argument("--loss", type=str, default='train_compressor_vmim')
+
+args = parser.parse_args()
+
+SOURCE_FILE = Path(__file__)
+SOURCE_DIR = SOURCE_FILE.parent
+ROOT_DIR = SOURCE_DIR.parent.resolve()
+DATA_DIR = ROOT_DIR / "data"
+
+print('######## Data direct ########', ROOT_DIR /'notebook/tensorflow_dataset' )
+print('######## CONFIG LSST Y 10 ########')
+
[…]

### 2023-10-26 script/dev/train_estimator_array.py

@@ -1,0 +1,301 @@
+from functools import partial
+from tqdm import tqdm
+import pickle
+import argparse
+from pathlib import Path
+
+import numpy as np
+import jax
+import jax.numpy as jnp
+
+from haiku._src.nets.resnet import ResNet34, ResNet18
+import optax
+import haiku as hk
+from numpyro.handlers import seed, condition, trace
+
+import tensorflow_datasets as tfds
+import tensorflow as tf
+
+import tensorflow_probability as tfp; tfp = tfp.experimental.substrates.jax
+tfb = tfp.bijectors
+tfd = tfp.distributions
+
+from sbi_lens.gen_dataset.lensing_lognormal_dataset import LensingLogNormalDataset
+from sbi_lens.gen_dataset.utils import augmentation_noise, augmentation_flip
+
+from sbi_lens.normflow.models import AffineSigmoidCoupling
+from sbi_lens.normflow.models import (
+  ConditionalRealNVP,
+  AffineCoupling
+)
+from sbi_lens.normflow.train_model import TrainModel
+from sbi_lens.config import config_lsst_y_10
+from sbi_lens.simulator import lensingLogNormal
+
+###################
+import logging
+import tensorflow_probability as tfp; tfp = tfp.substrates.jax
+
+tfp.distributions.TransformedDistribution(
+    tfp.distributions.Normal(0.0, 1.0), tfp.bijectors.Identity()
+)
+
+logger = logging.getLogger("root")
+
+
+class CheckTypesFilter(logging.Filter):
+    def filter(self, record):
+        return "check_types" not in record.getMessage()
+
+
+logger.addFilter(CheckTypesFilter())
+###################
+
+
+# script arguments
+parser = argparse.ArgumentParser()
+parser.add_argument("--total_steps", type=int, default=9000)
+parser.add_argument("--resnet", type=str, default='resnet18')
+parser.add_argument("--loss", type=str, default='train_compressor_vmim')
+parser.add_argument("--seed", type=int, default=4)
+parser.add_argument("--filename", type=str, default='res' )
+
+args = parser.parse_args()
+
+SOURCE_FILE = Path(__file__)
+SOURCE_DIR = SOURCE_FILE.parent
[…]

### 2023-10-26 script/job_compressor_gnll.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+#SBATCH -A ykz@v100
+#SBATCH --job-name=name_of_the_job         
+#SBATCH --output=name_of_the_job%j.out     
+#SBATCH --error=name_of_the_job%j.out     
+#SBATCH --constraint v100-32g
+#SBATCH --qos=qos_gpu-t4
+#SBATCH --hint=nomultithread        
+#SBATCH -t 60:00:00
+#SBATCH --nodes=1
+#SBATCH --gpus-per-task=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --gres=gpu:1 
+#SBATCH --cpus-per-task=20
+
+module purge
+module load python cudnn/8.5.0.96-11.7-cuda
+
+python /gpfsdswork/projects/rech/ykz/ulm75uc/VMIM-vs-MSE-/script/train_compressor_gnll.py
+
+

### 2023-10-27 paper/paper.tex
change disp sections
27/10

@@
 \section{Introduction}
-Weak gravitational lensing by the large-scale structure is caused by the presence of foreground matter bending the light emitted by background galaxies.
+Weak gravitational lensing by large-scale structures (LSS) is caused by the presence of foreground matter bending the light emitted by background galaxies.
 Being sensitive to the large-scale structure of the universe, it is one of the most promising tools for investigating the nature of dark energy, the origin of the accelerating expansion of the universe and estimating cosmological parameters.
 Future cosmological surveys, such as the Legacy Survey of Space and Time (LSST) of the Vera C.
-Rubin Observatory \citep{ivezic2019lsst}, the Nancy Grace Roman Space Telescope \citep{spergel2015wide}, and the Euclid Mission \citep{laureijs2011euclid}, will rely on weak gravitational lensing as one of the principal physical probes to address unresolved questions in current cosmology.
+Rubin Observatory \citep{ivezic2019lsst}, Nancy Grace Roman Space Telescope \citep{spergel2015wide}, and the Euclid Mission \citep{laureijs2011euclid}, will rely on weak gravitational lensing as one of the principal physical probes to address unresolved questions in current cosmology.
  As these surveys become deeper, they will be able to access more non-Gaussian features of the matter fields.
-This makes the standard weak lensing analyses, which rely on two-point statistics such as the 2-point shear correlation or the angular power spectrum, sub-optimal.
+This makes the standard weak-lensing analyses, which rely on two-point statistics such as the 2-point shear correlation or the angular power spectrum, sub-optimal.
 These analyses are unable to fully capture the non-Gaussian information imprinted in the lensing signal and can only access Gaussian information.
 
-To overcome this limitation, several higher-order statistics have been introduced.
-These include weak lensing peak counts \citep{liu2015cosmology,  liu2015cosmological, lin2015new, kacprzak2016cosmology, peel2017cosmological, shan2018kids, martinet2018kids, ajani2020constraining, harnois2021cosmic, zurcher2022dark}, wavelet and scattering transform \citep{ajani2021starlet, cheng2021weak}, the one point PDF \citep{liu2019constraining, uhlemann2020fisher, boyle2021nuw}, Minkowski functionals \citep{kratochvil2012probing, petri2013cosmology}, moments of mass maps \citep{gatti2021dark},  and 3 point statistics \citep{takada2004cosmological, semboloni2011weak, rizzato2019tomographic, halder2021integrated}.
+To overcome this limitation, several higher order statistics have been introduced.
+These include weak-lensing peak counts \citep{liu2015cosmology,  liu2015cosmological, lin2015new, kacprzak2016cosmology, peel2017cosmological, shan2018kids, martinet2018kids, ajani2020constraining, harnois2021cosmic, zurcher2022dark}, wavelet and scattering transform \citep{ajani2021starlet, cheng2021weak}, the one-point probability distribution function \citep[PDF][]{liu2019constraining, uhlemann2020fisher, boyle2021nuw}, Minkowski functionals \citep{kratochvil2012probing, petri2013cosmology}, moments of mass maps \citep{gatti2021dark},  and three-point  point statistics \citep{takada2004cosmological, semboloni2011weak, rizzato2019tomographic, halder2021integrated}.
 Although these methods have proven to improve cosmological constraints, they rely on summary statistics, for which an analytical description is often lacking.
@@
 To address the need for result validation in the LFI context, we employ a Bayesian forward modeling approach that enables inference on the joint posterior of the cosmological parameters and the convergence field.
-The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Python package designed for weak lensing inference with differentiable simulator implemented in Jax.
+The forward model is based on \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Python package designed for weak-lensing inference with differentiable simulator implemented in Jax.
 \texttt{SBILens} allows the sampling of convergence maps in a tomographic setting, accounting for the cross-correlation between different redshift bins.
 
-The paper is structured as follows: in \autoref{Sec:Motivation}, we illustrate the motivation behind this work.
-In \autoref{Sec:the SBILens framework}, we introduce the \texttt{SBILens} framework and describe the simulated data used in this work.
-In \autoref{Sec:experiment}, we detail the inference strategy and the three different approaches we used: the power spectrum, the map-based inference based on the Bayesian forward-modeling, and the map-based inference based on the LFI.
+The paper is structured as follows: in Section \ref{Sec:Motivation}, we illustrate the motivation behind this work.
+In Section \ref{Sec:the SBILens framework}, we introduce the \texttt{SBILens} framework and describe the simulated data used in this work.
+In Section \ref{Sec:experiment}, we detail the inference strategy and the three different approaches we used: the power spectrum, the map-based inference based on the Bayesian forward-modeling, and the map-based inference based on the LFI.
 In the same section, we also provide a detailed overview of different neural compression strategies.
-In \autoref{Sec:results}, we discuss the results and validate the LFI approaches.
-Finally, we conclude in \autoref{Sec:conclusion}.
+In Section \ref{Sec:results}, we discuss the results and validate the LFI approaches.
+Finally, we conclude in Section \ref{Sec:conclusion}.
 %------------------------------------------------------------------
@@
 \begin{equation}
-        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x},\bm{z}|\bm{\theta}) p(\bm{z}|\bm{\theta}).
+        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x}|\bm{z},\bm{\theta}) p(\bm{z}|\bm{\theta})p(\bm{\theta}).
 \end{equation}
@@
 
-Therefore, our primary focus will be to investigate the effectiveness and validity of these compression techniques in the context of weak lensing applications (within the limits of our setting).
+Therefore, our primary focus will be to investigate the effectiveness and validity of these compression techniques in the context of weak-lensing applications (within the limits of our setting).
 
@@
 \section{The SBILens framework}\label{Sec:the SBILens framework}
-To investigate the questions above, we have developed the Python package \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\texttt{SBILens}}, a weak lensing differentiable simulator based on a log-normal model.
+To investigate the questions above, we have developed the Python package \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\texttt{SBILens}}, a weak-lensing differentiable simulator based on a log-normal model.
 \texttt{SBILens} enables the sampling of convergence maps in a tomographic setting while considering the cross-correlation between different redshift bins.
@@
 \begin{equation}\label{power_spectrum_definition}
-    \left \langle \tilde{\kappa}^{(i)}_{ln} (\ell)\tilde{\kappa}^{(j)}_{ln}(\ell')\right \rangle =C^{ij}_{ln}(\ell)\delta^{K}(\ell-\ell')
+    \left \langle \tilde{\kappa}^{(i)}_{ln} (\ell)\tilde{\kappa}^{*(j)}_{ln}(\ell')\right \rangle =C^{ij}_{ln}(\ell)\delta^{K}(\ell-\ell')
 \end{equation}
@@
 An example of a tomographic convergence map simulated using the \texttt{SBILens} package is shown in \autoref{fig:convergence_maps}.
-\denise{TODO: change disp images}
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\begin{figure}
+\begin{figure*}
     \begin{center}
-    \includegraphics[width=\textwidth]{figures/bin1and2.pdf}
-    \includegraphics[width=\textwidth]{figures/bin3.pdf}
-    \includegraphics[width=0.5\textwidth]{figures/bin5.pdf}
+    \includegraphics[width=\textwidth]{figures/example_kmaps.pdf}
     \caption{
-     Example of convergence map simulated using the \texttt{SBILens} package.
+     Example of convergence maps simulated using the \texttt{SBILens} package.
     }
@@
      \end{center}
-\end{figure}
+\end{figure*}
 %--------------------------------------------------------------------
@@
     Convergence power spectra for different tomographic bin combinations.
-The solid yellow line shows the measurement from 20 simulated maps using the survey setting described in \autoref{Sec:the SBILens framework}, while the black dashed line shows the theoretical predictions computed using jax-cosmo.
+The solid yellow line shows the measurement from 20 simulated maps using the survey setting described in Section \ref{Sec:the SBILens framework}, while the black dashed line shows the theoretical predictions computed using jax-cosmo.
 In this figure, the shaded regions represent the standard deviation from 20 independent map realizations.
@@
 Additionally, we conduct a power spectrum study.
-Indeed, as discussed in \autoref{Sec:Lognormal Modeling}, log-normal fields offer the advantage of rapidly generating convergent fields while accounting for non-Gaussianities.
+Indeed, as discussed in Section \ref{Sec:Lognormal Modeling}, log-normal fields offer the advantage of rapidly generating convergent fields while accounting for non-Gaussianities.
 To emphasize this claim further, along with the full-field analysis, we include a power spectrum analysis.
@@
 %--------------------------------------------------------------------
-To construct the explicit map-based inference strategy in the Hierarchical Bayesian framework, we built a likelihood based on the data model described in \autoref{Sec:Lognormal Modeling}. 
+To construct the explicit map-based inference strategy in the Hierarchical Bayesian framework, we built a likelihood based on the data model described in Section \ref{Sec:Lognormal Modeling}. 
 As mentioned before, this means that the simulator serves as the physical model capable of generating the non-linear representation of the convergence map.
@@
 Indeed, minimizing the $L_{2}$ norm is equivalent to training the model to estimate the mean of the posterior distribution.
-We prove this statement in \autoref{Sec:appendix_Mean Square Error}.
+We prove this statement in Section \ref{Sec:appendix_Mean Square Error}.
 %--------------------------------------------------------------------
@@
 where $\bm{t}$ represents the summary statistics and $\bm{\theta}$ denotes the cosmological parameters.
-In \autoref{Sec:appendix_Maximum Absolute Error}, we demonstrate that minimizing this loss function is equivalent to training the model to estimate the median of the posterior distribution.
+In Section \ref{Sec:appendix_Maximum Absolute Error}, we demonstrate that minimizing this loss function is equivalent to training the model to estimate the median of the posterior distribution.
 While extensively employed in various previous studies \citep{2018PhRvD..97j3515G, fluri2018cosmological, ribli2019weak}, similar to the previous consideration for the MSE, it is important to note that the effectiveness of this loss function in extracting sufficient statistics needs to be demonstrated.
@@
 
-An application of this method in the context of weak lensing can be traced back to the work of \citet{fluri2021cosmological, fluri2022full}.
+An application of this method in the context of weak-lensing can be traced back to the work of \citet{fluri2021cosmological, fluri2022full}.
 Following \citet{charnock2018automatic}, they presented a compression approach that relies on the optimization of the Cramér-Rao bound.
@@
 %--------------------------------------------------------------------
-\subsection{Compression strategy}
+\subsubsection{Compression strategy}
 %--------------------------------------------------------------------
@@
 %--------------------------------------------------------------------
-\subsubsection{Inference strategy}\label{Sec:Inference_strategy}
+\subsection{Inference strategy}\label{Sec:Inference_strategy}
 %--------------------------------------------------------------------
@@
         \end{tabular}
-        \caption{Table summarizing the different neural compression schemes used for weak lensing applications.
+        \caption{Table summarizing the different neural compression schemes used for weak-lensing applications.
 The table includes all studies conducted within the context of implicit analysis (Likelihood-free) and standard Likelihood-based analysis.
@@
 We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10.
-The results are obtained using the simulation procedure outlined in \autoref{Sec:the SBILens framework} and the parameter inference strategy described in \autoref{Sec:Inference_strategy}.
+The results are obtained using the simulation procedure outlined in Section \ref{Sec:the SBILens framework} and the parameter inference strategy described in Section \ref{Sec:Inference_strategy}.
 We begin by analyzing the impact of the different compression strategies on the final cosmological constraints.
@@
 %--------------------------------------------------------------------
-We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard two-point statistics and two map-based approaches, the explicit inference and the implicit inference strategy.
+We now compare the constraining power of the three different approaches described in Section \ref{Sec:experiment}: the standard two-point statistics and two map-based approaches, the explicit inference and the implicit inference strategy.
 As outlined before, our interest is to prove that the two map-based approaches lead to very comparable posterior distributions.
@@
 We report the marginalized summary constraints in \autoref{tab:summaries}.
-The results concern the cosmological parameters that are better constrained from weak lensing: $\Omega_c$, $\sigma_8$, $w_0$.
+The results concern the cosmological parameters that are better constrained from weak-lensing: $\Omega_c$, $\sigma_8$, $w_0$.
 We note that the VMIM compressed summary statistics prefer values of $\Omega_c$, $\sigma_8$, and $w_0$ that are closer to our fiducial cosmology than those inferred by MSE and MAE. \\
@@
         Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup.
-The constraints are obtained by applying the $C_{\ell}$ (violet contours), the full field explicit inference (orange contours), and the full field implicit inference strategy (blue dashed contours), described in \autoref{Sec:experiment}.
+The constraints are obtained by applying the $C_{\ell}$ (violet contours), the full field explicit inference (orange contours), and the full field implicit inference strategy (blue dashed contours), described in Section \ref{Sec:experiment}.
         The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
@@
         Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup.
-The constraints are obtained from three CNN map compressed statistics: the MSE (black dashed contours), the MAE (magenta dashed contours), and VMIM (blue contours), described in \autoref{Sec:experiment}.
+The constraints are obtained from three CNN map compressed statistics: the MSE (red contours), the MAE (black dashed contours), VMIM (yellow contours), the GNLL (blue dotted contours), described in Section \ref{Sec:experiment}.
         The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
@@
             \hline\hline
-            $\Omega_c-  \sigma_8$ & 1222 & 2520 & 2526 & 2043 & 2316\\
-            $\Omega_c-  w_0$      & 100  & 198  & 190  & 152  & 162\\
-            $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
+            $\Omega_c-  \sigma_8$ & 1222 & 2520 & 2526 & 2043 & 2316 & 1882 \\
+            $\Omega_c-  w_0$      & 100  & 198  & 190  & 152  & 162  & 178  \\
+            $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149  & 153  \\
             \hline
@@
             \hline
-            $\Omega_c$ & $0.274^{+0.026}_{-0.025}$  & $0.283^{+0.031}_{-0.029}$  & $0.279^{+0.030}_{-0.028}$ & \\
+            $\Omega_c$ & $0.274^{+0.026}_{-0.025}$  & $0.283^{+0.031}_{-0.029}$  & $0.279^{+0.030}_{-0.028}$ & $0.275^{+0.029}_{-0.025}$\\
             \hline
-            $\Omega_b$ & $ \left( 49.9^{+6.0}_{-6.4} \right) \times 10^{-3}$ &  $\left( 49.2\pm 6.1 \right) \times 10^{-3}$ &  $\left( 50.1^{+5.9}_{-6.1} \right) \times 10^{-3}$ &  \\
+            $\Omega_b$ & $ \left( 49.9^{+6.0}_{-6.4} \right) \times 10^{-3}$ &  $\left( 49.2\pm 6.1 \right) \times 10^{-3}$ &  $\left( 50.1^{+5.9}_{-6.1} \right) \times 10^{-3}$ & $\left( 49.5^{+5.7}_{-6.2} \right) \times 10^{-3}$ \\
             \hline
-            $\sigma_8$ & $0.819^{+0.031}_{-0.029}$  & $0.808^{+0.034}_{-0.032}$  &  $0.808^{+0.032}_{-0.030}$ &   \\
+            $\sigma_8$ & $0.819^{+0.031}_{-0.029}$  & $0.808^{+0.034}_{-0.032}$  &  $0.808^{+0.032}_{-0.030}$ &  $0.834^{+0.036}_{-0.035}$ \\
             \hline
-            $w_0$      & $-1.00^{+0.20}_{-0.22}$  & $-1.13^{+0.23}_{-0.22}$ &   $-1.16\pm 0.22$  &  \\
+            $w_0$      & $-1.00^{+0.20}_{-0.22}$  & $-1.13^{+0.23}_{-0.22}$ &   $-1.16 \pm 0.22$  &  $-0.73^{+0.19}_{-0.22}$ \\
             \hline
-            $h_0$      & $0.666^{+0.061}_{-0.057}$ & $0.662^{+0.056}_{-0.057}$  & $0.664^{+0.056}_{-0.058}$  & \\
+            $h_0$      & $0.666^{+0.061}_{-0.057}$ & $0.662^{+0.056}_{-0.057}$  & $0.664^{+0.056}_{-0.058}$  & $0.670^{+0.059}_{-0.061}$\\
             \hline
-            $n_s$      &  $0.963^{+0.041}_{-0.038}$ &  $0.956 \pm 0.041$ &  $0.955^{+0.039}_{-0.041}$  & \\
+            $n_s$      &  $0.963^{+0.041}_{-0.038}$ &  $0.956 \pm 0.041$ &  $0.955^{+0.039}_{-0.041}$  & $0.954\pm 0.042$\\
             \hline
@@
 \\
-To construct the forward model for explicit inference and simulate the mock data required to train the implicit model, we developed \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Jax-based weak lensing simulator optimized for inference applications that need access to the model's derivatives.
-Our analysis is based on synthetic weak lensing data with five tomographic bins, mimicking a survey like LSST-Y10.
+To construct the forward model for explicit inference and simulate the mock data required to train the implicit model, we developed \href{https://github.com/DifferentiableUniverseInitiative/sbi_lens}{\url{SBILens}}, a Jax-based weak-lensing simulator optimized for inference applications that need access to the model's derivatives.
+Our analysis is based on synthetic weak-lensing data with five tomographic bins, mimicking a survey like LSST-Y10.
 
